### 2025-02-05 main.tex
Create files

%\documentclass[onecolumn]{aastex631}
\documentclass[linenumbers, onecolumn]{aastex631}

%\defcitealias{H21}{H21}

\newcommand{\ra}{\mathrm{RA}}
\newcommand{\dec}{\mathrm{Dec}}
\newcommand{\lsst}{\textit{LSST}}
\newcommand{\gaia}{\textit{Gaia}}
\newcommand{\AU}{\mathrm{au}}
\newcommand{\eqq}[1]{Equation~(\ref{#1})}
\newcommand{\ie}{\textit{i.e.\/}}
\newcommand{\eg}{\textit{e.g.\/}}
\newcommand\edited[1]{{\color{blue} {#1}}}
\newcommand\gary[1]{{\color{red} {\textbf{GMB}: #1}}}
% Turn off change highlighting
%\newcommand\edited[1]{#1}

\newcommand{\vecb}{\mathbf{b}}
\newcommand{\bhat}{\mathbf{\hat b}}
\newcommand{\vecq}{\mathbf{q}}
\newcommand{\vcirc}{v_c}
\newcommand{\Msun}{M_\odot}

\usepackage{natbib} 
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{color}
\usepackage{xcolor}
\usepackage{float}
\usepackage{hyperref}
%\usepackage{lineno}
%\linenumbers

\shorttitle{Asteroid Brownian motion}

\begin{document}

\title{Short-term Brownian motion of main-belt asteroids}

\author[0000-0002-8613-8259]{Gary M. Bernstein}
\affiliation{Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
\email{garyb@physics.upenn.edu}
\correspondingauthor{Gary M. Bernstein}

\begin{abstract}
  xxx
\end{abstract}

\keywords{asteroids}

\section{Introduction}

$ \vecb\,\bhat\,\vcirc$

\section{Conclusions}

\begin{acknowledgments}
  This work was supported by NSF grant AST-????.
  This research has made use of data and/or services provided by the International Astronomical Union's Minor Planet Center. 
\end{acknowledgments}

\newpage
\bibliographystyle{aasjournal}
%\bibliography{references}

\end{document}

### 2025-02-11 main.tex
Finally got perturbation algebra right

@@
 
+\newcommand{\vecI}{\mathbf{I}}
 \newcommand{\vecb}{\mathbf{b}}
+\newcommand{\vece}{\mathbf{e}}
 \newcommand{\bhat}{\mathbf{\hat b}}
+\newcommand{\rhat}{\mathbf{\hat r}}
+\newcommand{\phat}{\boldsymbol{\hat\phi}}
+\newcommand{\uhat}{\boldsymbol{\hat u}}
+\newcommand{\zhat}{\mathbf{\hat z}}
+\newcommand{\vecp}{\mathbf{p}}
 \newcommand{\vecq}{\mathbf{q}}
+\newcommand{\vecr}{\mathbf{r}}
+\newcommand{\vecv}{\mathbf{v}}
+\newcommand{\vecx}{\mathbf{x}}
+
+\newcommand{\matA}{A}
+\newcommand{\matB}{B}
+
 \newcommand{\vcirc}{v_c}
+\newcommand{\vrel}{v_{\rm r}}
 \newcommand{\Msun}{M_\odot}
-
+\newcommand{\covm}{C}
+\newcommand{\lop}{\varpi}
+\newcommand{\Lhat}{\hat L}
 \usepackage{natbib} 
@@
 
-$ \vecb\,\bhat\,\vcirc$
+We aim to calculate the Brownian motion variance to an accuracy of tens of percent.
+Greater accuracy in the calculation is not warranted since one of the principal inputs---the number of MBAs vs mass---is uncertain to at least this level because of unknowns in the conversion from absolute magnitude $H$ to mass.
+We will therefore be at liberty to drop terms that modify the perturbations by $O(e^2)$ of the target asteroid.
+
+We adopt a cylindrical coordinate system $(r,\phi,z),$ with $\zhat$
+normal to the initial orbital plane, and $\phi=z=0$ toward the
+perhelion of the initial orbit, \ie\ both the initial ascending node
+$\Omega_0$ and longitude of perihelion $\lop_0$ equal to zero.  The
+unit vectors $\rhat, \phat$ rotate with the target asteroid.  The mean
+anomaly at time $t=0$ is $M_0.$
+
+All distances will be given in units of the original semimajor axis $a_0$, and all velocities in units of the circular velocity $\vcirc \equiv \sqrt{G\Msun/a_0}.$  In these units, $G\Msun=1,$ the period of the initial orbit is $2\pi$ and the (unperturbed) mean anomaly is $M=t+M_0.$
+
+We describe all encounters with other asteroids in the impulse approximation, defining $\vecI$ as the $\Delta\vecv$ imparted on the target by the deflector.
+In our units, the gravitational impulse imparted by a deflector of mass $M_d$ approaching at impact parameter $\vecb=b\bhat$ and relative velocity $\vrel$ is
+\begin{equation}
+  \vecI = 2 \frac{M_d}{\Msun} (bv)^{-1} \bhat.
+  \label{eq:impulse}
+\end{equation}
+All of our results will be derived at leading order in $\vecI,$ which we will find is very well justified by the small size of the Brownian motion timescales of decades.
+
+One of our tasks will be to derive, from the known asteroid population, the rate (per target) of encounters vs the imparted impulse,
+\begin{equation}
+  \frac{dN}{dt\,dI_r\,dI_\phi\,dI_z},
+  \label{eq:dN}
+\end{equation}
+where the components of $\vecI$ are given in the cylindrical basis vectors about the asteroid's position at the impulse.
+We will approximate this function as independent of the orbit of the target MBA for asteroids with $2.2??\,\AU < a < 4.2\,\AU.$  We will further assume that the impulses on a given target are drawn independently from this distribution, \ie\ a Poisson process defined by this rate.
+The only properties of the impulse distribution are its second moments for components $\alpha \in \{r,\phi,z\}$
+\begin{equation}
+  \left \langle n I_\alpha^2 \right\rangle \equiv \int d^2I \frac{dN}{dt\,dI_r\,dI_\phi\,dI_z} I_\alpha^2.
+\label{eq:nvsq}
+\end{equation}
+The average of the cross terms $I_rI_z$ and $I_\phi, I_z$ will vanish if the deflector distribution is symmetric in inclination, and we will find numerically that the mean $I_rI_\phi$ is small enough to ignore.
+
+From this knowledge of the impulse distribution, our goal is to obtain the covariance matrix of the deviations in the target's observed position, relative to the initial orbit, after some time $t_{\rm obs}.$ The position observables are the range, plus the latitude and longitude of the target.
+We will simplify our results by assuming a heliocentric observer, so that the observational position vector is $\vecp\equiv (r,\theta=z/r,\phi).$ The quantity we seek is the covariance matrix $\covm^p$ of the observations attributable to the accumulated gravitational perturbations:
+\begin{equation}
+  \covm^p  \equiv  \left\langle  \Delta\vecp \Delta\vecp^T\right\rangle,  \label{eq:Cp}
+\end{equation}
+where the angle brackets indicate an average over possible realizations of the impulse history, and the mean anomaly $M$ at the time of observation.
+To do so, we will introduce an intermediate set of 6 parameters $\vecq$ describing the deviations of the orbital elements induced by the impulses.
+The $\vecq$ components will be selected to be equal to zero before any impulses are applied, and each responds linearly to $\vecI$ at $|I|\ll 1.$
+We will derive the matrix $\matA$ that describes the orbital-element shifts at time $t$ that arise from an impulse at time $t_i$: 
+\begin{equation}
+  \Delta \vecq(t,t_i) = \matA(e_0, t, t_i) \cdot \vecI.
+  \label{eq:A}
+\end{equation}
+In our first-order perturbation theory, $\vecq(t)$ will be the sum of the $\Delta\vecq(t_i)$ imparted by all impulses applied at times $0<t_i<t.$  Because the impulses are uncorrelated, $\vecq$ will therefore be the result of a random walk.
+The distribution will have a covariance matrix $\covm^q(t) \equiv \left\langle \vecq(t) \vecq^T(t) \right\rangle$ whose elements are
+\begin{eqnarray}
+  \covm^q_{jk}(e_o,t) & = & \left\langle \sum_{i,\gamma} \matA_{j\gamma}(e,t,t_i) \matA_{k\gamma}(e,t,t_i) I^2_{i,\gamma} \right\rangle \\
+           & = & \sum_\gamma \int dt_i \matA_{j\gamma}(e,t,t_i) \matA_{k\gamma}(e,t,t_i) \left\langle n I_\gamma^2\right\rangle.
+\label{eq:Cqjk}
+\end{eqnarray}
+In the first line, the sum $i$ runs over the impulses and $\gamma$ runs over the components $r,\phi,z$ of the impulse.
+The second line evaluates the expectation value of averaging over realizations of the random walk of impulses, exploiting the independence of the individual impulses from each other.
+
+The last element of our calculation will be a conversion from the element shifts $\vecq$ into the observed displacements $\Delta\vecp$.
+In linear perturbation theory this will again be expressible as a matrix
+\begin{equation}
+  \Delta\vecp(e_0,t) = \matB(e_0,t) \cdot \vecq(t).
+\label{eq:B}
+\end{equation}
+Combining this with \eqq{eq:Cp} and \eqq{eq:Cqjk} yields the desired result
+\begin{eqnarray}
+  \covm^x_{\alpha\beta}(t,e_0) & = & \sum_{jk} B_{\alpha j}(e,t) B_{\beta k}(e,t) \covm^q_{jk}(t) \nonumber \\
+  & = & \sum_\gamma \left\langle nI_\gamma^2\right\rangle \int dt_i \left[AB(e_0,t,t_i)\right]_{\alpha\gamma}  \left[AB(e_0,t,t_i)\right]_{\beta\gamma} 
+\label{eq:ABq}
+\end{eqnarray}
+which we would wish to average over the phase of the initial mean anomaly $M_0$ of the MBA.
+With this overall strategy, we will proceed to deriving the forms of $dN/dt\,d^3I,$ $\matA(e,t,t_i),$ and $\matB(e,t).$
+
+\subsection{Impulse distribution}
+
+\subsection{Element shifts from impulses}
+
+Five of our orbital element perturbations $\vecq$ will be taken from constants of the motion.
+The mean anomaly $M$ is not appropriate as the sixth, time-dependent element of $\vecq$ because the derivative $dM/dI$ can diverge as $e_0\rightarrow 0.$ Instead we introduce $\tau = M+\lop,$ which we will show does have a shift $\Delta\tau$ during an impulse that has finite and linear response to $\vecI$.
+Between impulses, $\tau$ advances with the mean motion as $a^{-3/2}t.$  We define a coordinate system that is uniformly rotating with the unperturbed $\tau_0=t+M_0.$ The two smoothly rotating unit vectors $\uhat_\parallel=(\cos \tau_0, \sin \tau_0)$ and
+$\uhat_\perp=(-\sin \tau_0, \cos \tau_0)$ satisfy $\uhat_\parallel \times \uhat_\perp = \zhat,$ and we will use subscripts $\parallel$ and $\perp$ to represent projections onto these components.
+In particular, the components of the ellipticity vector $e_\parallel$ and $e_\perp$ have the values $e_\parallel=e\cos M$ and $e_\perp=-e\sin M$ in the unperturbed orbit.
+The true anomaly $\nu,$ radius $r$, and azimuthal angle $\phi,$ and the velocity components of the initial orbit to $O(e)$ are
+\begin{eqnarray}
+  M & = & M_0 + t = \tau_0,  \nonumber \\
+  \nu & = & M - 2 e_\perp \nonumber \\
+  \phi & = & \nu + \lop = \tau - 2e_\perp, \nonumber\\
+  r & = & 1-e_\parallel, \nonumber \\
+  v_r & = & -e_\perp \nonumber \\
+  v_t & = & 1 + e_\parallel
+            \label{eq:kepler}
+\end{eqnarray}
+
+The first element adopted for $\vecq$ is the shift $\Delta a$, derivable from the orbital energy $E=-1/2a.$
+\begin{eqnarray}
+  \Delta E & = &  \vecv \cdot \vecI + O(I^2) \\
+  \quad \Rightarrow \quad \Delta a & = & 2\left(v_r I_r + v_\phi I_\phi\right) = -2e_\perp I_r + 2(1+e_\parallel) I_\phi.
+  \label{eq:da}
+\end{eqnarray}
+The initial angular momentum $\mathbf{L} = \sqrt{1-e^2}\,\zhat \approx \zhat$ is altered by
+\begin{eqnarray}
+  \Delta \mathbf{L} & = & \vecr \times \vecI = (1-e_\parallel) I_\phi \zhat - (1-e_\parallel) I_z \phat \\
+  \quad \Rightarrow \quad \Delta \Lhat_\parallel & = & -2e_\perp I_z
+  \label{eq:dLpar}\\
+  \Delta \Lhat_\perp & = & -(1-e_\parallel)  I_z
+                  \label{eq:dLperp}
+\end{eqnarray}
+We adopt the changes in direction $(\Delta\Lhat_x, \Delta\Lhat_y)$ as the next two elements of $\vecq$ specifying the inclination and ascending node $\Omega$ of the perturbed orbit.
+These are a rotation of $(\Delta\Lhat_\parallel, \Delta\Lhat_\perp)$ by the angle $\tau.$  The conversion from the first line above to the subsequent two makes use of the decomposition $\phat=\uhat_\perp + 2e_\perp \uhat_\parallel$ derivable from the value of $\phi$ in Equations~(\ref{eq:kepler}).
+
+The eccentricity vector $\vece = \vecv \times (\vecr \times \vecv) - \rhat$ is altered by the impulse according to
+\begin{eqnarray}
+  \Delta\vece & = & (2\rhat - rv_r \phat) I_t - \phat I_r \\
+\label{eq:depar}
+  \quad \Rightarrow \quad \Delta e_\parallel & = &  -2e_\perp I_r + 2 I_\phi \\
+\Delta e_\perp & = & -I_r - 3e_\perp I_\phi.
+\label{eq:deperp}
+\end{eqnarray}
+We adopt $\Delta e_x$ and $\Delta e_y$ as two additional components of $\vecq,$ again related through a rotation by $\tau$ to the quantities in Equations~(\ref{eq:depar}) and (\ref{eq:deperp}).
+
+The final element of $\vecq$ will be the shift $\Delta\tau$ induced in $M+\lop$ by the impulse.
+This can be obtained by enforcing the conditions that the radius $r$ or the azimuthal angle $\phi$ must remain constant during the impulse.
+To findthe latter, we need a formula for $\nu(M)$ as a power series in $e.$  Standard formulae in the literature lead to:
+\begin{eqnarray}
+  \nu & = & M + 2e \sin M + \frac{5e^2}{4} \sin 2M + O(e^3) \\
+  \label{eq:nue2}
+    & = & M + 2(e\sin M) + \frac{5}{2} (e \sin M) (e \cos M) + O(e^3) 
+\end{eqnarray}
+We wish to find the first-order shift in $\phi=\nu + \lop$ and set it to zero.
+But the derivatives of $M$ and $\lop$ with impulse can become infinite.
+Instead we introduce a perturbation $\Delta\tau = \Delta(M+\lop).$ We can now express several post-impact quantities as deviations from the unperturbed quantities (subscripted with 0):
+\begin{eqnarray}
+  M & = & \tau-\lop = \tau_0 -\lop + \Delta\tau \\
+  e \sin M & = & e \sin (\tau_0 -\lop) + \Delta\tau \left[ e \cos  (\tau_0 -\lop)\right] \\
+    & = & -e_{0\perp}-\Delta e_\perp + e_{0\parallel} \Delta\tau +O(e^2) \\
+  e \cos M & = & e \cos (\tau_0 -\lop) - \Delta\tau \left[ e \sin  (\tau_0 -\lop)\right] \\
+    & = & e_{0\parallel}+\Delta e_\parallel + e_{0\perp} \Delta\tau +O(e^2) \\
+\Rightarrow \quad \phi =\nu+\lop & = & \tau_0 + \Delta \tau + 2( -e_{0\perp}-\Delta e_\perp + e_{0\parallel} \Delta\tau )
+                              + \frac{5}{2}\left( -e_{0\perp}\Delta e_\parallel-e_{0\parallel}\Delta e_\perp\right) + O(e^2).
+\end{eqnarray}
+Forcing $\phi$ to be unchanged during the impulse requires
+\begin{eqnarray}
+  \Delta\tau & = & \frac{5 e_\perp}{2} \Delta e_\parallel + \left(2-\frac{3e_\parallel}{2}\right) \Delta e_\perp \\
+             & = & \left(-2+\frac{3e_\parallel}{2}\right) I_r - e_\perp I_\phi.
+                   \label{eq:dtau0}
+\end{eqnarray}
+We have dropped the 0 subscripts on $e_\parallel,e_\perp$ at this point for brevity---they will refer to the values for the unperturbed orbit at the time of impulse, unless noted otherwise.
+
+After the impulse, the mean anomaly $M$ advances at a rate of $a^{-3/2}t,$ meaning that an additional term $\Delta\tau = -3\Delta a (t-t_i)/2$ accrues by time $t$.
+The total perturbation of $\tau$ from the nominal value $t+M_0$ becomes
+\begin{equation}
+  \Delta\tau= \left[-2+\frac{3e_\parallel}{2} +3 e_\perp (t-t_i)\right] I_r  - \left[3(1+e_\parallel)(t-t_i) + e_\perp\right] I_\phi.
+  \label{eq:dtau}
+\end{equation}
+
+This equation gives the last row of the $\matA$ matrix giving element-level perturbations from individual impulses.  The preceding equations can be restated to give the other rows:
+\begin{eqnarray}
+  \Delta a & = & -2e_\perp I_r + 2(1+e_\perp) I_\phi, \\
+  \Delta \Lhat_\parallel & = & -2e_\perp I_z \\
+  \Delta \Lhat_\perp & = & -(1-e_\parallel) I_z \\
+  \Delta e_\parallel & = &  -2e_\perp I_r + 2 I_\phi \\
+  \Delta e_\perp & = & -I_r - 3e_\perp I_\phi.
+\end{eqnarray}
+
+\subsection{Astrometric and ranging shifts from elements shifts}
+\begin{equation}
+  \label{eq:Ee2}
+  \cos E  =  e\cos M - (e \sin M)^2 + O(e^3).
+\end{equation}
+
+Keeping only terms from Equations~(\ref{eq:kepler}) up to first order in $\vecq,$ the true anomaly $\nu,$ mean anomaly $M=t + \Delta t - \lop,$ and the azimuthal angle $\phi$ are related by
+\begin{eqnarray}
+  \nu  =  M + 2e\sin M  & = & t +  2e_0 \sin t  + \Delta t - \lop + 2 e_0(\Delta t - \lop) \cos t + 2\sin t \Delta e_x \\
+  \Rightarrow \quad \phi  =  \nu + \lop & = &  \left[t + 2e_0\sin t\right] + \left[(1+2e_0\cos t) \Delta t - 2\cos t \Delta e_y + 2\sin t \Delta e_x\right]
+                                                    \label{eq:phi}
+\end{eqnarray}
+where we have used $e_0\lop = \Delta e_y.$ The radial coordinate and polar angle $\theta$ are
+\begin{eqnarray}
+  r & = & (1 + \Delta a) \left[1- (e_0+\Delta e_x) \cos M \right] \nonumber \\
+\label{eq:r}
+  & = & \left[ 1-e_0\cos t \right] + \left[ (1-e_0\cos t) \Delta a + e_0\sin t \Delta t - \sin t \Delta e_y - \cos t \Delta e_x \right] \\
+  \theta = z/r & = & -\Lhat_x \cos\phi - \Lhat_y \sin\phi \nonumber \\
+\label{eq:theta}
+               & = & \left[ 0 \right] + \left[ -\cos t \Lhat_x  - \sin t \Lhat_y\right].
+\end{eqnarray}
+The second line of each deviation is divided into two bracketed terms, the first being the initial orbit and the second being the first-order terms in elements of $\vecq,$ which are therefore the elements of $\matB$.
+
+A heliocentric observer will measure ``range noise'' of $\covm^p_{rr}=\langle (\Delta r)^2\rangle$ in the measured distance to the target asteroid.
+We find this by squaring the second bracketed term for $r$ and averaging $t$ over an interval of size $2\pi$ to compensate for the fact that we have assumed $M=0$ at $t=0,$ whereas in fact the $M_0$ values are random.
+We can then derive
+\begin{equation}
+  \left\langle(\Delta r)^2\right\rangle = (\Delta a)^2 + \frac{1}{2}\left[ (\Delta e_x)^2 + (\Delta e_y)^2\right] + \frac{e_0}{2}(\Delta a \Delta e_x -  \Delta t \Delta e_y).
+  \label{eq:varr}
+\end{equation}
+The same heliocentric observer will see angular shifts of the target position satisfying
+\begin{eqnarray}
+  \label{eq:phiphi}
+  \left\langle (\Delta\phi)^2\right\rangle & = & (\Delta t)^2 + 2(\Delta e_x^2 + \Delta e_y^2)  -4e_0
+  \left(\Delta t \Delta e_y\right)\\
+  \label{eq:thetatheta}
+  \left\langle (\Delta\theta)^2\right\rangle & = & \frac{1}{2} \left( \Lhat_x^2 + \Lhat_y^2\right) \\
+  \label{eq:thetaphi}
+  \left\langle \Delta\theta\Delta\phi\right\rangle & = & -e_0\Delta t \Lhat_x + (\Delta e_y \Lhat_x -  \Delta e_x \Lhat_y)
+\end{eqnarray}
 

### 2025-02-16 main.tex
Writing impulse sections

@@
 
-\title{Short-term Brownian motion of main-belt asteroids}
+\title{Brownian motion of main-belt asteroids on human timescales}
 
@@
 
-We aim to calculate the Brownian motion variance to an accuracy of tens of percent.
-Greater accuracy in the calculation is not warranted since one of the principal inputs---the number of MBAs vs mass---is uncertain to at least this level because of unknowns in the conversion from absolute magnitude $H$ to mass.
-We will therefore be at liberty to drop terms that modify the perturbations by $O(e^2)$ of the target asteroid.
+High-precision measurements of positions of Solar System bodies have
+great value to multiple scientific pursuits.  They can reveal
+gravitational accelerations due to undiscovered bodies such as a
+hypothetical ``Planet X'' or passing primordial black holes, determine
+the masses of known bodies, tests the laws of gravity themselves, and characterize 
+the non-gravitational forces that drive many forms of planetary
+migration, \eg\ the maintainence of the near-Earth asteroid (NEA)
+population.  While most applications of precision orbit determination
+have used the 8 major planets as test particles, there are now $>10^6$
+known minor planets available for this task, whose astrometric
+information content is growing extremely rapidly through large-scale
+survey projects such as Gaia, PanSTARRS, and particularly the upcoming
+\textit{Legacy Survey of Space and Time} (LSST), which can produce
+milliarcsecond-scale accuracy for magnitude-limited small-body populations.  The great majority
+of currently-tracked objects are main-belt asteroids (MBAs), so we
+wish to investigate the potential power of MBAs as gravitational test
+bodies.
+
+The most advanced ephemerides for the Solar System consider the
+gravitational forces emanating from the Sun, from the major planets and their large
+moons, from up to $\approx300$ of the largest individual MBAs, and
+from distributed mass annuli representing the smaller members of
+various small-body populations (asteroid belts, Kuiper belt, etc.).
+Here we ask: \textit{what is the typical RMS deviation in angular or radial
+position that a typical MBA accrues because of its gravitational
+encounters with other individual MBAs?}  We are interested not in the
+perturbations that can be well described by a multipolar ring model
+for the full population; rather the Brownian motion that accumulates
+from closer encounters either with MBAs that \emph{are not included} in the
+ephemeris model at all.  For those larger asteroids that \emph{are} included
+in an ephemeris model, we need to consider the uncertainties in
+orbital motion that accrue from the uncertainty $\sigma_M$ in their
+masses.  Any such departures became a source of noise in any
+inferences that we make from use of these MBAs as
+dynamical tracer particles.
+
+While long-term diffusion of orbital elements has been extensively
+investigated in the context of planetary-system formation and
+planetary migration, we address here the more limited question of how
+much the MBAs will have their orbit elements and positions altered
+by Brownian motion over a time period $T \lesssim100$~yr of human
+observation, in the present dynamical environment.  We will make use
+of the currently known population of MBAs, correcting for
+incompleteness of current surveys at lower masses---or more precisely,
+at fainter absolute magnitudes $H,$ since only a handful of small
+asteroids have usefully accurate mass measures.
+
+\section{Calculation overview}
+We aim to calculate the Brownian motion variance to an accuracy of
+tens of percent.  Greater accuracy in the calculation is not warranted
+since one of the principal inputs---the number of MBAs vs mass---is
+uncertain to at least this level because of unknowns in the conversion
+from $H$ to mass.  We will therefore be at liberty to drop terms that
+modify the perturbations by $O(e^2)$ of the tracer asteroid that is
+being tracked.    We will assume throughout that the joint
+distribution of mass (or $H$) and orbital elements is separable into a
+mass distribution and an orbital distribution.  In this scenario, the
+Brownian motion is independent of the tracer mass, and involves an
+integral over the mass distribution of the deflecting bodies.
 
@@
 \end{equation}
-All of our results will be derived at leading order in $\vecI,$ which we will find is very well justified by the small size of the Brownian motion timescales of decades.
+All of our results will be derived at leading order in $\vecI,$ which
+is very well justified by the small size of MBA-induced impulses.
 
@@
 \end{equation}
-where the components of $\vecI$ are given in the cylindrical basis vectors about the asteroid's position at the impulse.
-We will approximate this function as independent of the orbit of the target MBA for asteroids with $2.2??\,\AU < a < 4.2\,\AU.$  We will further assume that the impulses on a given target are drawn independently from this distribution, \ie\ a Poisson process defined by this rate.
-The only properties of the impulse distribution are its second moments for components $\alpha \in \{r,\phi,z\}$
+where the components of $\vecI$ are given in the cylindrical basis
+vectors about the asteroid's position at the impulse.
+We will estimate this function from the known population,
+approximating it as constant within each of three subsets of tracer MBAs:
+the inner, middle, and outer populations bounded by the 4:1, 3:1,
+5:2, and 2:1 mean-motion resonances with Jupiter.
+We will further assume that the impulses on a given target are drawn
+independently from this distribution, \ie\ a Poisson process defined
+by this rate.  In this case, the only properties of the impulse
+distribution that we need are its second moments for components $\alpha \in \{r,\phi,z\}$
 \begin{equation}
@@
 \end{equation}
-The average of the cross terms $I_rI_z$ and $I_\phi, I_z$ will vanish if the deflector distribution is symmetric in inclination, and we will find numerically that the mean $I_rI_\phi$ is small enough to ignore.
+The average of the cross terms $I_rI_z$ and $I_\phi, I_z$ will vanish if the deflector distribution is symmetric in inclination, and we find numerically that the mean $I_rI_\phi$ is small enough to ignore.
 
-From this knowledge of the impulse distribution, our goal is to obtain the covariance matrix of the deviations in the target's observed position, relative to the initial orbit, after some time $t_{\rm obs}.$ The position observables are the range, plus the latitude and longitude of the target.
+From this knowledge of the impulse distribution, our goal is to obtain the covariance matrix of the deviations in the target's observed position, relative to the initial orbit, after some time $T.$ The position observables are the range, plus the latitude and longitude of the target.
 We will simplify our results by assuming a heliocentric observer, so that the observational position vector is $\vecp\equiv (r,\theta=z/r,\phi).$ The quantity we seek is the covariance matrix $\covm^p$ of the observations attributable to the accumulated gravitational perturbations:
@@
 which we would wish to average over the phase of the initial mean anomaly $M_0$ of the MBA.
-With this overall strategy, we will proceed to deriving the forms of $dN/dt\,d^3I,$ $\matA(e,t,t_i),$ and $\matB(e,t).$
 
-\subsection{Impulse distribution}
+Section~\ref{sec:impulse} describes the estimation of $\left\langle
+  nI_\alpha^2\right\rangle$ for MBA regions.
+Section~\ref{sec:propagation} derives the forms of $\matA(e,t,t_i),$
+and $\matB(e,t).$  Section~\ref{sec:results} combines these and
+summarizes the results and their implications.
+
+\section{Impulse distribution}
+\label{sec:impulse}
+\subsection{Derivation}
+Under the assumption that the distributions of masses and orbital
+elements of MBAs are separable, the quantities needed to describe the
+random walk of the asteroid's orbits can be expressed as an integral
+over the absolute magnitude $H$ of the deflectors, the velocity $v$,
+impact parameter $b,$ and polar/azimuthal angles $\theta,\phi$ of the
+unit vector of the impulse direction $\bhat$:
+\begin{eqnarray}
+  \left\langle I_\gamma^2 \right\rangle & = & \int
+                                              dH\,dv\,db\,d\theta\,d\phi
+                                              \frac{dn}{dH\,dv\,db\,d\theta\,d\phi}
+                                              I^2_\gamma(H,v,b,\theta\phi)
+  \\
+  & = & \int dH \frac{dN}{dH} \int dv
+        \frac{dn}{dA\,dv\,d\theta\,d\phi} \hat b_\gamma^2(\theta,\phi)
+        \int db_{b_{\rm min}}^{b_{\rm max}} 2\pi b\,db\,
+        \left(\frac{\sigma_M(H)}{bv}\right)^2 \\
+  & = & \int dH \frac{dN}{dH} \sigma_M(H)^2  \int dv
+        \frac{dn}{dA\,dv} v^{-2} \left\langle\hat b_\gamma^2(v)\right\rangle
+        2\pi \log(b_{\rm max}/b_{\rm min}).
+\end{eqnarray}
+The second row adopts the (numerically verified) assumption that the
+distribution of impact parameter $b$ will always be $\propto dA=2\pi
+b\,db$ in the range of $b$ of interest to us.  The mean rate of encounters
+between a deflector-tracer pair, as a function of the impact velocity
+and geometry is $dn/dA\,dv\,d\theta\,d\phi,$ which we will determine
+through numerical measurements with the orbits of all known MBAs. The
+$H$ distribution of MBAs is $dN/dH.$ We denote as $\sigma_M(H)$
+the RMS error in the mass of MBAs at $H$ used in the ephemeris model.
+We use $\sigma_M$ rather than $M$ since we are seeking the size of
+deviations from the ephemeris model.  For MBAs that are not included
+in the ephemeris model, $\sigma_M=M(H),$ the full mass of the deflector.
+
+In the third line, we execute the integral over $b$ and introduce the
+average geometry factors $\langle \hat b^2_\gamma\rangle$ of encounters as
+a function of $v$.  It must be true that
+\begin{equation}
+  \left\langle \hat b_r^2 \right\rangle
+  + \left\langle \hat b_\phi^2 \right\rangle
+  + \left\langle \hat b_z^2 \right\rangle = 1,
+\end{equation}
+but equality of the three is not required.
+
+For the values of $b_{\rm min}$ and $b_{\rm max},$ we adopt heuristics
+for the nature of impulses of interest.  We set $b_{\rm max}$ by the
+requiring that $b/v<1,$ such that the ``impulse'' is being applied
+over a time shorter than $1/2\pi$ of the orbital period.  Events
+lasting longer than this are either extended close encounters of 2
+MBAs with very similar orbits---which are rare and can be identified
+and modeled in advance; or they are slow, longer-range interactions
+that would be adequately modeled by a multipole model of the
+collective mass of the asteroid belts.  Thus we take $b_{\rm max}=v.$
+
+For $b_{\rm min},$ we adopt the criterion that encounters at
+sufficiently small $b$ will generate sufficiently large impulses $I$
+that observations of the tracer will detect this individual impulse at
+levels well above measurement noise.  This would mean that an
+ephemeris model could include the mass of the deflector asteroid, and
+this mass could be determined from fitting the data of the tracer.
+Fitting an ephemeris that includes all of these Individually
+detectable encounters is entirely feasible, as long as the relevant
+MBAs are known, with modestly accurate orbits.
+Such encounters would thus no longer be considered source of
+stochastic Brownian motion.  If we crudely choose some threshold
+$I_{\rm det}$ of impulse as being large enough to generate high-$S/N$
+deflections on its tracer, then our condition becomes $M(H)/bv <
+I_{\rm det},$ or $b_{\rm min}=M(H)/vI_{\rm det}.$  We then have
+\begin{equation}
+   \left\langle I_\gamma^2 \right\rangle = 
+2\pi \int dH \frac{dN}{dH} \sigma^2_M(H)  \int_{\sqrt{M(H)/I_{\rm
+      det}}}^\infty \frac{dv}{v^2} 
+        \frac{dn}{dA\,dv}  \left\langle\hat b_\gamma^2(v)\right\rangle
+        \log \left[v^2 I_{\rm det}/M(H)\right].
+        \label{eq:nI2}
+      \end{equation}
+The lower bound on the $v$ integral marks the speed below which
+$b_{\rm max}<b_{\rm min}.$
+
+\subsection{Numerical results}
+\eqq{eq:nI2} reduces the calculation to a double integral over the
+deflector $H$ distribution and the relative velocity $v.$  The
+constituents of this calculation are estimated as follows.
+
+\subsubsection{Encounter rates and geometries}
+The pairwise interaction rate $dn/dA\,dv$ and the impulse direction
+distributions $\langle \hat b^2_\gamma \rangle$ are estimated from a
+numerical integration of all of the MBA orbits available
+from the Minor Planet Center (MPC) as of ???? 2025.   We retain as
+potential deflectors the 1.26~million bodies with $1.8<a<4.2$~AU,
+observations at multiple oppositions, and uncertainty values $U\le5.$
+
+From this full MBA list, we select a subset of 5000 objects as
+``tracers'' randomly selected from the MBAs in each of three regimes:
+``Inner'' MBAs with semi-major axes between the 4:1 and 3:1
+mean-motion resonances of Jupiter; ``Middle'' MBAs between the 3:1 and
+5:2 resonances; and ``Outer'' MBAs between the 5:2 and 2:1
+resonances.  This is enough to be representative of the statistics of
+each MBA class, but few enough that the calculation can be done easily
+on a laptop computer.
+
+Using a simple leapfrog integrator with gravity from the Sun and the 8
+major planets, we advance all MBA's orbits from their heliocentric
+osculating elements at the epochs given by the MPC, to barycentric
+state vectors on 1 May 2025.  We then find all tracer-deflector pairs
+that approach within 0.03~AU of each during the $\pm1$-day period
+around this initial epoch, assuming inertial relative motion during
+this interval.  The circumstances of such encounters are saved: the
+identities of the two MBAs involved, the time of closest approach, the
+relative velocity $v,$ and the impact parameter vector $\vecb.$
+
+The leapfrog integrator advances all 1.26 million state vectors by 2
+days, and the process is repeated.  We continue at 2-day intervals
+until we have recorded 10 years' worth of encounters---17.7 million
+total impulses.  These events are split according to their combination
+of tracer class (Inner, Middle, or Outer belt) and deflector class
+(Inner, Middle, Outer, Other).
+
+For each type of impulse (\eg Inner-Inner), we combine the list of
+events with the counts of candidate tracers and deflectors to
+calculate $dn/dA\,dv,$ the event rates per eligible pair; and the mean
+geometry of the encounter, $\langle \hat b^2_\gamma \rangle,$ as a
+function of $v$.  Figure~\ref{fig:rtz} plots these functions of $v$
+for the Inner-Inner, Middle-Middle, and Outer-Outer scattering
+events.  Here it is apparent that the $\hat b_\phi$ component is
+larger than $\hat b_z$ and $\hat b_r,$ but there is significant
+variation with MBA class and with $v$.
+
+\subsubsection{Asteroid properties}
+We crudely approximate the population as having a geometric albedo of 0.25,
+density of 2500~kg~m$^{-3},$ and spherical shapes, which leads to the
+conversion
+\begin{equation}
+  M(H) = 1.2\times10^{-17} \times 10^{-0.6(H-15)}\, M_\odot.
+\end{equation}
+
+The two largest asteroids, Ceres and Vesta, have masses determined to
+high precision by the \textit{Dawn} spacecraft.  Thus despite holding
+most of the mass of the MBAs, their $\sigma_M$ values are just ??? and
+the unknown part of their contributions to other MBA's motion are
+quite small.  Aside from a handful of other asteroids in
+well-characterized binary systems, most other knowledge of MBA masses
+comes from mutual gravitational encounters.
+\citet{negin} forecast the extent to which LSST data will be able to
+constrain the masses of individual MBAs by measuring deflections of
+the $\approx 10$ tracer asteroids they encounter most closely over the
+course of the nominal 10-year survey.  They find that the $1\sigma$
+statistical error on a typical MBA is $\sigma_M=???\, M_\odot.$
+
+$\sigma_M(H)$ rolled off at ...
+
+Asteroid counts $dN/dH$
+
+\subsubsection{Individually detectable impulses}
+
+\subsubsection{Results}
+
+\section{Propagation to position shifts}
+\label{sec:propagation}
 
 \subsection{Element shifts from impulses}
+\label{sec:elements}
 
@@
 \subsection{Astrometric and ranging shifts from elements shifts}
+\label{sec:observe}
 \begin{equation}
@@
 
-\section{Conclusions}
+\section{Results and conclusions}
+\label{sec:results}
 

### 2025-02-16 main.tex
dN/dH done

@@
 observations at multiple oppositions, and uncertainty values $U\le5.$
+We designate each source as being an ``Inner'' MBA, with semi-major axes between the 4:1 and 3:1
+mean-motion resonances of Jupiter; ``Middle'' MBA between the 3:1 and
+5:2 resonances; ``Outer'' MBAs between the 5:2 and 2:1
+resonances; and the small remainder as ``Other.''  Our working
+assumption will be that the MPC objects are an unbiased sample of the
+orbital-element distribution within each of the Inner, Middle, and
+Outer belt regions.
 
-From this full MBA list, we select a subset of 5000 objects as
-``tracers'' randomly selected from the MBAs in each of three regimes:
-``Inner'' MBAs with semi-major axes between the 4:1 and 3:1
-mean-motion resonances of Jupiter; ``Middle'' MBAs between the 3:1 and
-5:2 resonances; and ``Outer'' MBAs between the 5:2 and 2:1
-resonances.  This is enough to be representative of the statistics of
-each MBA class, but few enough that the calculation can be done easily
+From this full MBA list, we select a random subset of 5000 objects
+from each of the Inner, Middle, and Outer belts to serve as a sample
+of ``tracers.''  We record all of the passages of any tracer MBA
+within 0.03~AU of any other ``deflector'' MBA (drawn from the full
+catalog of 1.26~million) over a 10-year period.  The 5000 tracers per
+belt region are enough to be representative of the statistics of
+the region, but few enough that the calculation can be done easily
 on a laptop computer.
+We will report $\langle nI_\gamma^2\rangle$
+separately for tracers of each region, allowing for the fact that a
+tracer of a given region can receive impulses from deflectors resident
+in all regions.
 
@@
 until we have recorded 10 years' worth of encounters---17.7 million
-total impulses.  These events are split according to their combination
-of tracer class (Inner, Middle, or Outer belt) and deflector class
-(Inner, Middle, Outer, Other).
+total impulses.
 
-For each type of impulse (\eg Inner-Inner), we combine the list of
+For each pair of tracer-deflector regions (\eg Inner-Inner), we combine the list of
 events with the counts of candidate tracers and deflectors to
@@
 
-Asteroid counts $dN/dH$
+\begin{figure}
+  \includegraphics{mbacounts.pdf}
+  \caption{Differential counts of MBAs cataloged by the MPC are
+    plotted vs absolute magnitude $H$  
+    for the Inner, Middle, Outer, and ``Other'' regions of
+    the main belt, and for their total.  We assume that they are
+    complete for $H\le17,$ and extrapolate to fainter sources using
+    the functional form from \eqq{eq:lsstmba} shown as the dotted
+    curve.}
+  \label{fig:counts}
+\end{figure}
+
+Figure~\ref{fig:counts} plots our assumptions for $dN/dH.$  The MPC
+catalog is likely close to complete in all three regions for $H\le17,$
+so we will use the MPC counts directly in this regime.  Note that the
+three regions have similar shapes of $dN/dH.$  For $H>17,$ we adopt
+the functional form for the cumulative MBA counts given by
+initial LSST predictions \citep{LSST_science_book}:
+\begin{equation}
+  N(<H) \propto \frac{10^{0.43(H-15.7)}}{10^{0.18(H-15.7)} +
+    10^{-0.18(H-15.7)}},
+\label{eq:lsstmba}
+\end{equation}
+normalizing this curve to the observed $H<17$ counts for each region.
 

### 2025-02-17 main.tex
More on impulse distribution

@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\textwidth]{isq.pdf}
+  \caption{Each panel shows statistics of the encounters between
+    tracer MBAs from one region and deflector MBAs from another.  As a
+    function of the relative velocity at impact, the three curves show
+    the mean squared value of the radial, azimuthal, and vertical
+    components of the impulse direction $\bhat.$  The horizontal
+    dashed line at value $1/3$ would be the result for isotropic
+    encounters, but we see that the azimuthal component of the impulse
+    tends to be larger, with a variety of behaviors.  The gray regions
+    trace the distribution $dN/dv$ of the rate of interactions per
+    tracer per orbit per interval in $v.$  These curves have a common
+    (arbitrary) normalization, so the relative heights properly
+    reflect the frequency of encounters between different MBA
+    regions.}
+  \label{fig:rtz}
+\end{figure}
+
 For each pair of tracer-deflector regions (\eg Inner-Inner), we combine the list of
@@
 high precision by the \textit{Dawn} spacecraft.  Thus despite holding
-most of the mass of the MBAs, their $\sigma_M$ values are just ??? and
+most of the mass of the MBAs, their $\sigma_M$ values are just
+$\approx3\times10^{-15} M_\odot$ and
 the unknown part of their contributions to other MBA's motion are
 quite small.  Aside from a handful of other asteroids in
-well-characterized binary systems, most other knowledge of MBA masses
+well-characterized binary systems or visited by spacecraft, most other knowledge of MBA masses
 comes from mutual gravitational encounters.
@@
 \begin{figure}
-  \includegraphics{mbacounts.pdf}
+  \centering
+  \includegraphics[width=0.8\textwidth]{mbacounts.pdf}
   \caption{Differential counts of MBAs cataloged by the MPC are
@@
 \subsubsection{Individually detectable impulses}
+For several \citet{negin} forecast the uncertainty $\sigma_v$ on the impulse 
+applied to each $\approx7000$ tracers using astrometric data expected from LSST
+over its 10-year duration.  The value of $\sigma_I$ depends upon the
+apparent magnitude of the tracer, the timing and geometry of the
+impulse, and the stochastic variations in the observing pattern and
+weather for LSST.
+For impulses occurring near the midpoint
+of the survey, $\approx80\%$ of all cases yield $\sigma_I<10^{-4}\,{\rm m}\,{\rm
+  s}^{-1},$ some an order of magnitude lower.  We conservatively
+assume that $I_{\rm det}=5\times 10^{-4}\,{\rm m}\,{\rm
+  s}^{-1},$ which when normalized by $v_c$ yields $I_{\rm
+  det}=10^{-8}$ at 10\% accuracy across the main belt.
 

### 2025-02-18 main.tex
sigmaM edits.

@@
 \begin{eqnarray}
-  \left\langle I_\gamma^2 \right\rangle & = & \int
+  \left\langle n I_\gamma^2 \right\rangle & = & \int
                                               dH\,dv\,db\,d\theta\,d\phi
@@
 \begin{equation}
-   \left\langle I_\gamma^2 \right\rangle = 
+   \left\langle n I_\gamma^2 \right\rangle = 
 2\pi \int dH \frac{dN}{dH} \sigma^2_M(H)  \int_{\sqrt{M(H)/I_{\rm
@@
 numerical integration of all of the MBA orbits available
-from the Minor Planet Center (MPC) as of ???? 2025.   We retain as
+from the Minor Planet Center (MPC) as of 6 January 2025.   We retain as
 potential deflectors the 1.26~million bodies with $1.8<a<4.2$~AU,
@@
 5:2 resonances; ``Outer'' MBAs between the 5:2 and 2:1
-resonances; and the small remainder as ``Other.''  Our working
+resonances; the remaining ``Other'' objects are a small minority that
+contribute very little to the total diffusion rate, and we will ignore.  Our working
 assumption will be that the MPC objects are an unbiased sample of the
@@
 on a laptop computer.
-We will report $\langle nI_\gamma^2\rangle$
-separately for tracers of each region, allowing for the fact that a
-tracer of a given region can receive impulses from deflectors resident
-in all regions.
+We will estimate the constituents of  $\langle nI_\gamma^2\rangle$ in
+\eqq{eq:nI2} separately for each of the nine combinations of tracers
+in Inner, Middle, and Outer regions with deflectors from each region.
 
@@
 osculating elements at the epochs given by the MPC, to barycentric
-state vectors on 1 May 2025.  We then find all tracer-deflector pairs
+state vectors on 1 May 2025.  Using $kd$-trees to accelerate
+pair-finding, we locate all tracer-deflector pairs
 that approach within 0.03~AU of each during the $\pm1$-day period
@@
 The leapfrog integrator advances all 1.26 million state vectors by 2
-days, and the process is repeated.  We continue at 2-day intervals
-until we have recorded 10 years' worth of encounters---17.7 million
+days, and the process is repeated.  We continue searching for pairs at
+2-day intervals until we have recorded 10 years' worth of encounters---17.7 million
 total impulses.
@@
 
-For each pair of tracer-deflector regions (\eg Inner-Inner), we combine the list of
+For each pair of tracer-deflector regions (\eg\ Inner-Inner), we combine the list of
 events with the counts of candidate tracers and deflectors to
@@
 larger than $\hat b_z$ and $\hat b_r,$ but there is significant
-variation with MBA class and with $v$.
+variation with MBA class and with $v$.  An important result is that
+the quantity $v^{-2} (dN/dA\,dv)$ is always well-behaved as $v\rightarrow 0,$
+so we do not need any special treatment to sum the integrals in
+\eqq{eq:nI2}.  The $v$ integrand peak is in the range $0.03<v/v_c<0.1.$
 
@@
 \end{equation}
-
-The two largest asteroids, Ceres and Vesta, have masses determined to
-high precision by the \textit{Dawn} spacecraft.  Thus despite holding
-most of the mass of the MBAs, their $\sigma_M$ values are just
-$\approx3\times10^{-15} M_\odot$ and
-the unknown part of their contributions to other MBA's motion are
-quite small.  Aside from a handful of other asteroids in
-well-characterized binary systems or visited by spacecraft, most other knowledge of MBA masses
-comes from mutual gravitational encounters.
-\citet{negin} forecast the extent to which LSST data will be able to
-constrain the masses of individual MBAs by measuring deflections of
-the $\approx 10$ tracer asteroids they encounter most closely over the
-course of the nominal 10-year survey.  They find that the $1\sigma$
-statistical error on a typical MBA is $\sigma_M=???\, M_\odot.$
-
-$\sigma_M(H)$ rolled off at ...
 
@@
 \end{figure}
-
+We also require the $H$ distribution of potential deflectors.
 Figure~\ref{fig:counts} plots our assumptions for $dN/dH.$  The MPC
@@
 
+Our integral for the uncertainty due to unmodelled MBA gravitation
+requires the RMS error $\sigma_M(H)$ between the true mass $M$ and the
+mass used (if any) in the ephemeris model.
+The two largest asteroids, Ceres and Vesta, have masses determined to
+high precision by the \textit{Dawn} spacecraft.  Thus despite holding
+half of the mass of the asteroid belt, their $\sigma_M$ values are just
+$\approx3\times10^{-15} M_\odot$ and
+the unknown part of their contributions to other MBA's motion are
+quite small.  Aside from a handful of other asteroids in
+well-characterized binary systems or visited by spacecraft, most other knowledge of MBA masses
+has and will come from mutual gravitational encounters.   An
+individual asteroid's mass can be estimated from its measured effect  on the
+few MBAs on which it imparts the largest impulses.  The accuracy
+$\sigma_M$ of this determination will depend upon the geometry of
+these encounters, the measurement errors on the tracers, and the size
+of the noise from unmodelled mass that we aim to determine.  These
+factors are independent of the mass of the deflector.  We will
+assign a value $\sigma_M$ to the RMS uncertainty in MBA mass
+attainable from following a few individual tracers, and calculate
+the Brownian-motion noise as a function of $\sigma_M.$  We take
+\begin{equation}
+  \sigma_M(H) = \textrm{min}\left[ \sigma_M, M(H) \right],
+\end{equation}
+because a given MBA could have a mass determination accurate to
+$\sigma_M$ if included in the ephemeris model, or if $M(H)<\sigma_M,$
+it would be omitted from the ephemeris model.
+
+To give some context, the most recent
+comprehensive investigation of MBA masses from mutual scattering
+events was by \citet{baer}, who report typical $\sigma_M\approx
+5\times10^{-13}$ for 30 MBAs.  We can expect substantial improvement
+in the LSST era when a millions of potential tracer asteroids attain
+mas-level astrometry.
+
 \subsubsection{Individually detectable impulses}
-For several \citet{negin} forecast the uncertainty $\sigma_v$ on the impulse 
-applied to each $\approx7000$ tracers using astrometric data expected from LSST
+\citet{negin} forecast the uncertainty $\sigma_v$ on an impulse 
+applied to each of $\approx7000$ tracers using astrometric data expected from LSST
 over its 10-year duration.  The value of $\sigma_I$ depends upon the
@@
 
+\citet{negin} forecast the extent to which LSST data will be able to
+constrain the masses of individual MBAs by measuring deflections of
+the $\approx 10$ tracer asteroids they encounter most closely over the
+course of the nominal 10-year survey.  They find that the $1\sigma$
+statistical error on a typical MBA is $\sigma_M=???\, M_\odot.$
+
+$\sigma_M(H)$ rolled off at ...
+
 \begin{acknowledgments}

### 2025-02-20 main.tex
draft finished

@@
 \newcommand{\matB}{B}
+\newcommand{\matC}{C}
 
+\newcommand{\nIr}{\left\langle nI_r^2\right\rangle}
+\newcommand{\nIphi}{\left\langle nI_\phi^2\right\rangle}
+\newcommand{\nIz}{\left\langle nI_z^2\right\rangle}
+
+\newcommand{\Var}{\textrm{Var}}
 \newcommand{\vcirc}{v_c}
@@
 gravitational accelerations due to undiscovered bodies such as a
-hypothetical ``Planet X'' or passing primordial black holes, determine
-the masses of known bodies, tests the laws of gravity themselves, and characterize 
+hypothetical ``Planet X'' \citep[\eg][]{holmanP9,inpopP9,trojans,occultations} or passing primordial
+black holes \citep{pbh}, determine
+the masses of known bodies \citep[\eg][]{goffin,baer}, test the laws of physics, and characterize 
 the non-gravitational forces that drive many forms of planetary
 migration, \eg\ the maintainence of the near-Earth asteroid (NEA)
-population.  While most applications of precision orbit determination
+population \citep{yorp}.  While most applications of precision orbit determination
 have used the 8 major planets as test particles, there are now $>10^6$
-known minor planets available for this task, whose astrometric
+known minor planets available for this task.  These small bodies' astrometric
 information content is growing extremely rapidly through large-scale
-survey projects such as Gaia, PanSTARRS, and particularly the upcoming
-\textit{Legacy Survey of Space and Time} (LSST), which can produce
-milliarcsecond-scale accuracy for magnitude-limited small-body populations.  The great majority
+survey projects that can produce
+milliarcsecond-scale accuracy for magnitude-limited small-body populations, such as Gaia \citep{gaiass3},
+PanSTARRS,\footnote{\url{https://www2.ifa.hawaii.edu/research/Pan-STARRS.shtml}}
+and particularly the upcoming Rubin observatory
+\textit{Legacy Survey of Space and Time} (LSST).\footnote{\url{https://rubinobservatory.org/}}  The great majority
 of currently-tracked objects are main-belt asteroids (MBAs), so we
@@
 
-The most advanced ephemerides for the Solar System consider the
+The most advanced ephemerides for the Solar System \citep{de440,inpop,pitjeva} consider the
 gravitational forces emanating from the Sun, from the major planets and their large
@@
 from distributed mass annuli representing the smaller members of
-various small-body populations (asteroid belts, Kuiper belt, etc.).
-Here we ask: \textit{what is the typical RMS deviation in angular or radial
-position that a typical MBA accrues because of its gravitational
+ the asteroid belt(s) and Kuiper belt.
+Here we ask: \textit{what is the RMS ephemeris error in angular or radial
+position a typical MBA accrues because of its gravitational
 encounters with other individual MBAs?}  We are interested not in the
@@
 for the full population; rather the Brownian motion that accumulates
-from closer encounters either with MBAs that \emph{are not included} in the
-ephemeris model at all.  For those larger asteroids that \emph{are} included
-in an ephemeris model, we need to consider the uncertainties in
-orbital motion that accrue from the uncertainty $\sigma_M$ in their
-masses.  Any such departures became a source of noise in any
+from closer encounters with: (a) MBAs that \emph{are not included} in the
+ephemeris model at all; and (b) asteroids that \emph{are} included
+in an ephemeris model, but have some uncertainty $\sigma_M$ in their masses.
+Any such departures became a source of noise in any
 inferences that we make from use of these MBAs as
-dynamical tracer particles.
+dynamical tracer particles.  We aim to quantify this noise, as a
+function of the RMS uncertainty $\sigma_M$ on the masses of individual
+asteroids, and assuming that deflectors with $M<\sigma_M$ are not
+included in the ephemeris at all.
+
+In principal, one can collect all of the positional information for
+all $>10^6$ known solar-system bodies, and fit them to a model with
+a free state vector and mass for each body.  Calculating the influence
+of $>7\times10^6$ free parameters on $10^8$ or so observables is,
+however, unlikely to be computationally feasible, and most of the
+bodies' masses will be lost in measurement noise and/or be degenerate
+with other bodies' masses.  The most ambitious attempt at a global
+solution that we are aware of is that of \citet{goffin}, who was able
+to include 250 MBAs as gravitating bodies in the fit to the full
+corpus of observations at the time.  For only $\approx 1/2$ of the
+bodies did the fit yield masses $>3\times$ the uncertainties.  With
+vast improvements in data and computing power since then and in the
+next decade, the number of detectable
+asteroid masses should increase, but there will always be
+uncertainties that accrue from the unmodelled deflectors and the
+finite $\sigma_M$ of modelled deflectors.  These remnant uncertainties
+will in turn serve as noise in attempts to model new sources of
+acceleration.
 
@@
 incompleteness of current surveys at lower masses---or more precisely,
-at fainter absolute magnitudes $H,$ since only a handful of small
-asteroids have usefully accurate mass measures.
+at fainter absolute magnitudes $H,$ since the $H$ distribution is
+known far more accurately than the $M$ distribution.
 
 \section{Calculation overview}
-We aim to calculate the Brownian motion variance to an accuracy of
-tens of percent.  Greater accuracy in the calculation is not warranted
-since one of the principal inputs---the number of MBAs vs mass---is
-uncertain to at least this level because of unknowns in the conversion
-from $H$ to mass.  We will therefore be at liberty to drop terms that
-modify the perturbations by $O(e^2)$ of the tracer asteroid that is
-being tracked.    We will assume throughout that the joint
+We aim to calculate the Brownian motion variance to an accuracy of a
+factor $\approx 2.$
+Greater accuracy in the calculation is not warranted
+since some of the principal inputs have substantial uncertainty.  The
+number of MBAs vs mass is 
+uncertain because of unknowns in the conversion
+from $H$ to mass; the status of future observations and $\sigma_M$ is
+also unclear.  Hence we can determine our quantities only to leading
+order in the MBA orbital eccentricity $e$ of the tracer population.
+We will assume throughout that the joint
 distribution of mass (or $H$) and orbital elements is separable into a
@@
 
-All distances will be given in units of the original semimajor axis $a_0$, and all velocities in units of the circular velocity $\vcirc \equiv \sqrt{G\Msun/a_0}.$  In these units, $G\Msun=1,$ the period of the initial orbit is $2\pi$ and the (unperturbed) mean anomaly is $M=t+M_0.$
+All distances will be given in units of the original semimajor axis $a_0$, and all velocities in units of the circular velocity $\vcirc \equiv \sqrt{G\Msun/a_0}.$  In these units, $G\Msun=1,$ the period of the initial orbit is $2\pi,$ and the (unperturbed) mean anomaly is $M=t+M_0.$
 
@@
 approximating it as constant within each of three subsets of tracer MBAs:
-the inner, middle, and outer populations bounded by the 4:1, 3:1,
+the inner, middle, and outer regions bounded by the 4:1, 3:1,
 5:2, and 2:1 mean-motion resonances with Jupiter.
@@
 \begin{equation}
-  \left \langle n I_\alpha^2 \right\rangle \equiv \int d^2I \frac{dN}{dt\,dI_r\,dI_\phi\,dI_z} I_\alpha^2.
+  \left \langle n I_\alpha^2 \right\rangle \equiv \int d^3I \frac{dN}{dt\,dI_r\,dI_\phi\,dI_z} I_\alpha^2.
 \label{eq:nvsq}
 \end{equation}
-The average of the cross terms $I_rI_z$ and $I_\phi, I_z$ will vanish if the deflector distribution is symmetric in inclination, and we find numerically that the mean $I_rI_\phi$ is small enough to ignore.
+The average of the cross terms $I_rI_z$ and $I_\phi I_z$ vanishes if
+the deflector distribution is symmetric in inclination as expected, and we find numerically that the mean $I_rI_\phi$ is small enough to ignore.
 
-From this knowledge of the impulse distribution, our goal is to obtain the covariance matrix of the deviations in the target's observed position, relative to the initial orbit, after some time $T.$ The position observables are the range, plus the latitude and longitude of the target.
+From this knowledge of the impulse distribution, our goal is to obtain
+the covariance matrix of the deviations of the target's position from
+the initial orbit, after some time $T.$ The position observables are
+the range, plus the celestial latitude and longitude of the target.
 We will simplify our results by assuming a heliocentric observer, so that the observational position vector is $\vecp\equiv (r,\theta=z/r,\phi).$ The quantity we seek is the covariance matrix $\covm^p$ of the observations attributable to the accumulated gravitational perturbations:
@@
 \end{equation}
-where the angle brackets indicate an average over possible realizations of the impulse history, and the mean anomaly $M$ at the time of observation.
+where the angle brackets indicate an average over possible
+realizations of the impulse history, and over the mean anomaly $M$ at the time of observation.
 To do so, we will introduce an intermediate set of 6 parameters $\vecq$ describing the deviations of the orbital elements induced by the impulses.
 The $\vecq$ components will be selected to be equal to zero before any impulses are applied, and each responds linearly to $\vecI$ at $|I|\ll 1.$
-We will derive the matrix $\matA$ that describes the orbital-element shifts at time $t$ that arise from an impulse at time $t_i$: 
+We will derive the matrix $\matA$ that describes the orbital-element shifts at time $T$ that arise from an impulse at time $t_i$: 
 \begin{equation}
-  \Delta \vecq(t,t_i) = \matA(e_0, t, t_i) \cdot \vecI.
+  \Delta \vecq(T,t_i) = \matA(e_0, T, t_i) \cdot \vecI.
   \label{eq:A}
 \end{equation}
-In our first-order perturbation theory, $\vecq(t)$ will be the sum of the $\Delta\vecq(t_i)$ imparted by all impulses applied at times $0<t_i<t.$  Because the impulses are uncorrelated, $\vecq$ will therefore be the result of a random walk.
-The distribution will have a covariance matrix $\covm^q(t) \equiv \left\langle \vecq(t) \vecq^T(t) \right\rangle$ whose elements are
+In our first-order perturbation theory, $\vecq(T)$ will be the sum of the $\Delta\vecq_i$ imparted by all impulses applied at times $0<t_i<T.$  Because the impulses are uncorrelated, $\vecq$ will therefore be the result of a random walk.
+The distribution will have a covariance matrix $\covm^q(T) \equiv \left\langle \vecq(T) \vecq^T(T) \right\rangle$ whose elements are
 \begin{eqnarray}
-  \covm^q_{jk}(e_o,t) & = & \left\langle \sum_{i,\gamma} \matA_{j\gamma}(e,t,t_i) \matA_{k\gamma}(e,t,t_i) I^2_{i,\gamma} \right\rangle \\
-           & = & \sum_\gamma \int dt_i \matA_{j\gamma}(e,t,t_i) \matA_{k\gamma}(e,t,t_i) \left\langle n I_\gamma^2\right\rangle.
+  \covm^q_{jk}(e_o,T) & = & \left\langle \sum_{i,\gamma} \matA_{j\gamma}(e_0,T,t_i) \matA_{k\gamma}(e_0,T,t_i) I^2_{i,\gamma} \right\rangle \\
+           & = & \sum_\gamma \int dt_i \matA_{j\gamma}(e_0,T,t_i) \matA_{k\gamma}(e_0,T,t_i) \left\langle n I_\gamma^2\right\rangle.
 \label{eq:Cqjk}
@@
 
-The last element of our calculation will be a conversion from the element shifts $\vecq$ into the observed displacements $\Delta\vecp$.
-In linear perturbation theory this will again be expressible as a matrix
+The last element of our calculation will be a conversion from the
+element shifts $\vecq$ into the observed displacements $\Delta\vecp$
+at time $T.$  In linear perturbation theory this will again be expressible as a matrix
 \begin{equation}
-  \Delta\vecp(e_0,t) = \matB(e_0,t) \cdot \vecq(t).
+  \Delta\vecp(e_0,T) = \matB(e_0,T) \cdot \vecq(T).
 \label{eq:B}
@@
 \begin{eqnarray}
-  \covm^x_{\alpha\beta}(t,e_0) & = & \sum_{jk} B_{\alpha j}(e,t) B_{\beta k}(e,t) \covm^q_{jk}(t) \nonumber \\
-  & = & \sum_\gamma \left\langle nI_\gamma^2\right\rangle \int dt_i \left[AB(e_0,t,t_i)\right]_{\alpha\gamma}  \left[AB(e_0,t,t_i)\right]_{\beta\gamma} 
+  \covm^p_{\alpha\beta}(e_0,T) & = & \sum_{jk} B_{\alpha j}(e_0,T) B_{\beta k}(e_0,T) \covm^q_{jk}(T) \nonumber \\
+  & = & \sum_\gamma \left\langle nI_\gamma^2\right\rangle \int dt_i \left[AB(e_0,T,t_i)\right]_{\alpha\gamma}  \left[AB(e_0,T,t_i)\right]_{\beta\gamma} 
 \label{eq:ABq}
@@
   nI_\alpha^2\right\rangle$ for MBA regions.
-Section~\ref{sec:propagation} derives the forms of $\matA(e,t,t_i),$
-and $\matB(e,t).$  Section~\ref{sec:results} combines these and
+Section~\ref{sec:propagation} derives the forms of $\matA(e_0,T,t_i),$
+and $\matB(e_0,T).$  Section~\ref{sec:results} combines these and
 summarizes the results and their implications.
@@
                                               dH\,dv\,db\,d\theta\,d\phi
-                                              \frac{dn}{dH\,dv\,db\,d\theta\,d\phi}
+                                             \frac{dN}{dH} \frac{dn}{dv\,db\,d\theta\,d\phi}
                                               I^2_\gamma(H,v,b,\theta\phi)
@@
         \frac{dn}{dA\,dv\,d\theta\,d\phi} \hat b_\gamma^2(\theta,\phi)
-        \int db_{b_{\rm min}}^{b_{\rm max}} 2\pi b\,db\,
+        \int_{b_{\rm min}}^{b_{\rm max}} 2\pi b\,db\,
         \left(\frac{\sigma_M(H)}{bv}\right)^2 \\
-  & = & \int dH \frac{dN}{dH} \sigma_M(H)^2  \int dv
+  & = & 2\pi \int dH \frac{dN}{dH} \sigma^2_M(H)  \int dv
         \frac{dn}{dA\,dv} v^{-2} \left\langle\hat b_\gamma^2(v)\right\rangle
-        2\pi \log(b_{\rm max}/b_{\rm min}).
+        \log(b_{\rm max}/b_{\rm min}).
 \end{eqnarray}
@@
 between a deflector-tracer pair, as a function of the impact velocity
-and geometry is $dn/dA\,dv\,d\theta\,d\phi,$ which we will determine
+and geometry, is $dn/dA\,dv\,d\theta\,d\phi,$ which we will determine
 through numerical measurements with the orbits of all known MBAs. The
 $H$ distribution of MBAs is $dN/dH.$ We denote as $\sigma_M(H)$
-the RMS error in the mass of MBAs at $H$ used in the ephemeris model.
+the RMS error in the mass of MBAs at $H$ from the value used in the ephemeris model.
 We use $\sigma_M$ rather than $M$ since we are seeking the size of
@@
 this mass could be determined from fitting the data of the tracer.
-Fitting an ephemeris that includes all of these Individually
-detectable encounters is entirely feasible, as long as the relevant
-MBAs are known, with modestly accurate orbits.
 Such encounters would thus no longer be considered source of
-stochastic Brownian motion.  If we crudely choose some threshold
+stochastic Brownian motion.  Finding all such encounters among known
+MBAs is feasible, and in fact is forecasted for the next decade's LSST
+observations by
+\citet{negin}. The fitting of the tracer data for the mass of the
+deflector is also a small computational effort.
+
+If we crudely choose some threshold
 $I_{\rm det}$ of impulse as being large enough to generate high-$S/N$
@@
 The lower bound on the $v$ integral marks the speed below which
-$b_{\rm max}<b_{\rm min}.$
+$b_{\rm max}<b_{\rm min}.$  Since $b_{\rm min}$ and $b_{\rm max}$
+appear logarithmically, the estimated diffusion is relatively insensitive to choices of
+their values.
 
@@
 \eqq{eq:nI2} reduces the calculation to a double integral over the
-deflector $H$ distribution and the relative velocity $v.$  The
+deflector $H$ and the tracer-deflector relative velocity $v.$  The
 constituents of this calculation are estimated as follows.
@@
 pair-finding, we locate all tracer-deflector pairs
-that approach within 0.03~AU of each during the $\pm1$-day period
+that pass within 0.03~AU of each other during the $\pm1$-day period
 around this initial epoch, assuming inertial relative motion during
@@
 \begin{equation}
-  M(H) = 1.2\times10^{-17} \times 10^{-0.6(H-15)}\, M_\odot.
+  M(H) = 1.2\times10^{-17} M_\odot \times 10^{-0.6(H-15)}
 \end{equation}
@@
 the functional form for the cumulative MBA counts given by
-initial LSST predictions \citep{LSST_science_book}:
+initial LSST predictions \citep{lsstbook}:
 \begin{equation}
@@
 The two largest asteroids, Ceres and Vesta, have masses determined to
-high precision by the \textit{Dawn} spacecraft.  Thus despite holding
-half of the mass of the asteroid belt, their $\sigma_M$ values are just
-$\approx3\times10^{-15} M_\odot$ and
-the unknown part of their contributions to other MBA's motion are
-quite small.  Aside from a handful of other asteroids in
+high precision by the \textit{Dawn} spacecraft \citep{Ceres,Vesta}.  Thus despite holding
+half of the $\approx 10^{-9} M_\odot$ mass of the asteroid belt, their $\sigma_M$
+values\footnote{More precisely, the uncertainties in their $GM$ values
+  relative to $GM_\odot$.} are 
+$\le3\times10^{-15} M_\odot$ and
+% Konilov papers: 1.2e-5 km^3/s^2 for Vesta, 4e-4 for Ceres
+this unknown part of their contribution to other MBA's motion is 
+unimportant.  Aside from a handful of other asteroids in
 well-characterized binary systems or visited by spacecraft, most other knowledge of MBA masses
@@
 few MBAs on which it imparts the largest impulses.  The accuracy
-$\sigma_M$ of this determination will depend upon the geometry of
-these encounters, the measurement errors on the tracers, and the size
-of the noise from unmodelled mass that we aim to determine.  These
+$\sigma_M$ of this determination will vary, depending upon the
+geometry  and timing of
+each deflector MBA's encounters, and the measurement errors on its tracers.  These
 factors are independent of the mass of the deflector.  We will
@@
 
-To give some context, the most recent
+To give some context, a
 comprehensive investigation of MBA masses from mutual scattering
-events was by \citet{baer}, who report typical $\sigma_M\approx
+events by \citet{goffin} reports typical $\sigma_M\approx
 5\times10^{-13}$ for 30 MBAs.  We can expect substantial improvement
@@
 
-\subsubsection{Results}
-
 \section{Propagation to position shifts}
@@
 Between impulses, $\tau$ advances with the mean motion as $a^{-3/2}t.$  We define a coordinate system that is uniformly rotating with the unperturbed $\tau_0=t+M_0.$ The two smoothly rotating unit vectors $\uhat_\parallel=(\cos \tau_0, \sin \tau_0)$ and
-$\uhat_\perp=(-\sin \tau_0, \cos \tau_0)$ satisfy $\uhat_\parallel \times \uhat_\perp = \zhat,$ and we will use subscripts $\parallel$ and $\perp$ to represent projections onto these components.
-In particular, the components of the ellipticity vector $e_\parallel$ and $e_\perp$ have the values $e_\parallel=e\cos M$ and $e_\perp=-e\sin M$ in the unperturbed orbit.
-The true anomaly $\nu,$ radius $r$, and azimuthal angle $\phi,$ and the velocity components of the initial orbit to $O(e)$ are
+$\uhat_\perp=(-\sin \tau_0, \cos \tau_0)$ satisfy $\uhat_\parallel
+\times \uhat_\perp = \zhat,$ and we will use subscripts $\parallel$
+and $\perp$ to represent projections onto these components.  In
+particular, the components $(e_\parallel,e_\perp)$ of the ellipticity
+vector $\vece$ have the values $e_\parallel=e\cos M$ and $e_\perp=-e\sin M$ in the unperturbed orbit.
+The true anomaly $\nu,$ radius $r$, and azimuthal angle $\phi,$ and
+the velocity components of the initial orbit are, to first order in $e$,
 \begin{eqnarray}
@@
 \begin{eqnarray}
-  \Delta\vece & = & (2\rhat - rv_r \phat) I_t - \phat I_r \\
+  \Delta\vece & = & (2\rhat - rv_r \phat) I_\phi - \phat I_r \\
 \label{eq:depar}
@@
 
-The final element of $\vecq$ will be the shift $\Delta\tau$ induced in $M+\lop$ by the impulse.
-This can be obtained by enforcing the conditions that the radius $r$ or the azimuthal angle $\phi$ must remain constant during the impulse.
-To findthe latter, we need a formula for $\nu(M)$ as a power series in $e.$  Standard formulae in the literature lead to:
+The final element of $\vecq$ will be the shift $\Delta\tau$ induced in
+$M+\lop$ by the impulse.  This can be obtained by enforcing the
+conditions that the radius $r$ or the azimuthal angle $\phi$ must
+remain constant during the impulse.  To find the latter, we need a formula for $\nu(M)$ as a power series in $e.$  Standard formulae in the literature lead to:
 \begin{eqnarray}
@@
 \end{eqnarray}
-We wish to find the first-order shift in $\phi=\nu + \lop$ and set it to zero.
+We wish to find the first-order shift in $\phi=\nu + \lop$ and set it
+to zero.
 But the derivatives of $M$ and $\lop$ with impulse can become infinite.
@@
   M & = & \tau-\lop = \tau_0 -\lop + \Delta\tau \\
-  e \sin M & = & e \sin (\tau_0 -\lop) + \Delta\tau \left[ e \cos  (\tau_0 -\lop)\right] \\
+  e \sin M & = & e \sin (\tau_0 -\lop) + \Delta\tau \left[ e \cos  (\tau_0 -\lop)\right] \nonumber \\
+\label{eq:esinM}
     & = & -e_{0\perp}-\Delta e_\perp + e_{0\parallel} \Delta\tau +O(e^2) \\
-  e \cos M & = & e \cos (\tau_0 -\lop) - \Delta\tau \left[ e \sin  (\tau_0 -\lop)\right] \\
+  e \cos M & = & e \cos (\tau_0 -\lop) - \Delta\tau \left[ e \sin  (\tau_0 -\lop)\right] \nonumber\\
+\label{eq:ecosM}
     & = & e_{0\parallel}+\Delta e_\parallel + e_{0\perp} \Delta\tau +O(e^2) \\
@@
                               + \frac{5}{2}\left( -e_{0\perp}\Delta e_\parallel-e_{0\parallel}\Delta e_\perp\right) + O(e^2).
+\label{eq:phi}
 \end{eqnarray}
@@
 
-This equation gives the last row of the $\matA$ matrix giving element-level perturbations from individual impulses.  The preceding equations can be restated to give the other rows:
+We now have all the information we need to create the $\matA$ matrix.
+Rotating the $\vecq$ elements back into the Cartesian frame, we
+obtain, to $O(e)$,
+\begin{equation}
+  \left( \begin{array}{c}
+           \Delta a \\ \Delta\tau \\ \Delta e_x \\ \Delta e_y \\ \Delta \Lhat_x \\ \Delta\Lhat_y
+         \end{array}\right) = \left( \begin{array}{ccc}
+ 2e\sin\tau & 2 + 2e\cos\tau & 0 \\
+ -2+e\left[1.5\cos\tau-3(T-t_i)\sin\tau\right] & e\sin\tau -3(1+e\cos\tau)(T-t_i) & 0\\
+ -\cos\tau-2e\sin^2\tau & 2\cos\tau - 1.5e\sin 2\tau & 0\\
+  -\sin\tau + e\sin 2\tau & 2\sin\tau-3e\cos^2\tau & 0 \\
+0 & 0 & \sin\tau + 0.5 e \sin 2\tau \\
+ 0 & 0 &  -\cos\tau +e(3-\cos 2\tau)/2 \\
+ \end{array}\right)
+  \left(\begin{array}{c} I_r \\ I_\phi \\ I_z \end{array}\right)
+\end{equation}
+We can now do the time integral and sum over impulse directions in \eqq{eq:Cqjk}, and average over $\tau$ to average over the orbital phase of the impulses. The diagonal elements are
 \begin{eqnarray}
-  \Delta a & = & -2e_\perp I_r + 2(1+e_\perp) I_\phi, \\
-  \Delta \Lhat_\parallel & = & -2e_\perp I_z \\
-  \Delta \Lhat_\perp & = & -(1-e_\parallel) I_z \\
-  \Delta e_\parallel & = &  -2e_\perp I_r + 2 I_\phi \\
-  \Delta e_\perp & = & -I_r - 3e_\perp I_\phi.
+  \Var(a) & = & 4T \nIphi \nonumber\\
+  \Var(\tau) & = & 4T\nIr + 3T^3\nIphi \nonumber\\
+  \Var(e_x) =  \Var(e_y) & = & \nIr/2+2\nIphi \nonumber\\
+  \Var(\Lhat_x) =  \Var(\Lhat_y) & = & \nIz/2
+\label{eq:Cq}
 \end{eqnarray}
+All of the off-diagonal terms (covariances) are either zero or $O(e)\left\langle nI^2\right\rangle,$ and they result in observable consequences that are smaller by $\left\langle e^2 \right\rangle$ than the diagonal terms' contributions.
 
+% \begin{equation}
+%   \matC^2 = \left(\begin{array}{cccccc}
+%     4T \nIphi & -3T^2\nIphi & e\left(\nIr + 5\nIphi\right) & e\nIr & 0 & 0 \\
+% -3T^2\nIphi & 4\nIr + 3T^3\nIphi & -e(11T\nIr/4 + 15T^2\nIphi/4) & e\left[(2T-3T^2/4)\nIr+(-T+9T^2/4)\nIphi\right] & 0 & 0\\
+%  e\left(\nIr + 5\nIphi\right) & -e(11T\nIr/4 + 15T^2\nIphi/4) & \nIr/2+2\nIphi & 0 & 0 & 0 \\
+%  e\nIr &  e\left[(2T-3T^2/4)\nIr+(-T+9T^2/4)\nIphi\right]  & 0 & \nIr/2+2\nIphi & 0 & 0 \\
+% 0 & 0 & 0 & 0 & \nIz/2 & 0 \\
+% 0 & 0 & 0 & 0 & 0 \nIz/2 \\
+%\end{array}\right)
+% \end{equation}
  \subsection{Astrometric and ranging shifts from elements shifts}
 \label{sec:observe}
+The radial coordinate of a Keplerian orbit is $r=a(1-e\cos E),$ and an expansion of the eccentric anomaly $E$ in powers of $e$ is
 \begin{equation}
@@
 \end{equation}
+From this we can derive the perturbations to $r$ as a function of our chosen basis for perturbations to the orbital elements.  Making use of Equations~(\ref{eq:ecosM}) and (\ref{eq:esinM}) we obtain
+\begin{equation}
+  \Delta r = (1-e\cos\tau) \Delta a + e\sin\tau \Delta\tau +  (-\cos\tau -2e\sin^2\tau) \Delta e_x + (-\sin\tau+e\sin 2\tau) \Delta e_y.
+\end{equation}
+If we square this expression, average over the orbital phase $\tau$ of the observation, and retain leading order in $e$, we get the variance of the radial coordinate from the unperturbed orbit.
+All of the terms involving covariances between orbital elements either average to zero over an orbit, or are suppressed by $e^2$ relative to other terms.
+We are left with
+\begin{eqnarray}
+  \Var(r) & \approx & \Var(a) + \frac{1}{2}\Var(e_x) + \frac{1}{2}\Var(e_y) + \frac{\langle e^2\rangle}{2} \Var(\tau) \\
+          & \approx & \frac{T}{2} \nIr + \left( 6T + \frac{3\langle e^2 \rangle}{2} T^3\right) \nIphi.
+                      \label{eq:varr}
+\end{eqnarray}
+The second line substitutes the results from Equations~(\ref{eq:Cq}).
+We retain the final term because, as $T>>1,$ this precession term that scales with $\langle e^2 \rangle T^3$ becomes comparable to or larger than the other terms that scale as $T$.
+The $T$-scaling terms all arise from epicyclic motions that are periodic with the orbit.
 
-Keeping only terms from Equations~(\ref{eq:kepler}) up to first order in $\vecq,$ the true anomaly $\nu,$ mean anomaly $M=t + \Delta t - \lop,$ and the azimuthal angle $\phi$ are related by
+The heliocentric azimuthal angle $\phi$ is given by \eqq{eq:phi}.  We can similarly square the perturbation and average over orbital phase and population factors to get
 \begin{eqnarray}
-  \nu  =  M + 2e\sin M  & = & t +  2e_0 \sin t  + \Delta t - \lop + 2 e_0(\Delta t - \lop) \cos t + 2\sin t \Delta e_x \\
-  \Rightarrow \quad \phi  =  \nu + \lop & = &  \left[t + 2e_0\sin t\right] + \left[(1+2e_0\cos t) \Delta t - 2\cos t \Delta e_y + 2\sin t \Delta e_x\right]
-                                                    \label{eq:phi}
+  \Var(\phi) & \approx & \Var(\tau) + 2\Var(e_x) + 2\Var(e_y) \\
+             & \approx & 6T\nIr + \left(8T+3T^3\right) \nIphi.
+                         \label{eq:varphi}
 \end{eqnarray}
-where we have used $e_0\lop = \Delta e_y.$ The radial coordinate and polar angle $\theta$ are
+The heliocentric latitude (or equivalently, polar angle) has a deviation
 \begin{eqnarray}
-  r & = & (1 + \Delta a) \left[1- (e_0+\Delta e_x) \cos M \right] \nonumber \\
-\label{eq:r}
-  & = & \left[ 1-e_0\cos t \right] + \left[ (1-e_0\cos t) \Delta a + e_0\sin t \Delta t - \sin t \Delta e_y - \cos t \Delta e_x \right] \\
-  \theta = z/r & = & -\Lhat_x \cos\phi - \Lhat_y \sin\phi \nonumber \\
-\label{eq:theta}
-               & = & \left[ 0 \right] + \left[ -\cos t \Lhat_x  - \sin t \Lhat_y\right].
+  \Delta\theta & = & -\Delta\Lhat_x \frac{x}{r} - \Delta\Lhat_y \frac{y}{r} \\
+  \Rightarrow \quad \Var(\theta) & \approx & \langle\cos^2\tau\rangle \Var(\Lhat_x) + \langle\sin^2\tau\rangle \Var(\Lhat_y) \\
+               & \approx & \frac{T}{2} \nIz.
+                           \label{eq:vartheta}
 \end{eqnarray}
-The second line of each deviation is divided into two bracketed terms, the first being the initial orbit and the second being the first-order terms in elements of $\vecq,$ which are therefore the elements of $\matB$.
 
-A heliocentric observer will measure ``range noise'' of $\covm^p_{rr}=\langle (\Delta r)^2\rangle$ in the measured distance to the target asteroid.
-We find this by squaring the second bracketed term for $r$ and averaging $t$ over an interval of size $2\pi$ to compensate for the fact that we have assumed $M=0$ at $t=0,$ whereas in fact the $M_0$ values are random.
-We can then derive
-\begin{equation}
-  \left\langle(\Delta r)^2\right\rangle = (\Delta a)^2 + \frac{1}{2}\left[ (\Delta e_x)^2 + (\Delta e_y)^2\right] + \frac{e_0}{2}(\Delta a \Delta e_x -  \Delta t \Delta e_y).
-  \label{eq:varr}
-\end{equation}
-The same heliocentric observer will see angular shifts of the target position satisfying
-\begin{eqnarray}
-  \label{eq:phiphi}
-  \left\langle (\Delta\phi)^2\right\rangle & = & (\Delta t)^2 + 2(\Delta e_x^2 + \Delta e_y^2)  -4e_0
-  \left(\Delta t \Delta e_y\right)\\
-  \label{eq:thetatheta}
-  \left\langle (\Delta\theta)^2\right\rangle & = & \frac{1}{2} \left( \Lhat_x^2 + \Lhat_y^2\right) \\
-  \label{eq:thetaphi}
-  \left\langle \Delta\theta\Delta\phi\right\rangle & = & -e_0\Delta t \Lhat_x + (\Delta e_y \Lhat_x -  \Delta e_x \Lhat_y)
-\end{eqnarray}
+Equations~(\ref{eq:varr}), (\ref{eq:varphi}), and (\ref{eq:vartheta}) comprise our desired result.  Several characteristics are noteworthy:
+\begin{itemize}
+\item $r$ is in units of $a_0$ and $\theta, \phi$ are angles.
+To convert their variances into linear displacements in the $r,\phi,z$ directions, multiply them all by $a^2,$ or equivalently scale the standard deviations by $a.$
+\item The terms proportional to $T$ show typical diffusive growth $\Delta \propto \sqrt{T},$ and arise from perturbations to the 5 constants of Keplerian motion.
+They represent epicyclic alterations to the original orbit that grow in amplitude as $\sqrt{T}.$  The $T^3$ terms dominate $\Var(\phi)$ by the end of a full orbit, resulting from accumulating delays/advances $\Delta\tau$ in orbital phase $M$ as $a$ and the mean motion diffuse.
+Thus it is the energy-changing impulses, namely $I_\phi,$ that cause the most Brownian motion.
+\item The radial variance also becomes dominated by the $T^3$ term if $e>0,$ as this term represents advances/delays in the radial epicycle phase.
+This is the only part of the Brownian motion that whose leading order in $e$ is non-zero, \ie\ which we would have missed by assuming circular initial orbits.
+\item The RMS vertical displacement grows strictly as $T^{1/2}.$
+\item It will be generally true that $\Var(\phi) > \Var(r) > \Var(\theta),$ a trend exacerbated by the empirical observation (Figure~\ref{fig:rtz}) that impulses favor the $\phat$ direction.
+\item There are no significant covariances among the $r,\theta,$ and $\phi$ displacements at the population-averaged level.
+  \item For Earth-based observers, the deviations in $(r,\theta,\phi)$ from the nominal orbit will be some calculable, time-dependent, nearly-orthogonal transformation of the heliocentric deviations derived above, so the results will not be grossly different aside from bringing the three directional components closer together in amplitude.
+\end{itemize}
 
@@
 \label{sec:results}
+We can now propagate the numerical estimates for $\left\langle nI_\gamma^2\right\rangle$ from Section~\ref{sec:impulse} through the analytical observational consequences in Section~\ref{sec:observe} to yield estimates of the unmodelled RMS shifts of MBAs from each region.
+We assume the following parameters:
+\begin{itemize}
+\item The deviations are accumulated over a time $T=10$~yr.
+\item The population has $\langle e^2 \rangle=0.03,$ which is the value for the cataloged MBAs of the Inner and Middle belts, and a slight overestimate for the Outer belt.
+\item The $5\sigma$ impulse size $I_{\rm det}$, at which we deem it possible to isolate the effect of a single deflector on a tracer and hence usefully include that deflector's mass in the ephemeris model, is taken to be $I_{\rm det} = 10^{-8}v_c.$ In practice this will depend on the geometry of the impulse and on the quality of the tracer's observations.
+But we find that changing $I_{\rm det}$ by a factor of 10 alters the noise from unmodeled impulses by $<20\%.$
+  \item We will vary $\sigma_M,$ the RMS uncertainty on the mass of modelled asteroids, such that the number $N$ of MBAs  with $M>\sigma_M$ varies from 100 to 10,000.
+The number of MBAs included in the ephemeris model would be $\approx N.$
+\end{itemize}
 
-\citet{negin} forecast the extent to which LSST data will be able to
-constrain the masses of individual MBAs by measuring deflections of
-the $\approx 10$ tracer asteroids they encounter most closely over the
-course of the nominal 10-year survey.  They find that the $1\sigma$
-statistical error on a typical MBA is $\sigma_M=???\, M_\odot.$
+\begin{figure}
+  \centering
+  \includegraphics[width=0.6\textwidth]{result.pdf}
+  \caption{Each line shows the forecast of the variance of the
+    azimuthal position of a Middle Main Belt asteroid caused by discrete
+    encounters with absolute magnitude $H$ greater than the value on
+    the (lower) $x$ axis, \ie\ the cumulative noise from all MBAs
+    smaller than the plotted $H_{\rm min.}$  The upper axis gives mass
+    instead of $H$.  The left axis gives values in meters, the right
+    the apparent shift in milliarseconds for a heliocentric observer.
+    Each line is for a different assumption about which asteroids' deflections are
+    part of the ephemeris model instead of being considered as noise: the $\sigma_M$
+    value gives the RMS mass uncertainty on modelled MBAs, and the $N$
+    is the number of MBAs with $M>\sigma_M$ that could be usefully
+    modelled.}
+  \label{fig:result}
+\end{figure}
 
-$\sigma_M(H)$ rolled off at ...
+Figure~\ref{fig:result} plots the resulting noise level $\sigma_\phi=\sqrt{\Var{\phi}}$ of the calculation for the Middle regions of MBAs, showing both the linear RMS displacements on the left axis and the angular displacements on the right axis.
+Each point on the curve is the $\sigma_\phi$ attributable to deflector MBAs that are smaller (fainter) than the $H$ value on the lower axis, or the mass on the top axis.
+Thus the left-most point on each curve marks the total noise integrated over the full population of deflectors.
+Different color lines show the results for the values of $N$ and $\sigma_M$ as marked astride each line.
+Notable features of these result are:
+\begin{itemize}
+\item We find that \emph{the linear displacements are nearly identical for Inner, Middle, and Outer regions of the asteroid belt.}  The angular displacements vary slightly, inversely to their distance from the Sun.
+\item We plot the azimuthal noise.
+\emph{The radial noise is $\approx 7\times$ lower than the azimuthal noise, and the vertical Brownian motion is $\approx45\times$ lower than the azimuthal, on average.}
+\item For longer time periods $T$, the azimuthal noise will scale as
+$T^{3/2},$ the radial noise slightly below this rate, and the vertical
+noise as $T^{1/2}.$
+\item The largest contributions to the unmodelled Brownian motion are
+made by deflectors with mass and $H$ at the ``shoulder'' of each
+curve, which will also correspond to the MBAs with masses $M\sim
+\sigma_M$.  Above this mass, the noise per deflector is constant but
+the deflectors become scarce at larger $M.$ Below this mass, $dN/dH$
+is rising to fainter/smaller bodies, but the factor $M^2(H) dN/dH$
+that determines the Brownian motion is dropping.
+\end{itemize}
+
+We take as representative of the present state of the art the case
+where the $N=100$ most massive MBAs (aside from Ceres and Vesta) have
+masses determined to $\sigma_M\approx 10^{-12.7} M_\odot$ by mutual
+encounters.  Our calculations suggest then that a 10-year ephemeris
+including these 100 MBAs will accrue RMS errors slightly under 1~km in
+the azimuthal position, or 0.4~mas in observed ecliptic longitude.
+RMS range and vertical ephemeris errors would be closer to 100 and 20
+meters, respectively.  In the next decade, LSST will significantly
+increase the number $N$ and lower $\sigma_M$ of deflector MBAs with
+usefully measured masses from mutual encounters by tracking hugely
+more potential tracers.  We find that, roughly speaking, $\sigma_\phi$
+drops with the number of modeled MBAs as $1/N,$ and
+\begin{equation} \sigma_\phi \equiv \sqrt{\Var{\phi}} \approx
+(100\,\textrm{m}) \times \left(\sigma_M /
+10^{-14}M_\odot\right)^{0.6}.
+  \label{eq:varscale}
+\end{equation}
+
+We need to check whether our supposition that impulses $I>I_{\rm det}$
+would stand out above the measurement noise of a single tracer remains
+true once we include the noise induced by the other unmodelled
+impulses.  Focusing on the effects of a tangential impulse of
+$I_\phi=I_{\rm det}/\sqrt{2},$ \eqq{eq:dtau} predicts an orbital phase
+shift of $\Delta\tau \approx 3TI_\phi$ at time $T$ past the impulse,
+translating directly into a position shift $\Delta\phi$ of the same
+size as per \eqq{eq:phi}.  Consider a survey of duration $T,$ and a
+single impulse $I_{\rm det}$ applied at time $T/2.$ If
+$I_\phi=10^{-8}/\sqrt{2}$ and $T=10$~yr, then
+$\Delta\phi\approx35$~mas.  This is nearly $100\times$ larger than the
+stochastic $\sigma_\phi$ expected from unmodelled masses at the
+current state of knowledge.  For primarily radial or vertical
+impulses, the astrometric signal will be smaller, but the Brownian
+motion noise is lower too.  Since $\Delta\phi/\sigma_\phi \propto
+T^{-1/2},$ this signal-to-noise ratio would still be $\gg 5$ for
+surveys of several decades.  We conclude that our choice of $I_{\rm
+det}$ is not ruined by the presence of Brownian motion noise on the
+tracer asteroids, rather it will be driven by the observational
+uncertainties.
+
+\subsection{Implications for future investigations}
+
+An important question is: will the positional noise due to unmodelled MBA encounters degrade inferences being made with measurements of tracer MBAs?
+If the Brownian noise is below other sources of measurement noise, then the answer must be ``no.''
+The per-epoch uncertainties on solar-system object astrometry in Gaia DR3\footnote{We take one Gaia transit as an ``epoch'' here.} are reported to be as low as 0.3~mas for very bright sources ($G=13$), rising to $\approx10$~mas at its $G=21$ limit where most detections lie \citep[][Figure 6]{gaiass3}.
+The overall RMS residual to the orbit fits are 5~mas (this is in the ``along-track'' direction).
+Thus the Brownian motion of $\approx0.4$~mas we estimate for $\phi$ (over a 10-year period) would be important for the single-epoch error budget of only the brightest Gaia MBAs.
+The Brownian motion is, however, highly correlated across epochs, whereas measurement errors are independent and decrease as the square root of the number of epochs, which is $\approx20$ over 3 years for a typical MBA in the Gaia DR3 catalog.
+The Brownian motion might become relevant for analysis of a significant fraction of Gaia MBAs during analysis of the complete 11-year data.
+
+For LSST, the measurement errors on MBA positions will have a floor of 1--2~mas per epoch set by refraction from atmospheric turbulence \citep{willow,trojans}.
+Objects with apparent magnitude $m\gtrsim20$ will have larger per-epoch errors due to photon noise in the images.
+There will, however, be hundreds of observations taken of each MBA over 10~yr, meaning that Brownian motion might be a significant contributor to inferences from the dynamics of brighter tracer MBAs.
+
+On the other hand, Gaia has greatly improved the astrometric precision for $\sim 10^5$ MBAs and LSST will do so for $\sim 10^7$ potential tracer MBAs, significantly lowering $\sigma_M$ and increasing the number $N$ and of deflector MBAs with usefully measured masses from mutual encounters.
+If either can successfully measure masses at $S/N\gtrsim 1$ for the most of the $\ge 1000$ largest MBAs, Figure~\ref{fig:result} indicates reduction of the unmodelled Brownian motion to $\approx0.02$~mas, well below the expected measurement errors for both experiments.
+The prospects for constraining individual MBAs' masses with LSST data on mutual encounters are examined by \citet{negin}, who conclude ????.
+
+\textbf{Fraction of objects generating $1\sigma$ detections in Negin project??}
+
+An even more ambitious proposal by \citet{occult} is to track $\approx10^6$ MBAs down to diameters of $\approx1$~km by watching them occult Gaia stars.
+The timing of each occultation can locate the center of mass of the target MBA along its direction of motion, with most of the information available from objects yielding uncertainties of 50--100~m on this position.
+The Brownian motion noise over a 10-year baseline will in this case become comparable to the per-epoch measurement errors.
+Reducing the typical mass uncertainty of individual asteroids to $\sigma_M\approx 10^{-15.4} M_\odot$ would be needed ($N=3000$) to reduce the Brownian noise to negligible levels for this survey.
+
+The extent to which the Brownian motion can confound an inference also depends on how similar the spatio-temporal dependence of the Brownian motion is to the signal under test.
+If, for example, one is searching for the tidal force of a Planet X, this will be manifested primarily as a collective apsidal precession and a quadrupolar deviation of the orbit from its ellipse that does not evolve with time.
+These are quite distinct from the random walk in $\Delta\tau$ that is the dominant effect of Brownian motion.
+At the other extreme, if one were trying to measure the Yarkovsky effect (which, to our knowledge, has never been detected on an MBA), this will be primarily realized as a time-invariant tangential force (torque) whose amplitude and direction depend on the (unknown) spin axis, the spin rate, and the radiant and thermal properties of the surface, which will lead to a linear increase or decrease in $\Delta\tau$ over the observation period.
+This may be  difficult to discern from the Brownian motion's random walk in $\Delta\tau$ over the same interval.
+The ratio of the Yarkovsky signal to the Brownian motion will increase as $\sqrt{T}$ and will also scale with the inverse diameter of the source.
+There will be some nominal diameter below which the unknown Yarkovsky signal leads to larger observable $\sigma_\phi$ than the $\sigma_\phi$ we have derived from Brownian motion.
+This will be good news for those wishing to measure individual bodies' Yarkovsky effect, and bad news for those wishing to use these bodies as test particles for other gravitational or non-gravitational forces.
+
+In conclusion, the Brownian motion from unmodelled or mis-modelled mutual impulses between MBAs causes RMS azimuthal displacements at the 1~km scale over 10 years, with current knowledge of the masses of large asteroids.
+This is large enough to become a significant part of the error budget for many inferences from the larger/brighter asteroids being used as tracer particles from Gaia or LSST images.
+If, however, we can use mutual events to constrain the masses of the $\approx1000$ largest  asteroids with $S/N\gtrsim1,$ then the unmodelled Brownian motion shrinks to $\approx100$~m for MBAs, which becomes insignificantly small for most uses of Gaia or LSST astrometry.
+Occultation-based astrometry might, at this stage, have measurement errors that are similar to the uncertainties from Brownian motion on decade time scales.
 
@@
 \bibliographystyle{aasjournal}
-%\bibliography{references}
+\bibliography{references}
 

### 2025-02-21 main.tex
include sigmaM forecast

@@
 \subsubsection{Asteroid properties}
+\label{sec:mbaprops}
 We crudely approximate the population as having a geometric albedo of 0.25,
@@
 
+\begin{figure}
+  \plottwo{sigmaI.pdf}{sigmaM.pdf}
+  \caption{Forecasts of uncertainties derived from LSST tracking of the known MBAs.
+The left-hand plot shows the distribution of (log) uncertainties $\sigma_I$ on the amplitude of the impulse applied to the tracer MBA, incorporating forecasted LSST observations and marginalizing over the tracer's initial state vector.
+The right-hand side shows the distribution of the uncertainties on the mass, $\sigma_M=(bv/2)\sigma_I,$ for deflector MBAs expected from LSST.
+The histograms show the results from using only the single tracer yielding the lowest $\sigma_M$ for that deflector; combining the 10 best-$\sigma_M$ tracers; or combining all impulses at $b<0.03$~AU.
+The RMS value of $\sigma_M$ over the deflector population is given in the legend.
+Masses are in solar units, and velocities in units of the circular velocity of the deflector.}
+  \label{fig:sigmaI}
+\end{figure}
+
 To give some context, a
@@
 events by \citet{goffin} reports typical $\sigma_M\approx
-5\times10^{-13}$ for 30 MBAs.  We can expect substantial improvement
+5\times10^{-13}$ for 30 MBAs. But we can expect substantial improvement
 in the LSST era when a millions of potential tracer asteroids attain
 mas-level astrometry.
+\citet{negin} use a simulated LSST sequence of observations to estimate the impulse (and mass) uncertainties from mutual encounters of the known asteroids that will occur during years 2--9 of its 10-year survey.
+The tracer MBAs for those events are essentially an ubiased selection from the known MBAs, and the left-hand plot of Figure~\ref{fig:sigmaM} shows the distribution of the uncertainty $\sigma_I$ derived from each tracer's LSST data after marginalizing over its initial state vector.
+This histogram marginalizes over the apparent magnitude and observing conditions of the tracer, the date of the encounter, and the direction $\bhat$ of the impulse.
+The right-hand plot of Figure~\ref{fig:sigmaI} forecasts the distribution of $\sigma_M$ that we can expect for deflectors.
+This forecast is made as follows: for each of the 5000 deflectors that we have selected from each MBA region, we have a list of all of its encounters with known MBAs with $b<0.03$~AU.
+For each such encounter, we select at random one $\sigma_I$ from the distribution found by \citet{negin}, and calculate $\sigma_M=(bv/2)\sigma_I.$  For each deflector, we then plot the histogram of (a) the individual encounter with lowest $\sigma_M$; (b) the combined constraint $\left(\sum \sigma_{M,i}^2\right)^{-1/2}$ from the 10 lowest $\sigma_M$ encounters for that deflector; and (c) the combined constraint for all $b<0.03$~AU encounters (typically $\sim1000$).
+The RMS $\sigma_M$ using LSST data from the 10 most-informative tracer encounters with a given deflector is estimated as $\sigma_M=2.6\times10^{-14}M_\odot.$ There are $\approx450$ MBAs with mass above this forecasted RMS $\sigma_M.$  This is a conservative estimate in that we ignore any observational information before LSST, and we also do not consider information from the $\approx10\times$ more MBAs that LSST will discover.
 
 \subsubsection{Individually detectable impulses}
-\citet{negin} forecast the uncertainty $\sigma_v$ on an impulse 
-applied to each of $\approx7000$ tracers using astrometric data expected from LSST
-over its 10-year duration.  The value of $\sigma_I$ depends upon the
-apparent magnitude of the tracer, the timing and geometry of the
-impulse, and the stochastic variations in the observing pattern and
-weather for LSST.
-For impulses occurring near the midpoint
-of the survey, $\approx80\%$ of all cases yield $\sigma_I<10^{-4}\,{\rm m}\,{\rm
+Examining the $\sigma_I$ distribution forecasted for LSST data on tracers (Figure~\ref{fig:sigmaI}, we find that for impulses occurring near the midpoint of the survey, $\approx80\%$ of all cases yield $\sigma_I<10^{-4}\,{\rm m}\,{\rm
   s}^{-1},$ some an order of magnitude lower.  We conservatively
-assume that $I_{\rm det}=5\times 10^{-4}\,{\rm m}\,{\rm
-  s}^{-1},$ which when normalized by $v_c$ yields $I_{\rm
-  det}=10^{-8}$ at 10\% accuracy across the main belt.
+assume that any encounter producing  an impulse $>5\times$ this value, $I_{\rm det}=5\times 10^{-4}\,{\rm m}\,{\rm
+  s}^{-1},$ can be used to directly constrain the relevant deflector and include it in the ephemeris model.
+When normalized by $v_c$ yields $I_{\rm  det}=10^{-8}$ at 10\% accuracy across the main belt.
 
@@
 RMS range and vertical ephemeris errors would be closer to 100 and 20
-meters, respectively.  In the next decade, LSST will significantly
-increase the number $N$ and lower $\sigma_M$ of deflector MBAs with
-usefully measured masses from mutual encounters by tracking hugely
-more potential tracers.  We find that, roughly speaking, $\sigma_\phi$
+meters, respectively.  We find that, roughly speaking, $\sigma_\phi$
 drops with the number of modeled MBAs as $1/N,$ and
@@
 \end{equation}
+ In the next decade, LSST will significantly
+increase the number $N$ and lower $\sigma_M$ of deflector MBAs with
+usefully measured masses from mutual encounters by tracking hugely
+more potential tracers.
+The conservative estimate given in Section~\ref{sec:mbaprops} is that LSST will lower $\sigma_M$ to $\approx10^{-13.7},$ which our calculations indicate lowers the $\sigma_\phi$ values to $\approx250$~m or 0.1~mas.
 
@@
 The Brownian motion might become relevant for analysis of a significant fraction of Gaia MBAs during analysis of the complete 11-year data.
+But with LSST estimates of deflector masses in hand, with $\sigma_\phi\approx0.1$~mas, the (retrospective) use of Gaia data will probably have negligible uncertainty from Brownian motion.
 
@@
 Objects with apparent magnitude $m\gtrsim20$ will have larger per-epoch errors due to photon noise in the images.
-There will, however, be hundreds of observations taken of each MBA over 10~yr, meaning that Brownian motion might be a significant contributor to inferences from the dynamics of brighter tracer MBAs.
+There will, however, be hundreds of observations taken of each MBA over 10~yr, meaning that Brownian motion might be a significant contributor to inferences from the dynamics of brighter tracer MBAs, if our $\sigma_M$ levels were fixed near current knowledge.
 
-On the other hand, Gaia has greatly improved the astrometric precision for $\sim 10^5$ MBAs and LSST will do so for $\sim 10^7$ potential tracer MBAs, significantly lowering $\sigma_M$ and increasing the number $N$ and of deflector MBAs with usefully measured masses from mutual encounters.
-If either can successfully measure masses at $S/N\gtrsim 1$ for the most of the $\ge 1000$ largest MBAs, Figure~\ref{fig:result} indicates reduction of the unmodelled Brownian motion to $\approx0.02$~mas, well below the expected measurement errors for both experiments.
-The prospects for constraining individual MBAs' masses with LSST data on mutual encounters are examined by \citet{negin}, who conclude ????.
-
-\textbf{Fraction of objects generating $1\sigma$ detections in Negin project??}
+On the other hand, LSST will greatly improved the astrometric precision for $\sim 10^7$ potential tracer MBAs, significantly lowering $\sigma_M,$ so a self-consistent analysis should consider the per-epoch contribution of unmodelled Brownian motion to grow to $\approx0.1$~mas by the end of the survey.
+Considering the coherence of the Brownian noise over $O(100)$ LSST observations, the Brownian noise could become comparable to measurement noise for inferences using the brighter tracer MBAs in LSST data, but will remain unimportant for the great majority of LSST's measured MBAs.
 
@@
 These are quite distinct from the random walk in $\Delta\tau$ that is the dominant effect of Brownian motion.
-At the other extreme, if one were trying to measure the Yarkovsky effect (which, to our knowledge, has never been detected on an MBA), this will be primarily realized as a time-invariant tangential force (torque) whose amplitude and direction depend on the (unknown) spin axis, the spin rate, and the radiant and thermal properties of the surface, which will lead to a linear increase or decrease in $\Delta\tau$ over the observation period.
-This may be  difficult to discern from the Brownian motion's random walk in $\Delta\tau$ over the same interval.
+The Brownian noise will be somewhat orthogonal to the signal under study in the data space, and therefore cause less degradation to the Planet X inference than the per-epoch $\sigma_\phi$ would suggest.
+Another example would be a gravitational perturbation whose signature is primarily a nodal precession---we have seen that Brownian motion out of the orbital plane is $>40\times$ smaller than the $\phi$ component.
+At the other extreme, if one were trying to measure the Yarkovsky effect (which, to our knowledge, has never been detected on an MBA), this will be primarily realized as a time-invariant tangential force (torque) whose amplitude and direction depend on the (unknown) spin axis, the spin rate, and the radiant and thermal properties of the surface.
+This will lead to a linear increase or decrease in $\Delta\tau$ over the observation period.
+This signal may be  difficult to discern from the Brownian motion's random walk in $\Delta\tau$ over the same interval.
 The ratio of the Yarkovsky signal to the Brownian motion will increase as $\sqrt{T}$ and will also scale with the inverse diameter of the source.
@@
 This is large enough to become a significant part of the error budget for many inferences from the larger/brighter asteroids being used as tracer particles from Gaia or LSST images.
-If, however, we can use mutual events to constrain the masses of the $\approx1000$ largest  asteroids with $S/N\gtrsim1,$ then the unmodelled Brownian motion shrinks to $\approx100$~m for MBAs, which becomes insignificantly small for most uses of Gaia or LSST astrometry.
+If, however, we can use mutual events from LSST to constrain the masses of the $\approx500$ largest  asteroids with $S/N\gtrsim1,$ then the unmodelled Brownian motion shrinks to $\approx250$~m (or 0.1~mas) for MBAs, which becomes insignificantly small for most uses of Gaia or LSST astrometry.
 Occultation-based astrometry might, at this stage, have measurement errors that are similar to the uncertainties from Brownian motion on decade time scales.

### 2025-02-22 main.tex
Yarkovsky text

@@
 
-One of our tasks will be to derive, from the known asteroid population, the rate (per target) of encounters vs the imparted impulse,
+One of our tasks will be to derive, from the known asteroid population, the rate (per tracer) of encounters vs the imparted impulse,
 \begin{equation}
@@
 approximating it as constant within each of three subsets of tracer MBAs:
-the inner, middle, and outer regions bounded by the 4:1, 3:1,
+the Inner, Middle, and Outer regions bounded by the 4:1, 3:1,
 5:2, and 2:1 mean-motion resonances with Jupiter.
@@
 the range, plus the celestial latitude and longitude of the target.
-We will simplify our results by assuming a heliocentric observer, so that the observational position vector is $\vecp\equiv (r,\theta=z/r,\phi).$ The quantity we seek is the covariance matrix $\covm^p$ of the observations attributable to the accumulated gravitational perturbations:
+We will simplify our results by assuming a heliocentric observer, so that the observational position vector is $\vecp\equiv (r,\phi,\theta=z/r).$ The quantity we seek is the covariance matrix $\covm^p$ of the observations attributable to the accumulated gravitational perturbations:
 \begin{equation}
@@
 \begin{equation}
-  \Delta \vecq(T,t_i) = \matA(e_0, T, t_i) \cdot \vecI.
+  \Delta \vecq(e_0,T,t_i) = \matA(e_0, T, t_i) \vecI.
   \label{eq:A}
@@
 at time $T.$  In linear perturbation theory this will again be expressible as a matrix
-\begin{equation}
-  \Delta\vecp(e_0,T) = \matB(e_0,T) \cdot \vecq(T).
+\begin{eqnarray}
 \label{eq:B}
-\end{equation}
-Combining this with \eqq{eq:Cp} and \eqq{eq:Cqjk} yields the desired result
-\begin{eqnarray}
-  \covm^p_{\alpha\beta}(e_0,T) & = & \sum_{jk} B_{\alpha j}(e_0,T) B_{\beta k}(e_0,T) \covm^q_{jk}(T) \nonumber \\
-  & = & \sum_\gamma \left\langle nI_\gamma^2\right\rangle \int dt_i \left[AB(e_0,T,t_i)\right]_{\alpha\gamma}  \left[AB(e_0,T,t_i)\right]_{\beta\gamma} 
-\label{eq:ABq}
+  \Delta\vecp(e_0,T) & = & \matB(e_0,T) \vecq(T) \\
+  \Rightarrow \quad 
+  \covm^p_{\alpha\beta}(e_0,T) & = & \sum_{jk} B_{\alpha j}(e_0,T) B_{\beta k}(e_0,T) \covm^q_{jk}(T) 
+\label{eq:covp}
 \end{eqnarray}
-which we would wish to average over the phase of the initial mean anomaly $M_0$ of the MBA.
+which we would wish to average over the phase of the MBA orbit or mean anomaly $M$ at the time of observation $T.$
 
@@
         \int_{b_{\rm min}}^{b_{\rm max}} 2\pi b\,db\,
-        \left(\frac{\sigma_M(H)}{bv}\right)^2 \\
-  & = & 2\pi \int dH \frac{dN}{dH} \sigma^2_M(H)  \int dv
+        \left(\frac{2\sigma_M(H)}{bv}\right)^2 \\
+  & = & 8\pi \int dH \frac{dN}{dH} \sigma^2_M(H)  \int dv
         \frac{dn}{dA\,dv} v^{-2} \left\langle\hat b_\gamma^2(v)\right\rangle
@@
 $I_{\rm det}$ of impulse as being large enough to generate high-$S/N$
-deflections on its tracer, then our condition becomes $M(H)/bv <
-I_{\rm det},$ or $b_{\rm min}=M(H)/vI_{\rm det}.$  We then have
+deflections on its tracer, then our condition becomes $2M(H)/bv <
+I_{\rm det},$ or $b_{\rm min}=2M(H)/vI_{\rm det}.$  We then have
 \begin{equation}
    \left\langle n I_\gamma^2 \right\rangle = 
-2\pi \int dH \frac{dN}{dH} \sigma^2_M(H)  \int_{\sqrt{M(H)/I_{\rm
+2\pi \int dH \frac{dN}{dH} \sigma^2_M(H)  \int_{\sqrt{2M(H)/I_{\rm
       det}}}^\infty \frac{dv}{v^2} 
         \frac{dn}{dA\,dv}  \left\langle\hat b_\gamma^2(v)\right\rangle
-        \log \left[v^2 I_{\rm det}/M(H)\right].
+        \log \left[v^2 I_{\rm det}/2M(H)\right].
         \label{eq:nI2}
@@
 We will estimate the constituents of  $\langle nI_\gamma^2\rangle$ in
-\eqq{eq:nI2} separately for each of the nine combinations of tracers
-in Inner, Middle, and Outer regions with deflectors from each region.
+\eqq{eq:nI2} separately for each of the nine combinations of (tracer region, deflector region).
 
@@
 2-day intervals until we have recorded 10 years' worth of encounters---17.7 million
-total impulses.
+total impulses, or roughly 1000 encounters per tracer.
 
@@
 geometry of the encounter, $\langle \hat b^2_\gamma \rangle,$ as a
-function of $v$.  Figure~\ref{fig:rtz} plots these functions of $v$
-for the Inner-Inner, Middle-Middle, and Outer-Outer scattering
-events.  Here it is apparent that the $\hat b_\phi$ component is
+function of $v$, all plotted in Figure~\ref{fig:rtz}.
+Here it is apparent that the $\hat b_\phi$ component is
 larger than $\hat b_z$ and $\hat b_r,$ but there is significant
@@
 mas-level astrometry.
-\citet{negin} use a simulated LSST sequence of observations to estimate the impulse (and mass) uncertainties from mutual encounters of the known asteroids that will occur during years 2--9 of its 10-year survey.
-The tracer MBAs for those events are essentially an ubiased selection from the known MBAs, and the left-hand plot of Figure~\ref{fig:sigmaM} shows the distribution of the uncertainty $\sigma_I$ derived from each tracer's LSST data after marginalizing over its initial state vector.
+\citet{negin} use a simulated LSST sequence of observations to estimate the impulse (and mass) uncertainties from analyzing LSST's data on mutual encounters of the known asteroids.
+The tracer MBAs for those events are essentially an ubiased selection from the known MBAs, and the left-hand plot of Figure~\ref{fig:sigmaI} shows the distribution of the uncertainty $\sigma_I$ derived from each tracer's LSST data after marginalizing over its initial state vector.
 This histogram marginalizes over the apparent magnitude and observing conditions of the tracer, the date of the encounter, and the direction $\bhat$ of the impulse.
 The right-hand plot of Figure~\ref{fig:sigmaI} forecasts the distribution of $\sigma_M$ that we can expect for deflectors.
-This forecast is made as follows: for each of the 5000 deflectors that we have selected from each MBA region, we have a list of all of its encounters with known MBAs with $b<0.03$~AU.
-For each such encounter, we select at random one $\sigma_I$ from the distribution found by \citet{negin}, and calculate $\sigma_M=(bv/2)\sigma_I.$  For each deflector, we then plot the histogram of (a) the individual encounter with lowest $\sigma_M$; (b) the combined constraint $\left(\sum \sigma_{M,i}^2\right)^{-1/2}$ from the 10 lowest $\sigma_M$ encounters for that deflector; and (c) the combined constraint for all $b<0.03$~AU encounters (typically $\sim1000$).
-The RMS $\sigma_M$ using LSST data from the 10 most-informative tracer encounters with a given deflector is estimated as $\sigma_M=2.6\times10^{-14}M_\odot.$ There are $\approx450$ MBAs with mass above this forecasted RMS $\sigma_M.$  This is a conservative estimate in that we ignore any observational information before LSST, and we also do not consider information from the $\approx10\times$ more MBAs that LSST will discover.
+This forecast is made as follows: we treat each of the 5000 tracer MBAs that we have selected from each region as deflectors, since we have a list of all of their encounters with known MBAs with $b<0.03$~AU.
+For each such encounter, we select at random one $\sigma_I$ from the distribution found by \citet{negin} (Figure~\ref{fig:sigmaI}, left), and calculate $\sigma_M=(bv/2)\sigma_I.$  For each deflector, we then calculate (a) the individual encounter with lowest $\sigma_M$; (b) the combined constraint $\left(\sum \sigma_{M,i}^2\right)^{-1/2}$ from the 10 lowest $\sigma_M$ encounters for that deflector; and (c) the combined constraint for all $b<0.03$~AU encounters (typically $\sim1000$ of them).
+The right-hand plot of Figure~\ref{fig:sigmaI} gives the histograms of these three measures of $\sigma_M.$  We see that most of the available information on any given deflector's mass comes from its 10 most-informative encounters.
+It would be computationally very easy to solve for the masses of $O(10^4)$ MBAs by including the deflections they impart on $O(10^5)$ other MBAs in an ephemeris fit.
+The RMS $\sigma_M$ using LSST data from the 10 most-informative tracer encounters with a given deflector is estimated as $\sigma_M=2.6\times10^{-14}M_\odot.$ There are $\approx450$ MBAs with mass above this forecasted RMS $\sigma_M.$  This is a conservative estimate in that we ignore any observational information before LSST, and we also do not consider the information increase that will come from tracking the $\approx10$-fold increase in number of known MBAs that LSST will enable.
 
 \subsubsection{Individually detectable impulses}
-Examining the $\sigma_I$ distribution forecasted for LSST data on tracers (Figure~\ref{fig:sigmaI}, we find that for impulses occurring near the midpoint of the survey, $\approx80\%$ of all cases yield $\sigma_I<10^{-4}\,{\rm m}\,{\rm
+Examining the $\sigma_I$ distribution forecasted for LSST data on tracers (Figure~\ref{fig:sigmaI}), we find that for impulses occurring near the midpoint of the survey, $\approx80\%$ of all cases yield $\sigma_I<10^{-4}\,{\rm m}\,{\rm
   s}^{-1},$ some an order of magnitude lower.  We conservatively
@@
   s}^{-1},$ can be used to directly constrain the relevant deflector and include it in the ephemeris model.
-When normalized by $v_c$ yields $I_{\rm  det}=10^{-8}$ at 10\% accuracy across the main belt.
+When normalized by $v_c,$ we have $I_{\rm  det}=10^{-8}$ at 10\% accuracy across the main belt.
 
@@
 \end{eqnarray}
-We adopt the changes in direction $(\Delta\Lhat_x, \Delta\Lhat_y)$ as the next two elements of $\vecq$ specifying the inclination and ascending node $\Omega$ of the perturbed orbit.
+We adopt the changes in direction $(\Delta\Lhat_x, \Delta\Lhat_y)$ as two elements of $\vecq$ specifying the inclination and ascending node $\Omega$ of the perturbed orbit.
 These are a rotation of $(\Delta\Lhat_\parallel, \Delta\Lhat_\perp)$ by the angle $\tau.$  The conversion from the first line above to the subsequent two makes use of the decomposition $\phat=\uhat_\perp + 2e_\perp \uhat_\parallel$ derivable from the value of $\phi$ in Equations~(\ref{eq:kepler}).
@@
 We now have all the information we need to create the $\matA$ matrix.
-Rotating the $\vecq$ elements back into the Cartesian frame, we
-obtain, to $O(e)$,
+Rotating to $\vecq=\{\Delta a, \Delta \tau, \Delta e_x, \Delta e_y, \Delta\Lhat_x, \Delta\Lhat_y\}$ and with $\vecI=\{I_r, I_\phi, I_z\}$ we have, to $O(e)$,
 \begin{equation}
-  \left( \begin{array}{c}
-           \Delta a \\ \Delta\tau \\ \Delta e_x \\ \Delta e_y \\ \Delta \Lhat_x \\ \Delta\Lhat_y
-         \end{array}\right) = \left( \begin{array}{ccc}
+         \matA(e,T,t_i) = \left( \begin{array}{ccc}
  2e\sin\tau & 2 + 2e\cos\tau & 0 \\
@@
  \end{array}\right)
-  \left(\begin{array}{c} I_r \\ I_\phi \\ I_z \end{array}\right)
 \end{equation}
-We can now do the time integral and sum over impulse directions in \eqq{eq:Cqjk}, and average over $\tau$ to average over the orbital phase of the impulses. The diagonal elements are
+We can now do the time integral and sum over impulse directions in \eqq{eq:Cqjk}, and average over $\tau$ to average over the orbital phase of the impulses. The diagonal elements of $\covm^q$ are
 \begin{eqnarray}
@@
 \end{eqnarray}
-All of the off-diagonal terms (covariances) are either zero or $O(e)\left\langle nI^2\right\rangle,$ and they result in observable consequences that are smaller by $\left\langle e^2 \right\rangle$ than the diagonal terms' contributions.
+All of the off-diagonal terms (covariances) are either zero or $O\left(e\left\langle nI^2\right\rangle\right),$ and they result in observable consequences that are smaller by $\left\langle e^2 \right\rangle$ than the diagonal terms' contributions.
 
@@
 We retain the final term because, as $T>>1,$ this precession term that scales with $\langle e^2 \rangle T^3$ becomes comparable to or larger than the other terms that scale as $T$.
-The $T$-scaling terms all arise from epicyclic motions that are periodic with the orbit.
-
 The heliocentric azimuthal angle $\phi$ is given by \eqq{eq:phi}.  We can similarly square the perturbation and average over orbital phase and population factors to get
@@
 \item There are no significant covariances among the $r,\theta,$ and $\phi$ displacements at the population-averaged level.
-  \item For Earth-based observers, the deviations in $(r,\theta,\phi)$ from the nominal orbit will be some calculable, time-dependent, nearly-orthogonal transformation of the heliocentric deviations derived above, so the results will not be grossly different aside from bringing the three directional components closer together in amplitude.
+  \item For Earth-based observers, the deviations in $(r,\theta,\phi)$ from the nominal orbit will be some calculable, time-dependent, nearly-orthonormal matrix transformation of the heliocentric deviations derived above, so the results will not be grossly different aside from bringing the three directional components closer together in amplitude.
 \end{itemize}
@@
     the (lower) $x$ axis, \ie\ the cumulative noise from all MBAs
-    smaller than the plotted $H_{\rm min.}$  The upper axis gives mass
-    instead of $H$.  The left axis gives values in meters, the right
+    fainter (smaller) than the plotted $H_{\rm min.}$  The upper axis gives the $M_{\rm max}$ equivalent to $H_{\rm min}.$
+.  The left axis gives values in meters, the right
     the apparent shift in milliarseconds for a heliocentric observer.
@@
     is the number of MBAs with $M>\sigma_M$ that could be usefully
-    modelled.}
+    modelled.  The black dashed curve is a conservative estimate of what can be achieved using LSST data to constrain larger MBA's masses.}
   \label{fig:result}
@@
 
-Figure~\ref{fig:result} plots the resulting noise level $\sigma_\phi=\sqrt{\Var{\phi}}$ of the calculation for the Middle regions of MBAs, showing both the linear RMS displacements on the left axis and the angular displacements on the right axis.
+Figure~\ref{fig:result} plots the resulting noise level $\sigma_\phi=\sqrt{\Var(\phi)}$ of the calculation for the Middle regions of MBAs, showing both the linear RMS displacements on the left axis and the angular displacements on the right axis.
 Each point on the curve is the $\sigma_\phi$ attributable to deflector MBAs that are smaller (fainter) than the $H$ value on the lower axis, or the mass on the top axis.
@@
 \item We plot the azimuthal noise.
-\emph{The radial noise is $\approx 7\times$ lower than the azimuthal noise, and the vertical Brownian motion is $\approx45\times$ lower than the azimuthal, on average.}
+\emph{The radial noise results look almost identical, but $\approx 7\times$ lower RMS, and likewise the vertical Brownian motion is  very similar but scaled $\approx45\times$ lower than the azimuthal RMS.}
 \item For longer time periods $T$, the azimuthal noise will scale as
-$T^{3/2},$ the radial noise slightly below this rate, and the vertical
-noise as $T^{1/2}.$
+$T^{3/2},$ the vertical
+noise as $T^{1/2},$ and the radial noise between these.
 \item The largest contributions to the unmodelled Brownian motion are
@@
 encounters.  Our calculations suggest then that a 10-year ephemeris
-including these 100 MBAs will accrue RMS errors slightly under 1~km in
-the azimuthal position, or 0.4~mas in observed ecliptic longitude.
-RMS range and vertical ephemeris errors would be closer to 100 and 20
-meters, respectively.  We find that, roughly speaking, $\sigma_\phi$
+including these 100 MBAs will accrue RMS errors of $\approx 2$~km in
+the azimuthal position, or 1~mas in observed ecliptic longitude.
+RMS range and vertical ephemeris errors would be closer to 300 and 40~m, respectively.  We find that, roughly speaking, $\sigma_\phi$
 drops with the number of modeled MBAs as $1/N,$ and
-\begin{equation} \sigma_\phi \equiv \sqrt{\Var{\phi}} \approx
-(100\,\textrm{m}) \times \left(\sigma_M /
+\begin{equation} \sigma_\phi \equiv \sqrt{\Var(\phi}) \approx
+(200\,\textrm{m}) \times \left(\sigma_M /
 10^{-14}M_\odot\right)^{0.6}.
@@
 more potential tracers.
-The conservative estimate given in Section~\ref{sec:mbaprops} is that LSST will lower $\sigma_M$ to $\approx10^{-13.7},$ which our calculations indicate lowers the $\sigma_\phi$ values to $\approx250$~m or 0.1~mas.
+The conservative estimate given in Section~\ref{sec:mbaprops} is that LSST will lower $\sigma_M$ to $\approx10^{-13.7},$ which our calculations indicate lowers the $\sigma_\phi$ values to $\approx500$~m or 0.25~mas.
 
@@
 $I_\phi=10^{-8}/\sqrt{2}$ and $T=10$~yr, then
-$\Delta\phi\approx35$~mas.  This is nearly $100\times$ larger than the
-stochastic $\sigma_\phi$ expected from unmodelled masses at the
+$\Delta\phi\approx35$~mas.  This is much larger than the
+stochastic $\sigma_\phi\approx1$~mas expected from unmodelled masses at the
 current state of knowledge.  For primarily radial or vertical
@@
 T^{-1/2},$ this signal-to-noise ratio would still be $\gg 5$ for
-surveys of several decades.  We conclude that our choice of $I_{\rm
+surveys of several decades' duration.  We conclude that our choice of $I_{\rm
 det}$ is not ruined by the presence of Brownian motion noise on the
@@
 Thus the Brownian motion of $\approx0.4$~mas we estimate for $\phi$ (over a 10-year period) would be important for the single-epoch error budget of only the brightest Gaia MBAs.
-The Brownian motion is, however, highly correlated across epochs, whereas measurement errors are independent and decrease as the square root of the number of epochs, which is $\approx20$ over 3 years for a typical MBA in the Gaia DR3 catalog.
+The Brownian motion is, however, highly correlated across epochs, whereas each epoch's measurement errors are independent.
+The Brownian motion noise hence does not average down as the square root of the number of epochs, which is $\approx20$ over 3 years for a typical MBA in the Gaia DR3 catalog.
 The Brownian motion might become relevant for analysis of a significant fraction of Gaia MBAs during analysis of the complete 11-year data.
-But with LSST estimates of deflector masses in hand, with $\sigma_\phi\approx0.1$~mas, the (retrospective) use of Gaia data will probably have negligible uncertainty from Brownian motion.
+But with LSST estimates of deflector masses in hand c.\ 2035, with $\sigma_\phi\approx0.25$~mas, the (retrospective) use of Gaia data will probably again will have Brownian motion subdominant to measurement errors the fainter few magnitudes of its catalog.
 
@@
 Objects with apparent magnitude $m\gtrsim20$ will have larger per-epoch errors due to photon noise in the images.
-There will, however, be hundreds of observations taken of each MBA over 10~yr, meaning that Brownian motion might be a significant contributor to inferences from the dynamics of brighter tracer MBAs, if our $\sigma_M$ levels were fixed near current knowledge.
-
-On the other hand, LSST will greatly improved the astrometric precision for $\sim 10^7$ potential tracer MBAs, significantly lowering $\sigma_M,$ so a self-consistent analysis should consider the per-epoch contribution of unmodelled Brownian motion to grow to $\approx0.1$~mas by the end of the survey.
+There will, however, be hundreds of observations taken of each MBA over 10~yr, meaning that Brownian motion might be a significant contributor to inferences from the dynamics of most of its tracer MBAs, if our $\sigma_M$ levels were fixed near current knowledge.
+On the other hand, analyses of LSST astrometry of MBAs should certainly take advantage of its ability to improve knowledge of the masses of individual larger MBAs, and lower the ephemeris errors to $\approx0.25$~mas, well below single-epoch atmospheric turbulence noise..
 Considering the coherence of the Brownian noise over $O(100)$ LSST observations, the Brownian noise could become comparable to measurement noise for inferences using the brighter tracer MBAs in LSST data, but will remain unimportant for the great majority of LSST's measured MBAs.
 
-An even more ambitious proposal by \citet{occult} is to track $\approx10^6$ MBAs down to diameters of $\approx1$~km by watching them occult Gaia stars.
+An even more ambitious proposal by \citet{occultations} is to track $\approx10^6$ MBAs down to diameters of $\approx1$~km by watching them occult Gaia stars.
 The timing of each occultation can locate the center of mass of the target MBA along its direction of motion, with most of the information available from objects yielding uncertainties of 50--100~m on this position.
@@
 
-The extent to which the Brownian motion can confound an inference also depends on how similar the spatio-temporal dependence of the Brownian motion is to the signal under test.
+The algorithms used to make inferences based on range or astrometric data will need to be changed when Brownian motion becomes a significant source of error, to accomodate the fact that positional errors on a given tracer are highly correlated between epochs.
+The full covariance matrix of the tracer's observations will be needed rather than typical practice of considering measurements as independent.
+A small modification to our $\matB$ matrices would be needed to calculate the covariance of the Brownian noise between observations at different times $T$ and $T^\prime.$  The Brownian motion should be fairly uncorrelated between different MBAs, because the varying encounter geometries randomize the influence of any particular deflector's mass error on the trajectories of the tracers it influences.
+
+The extent to which the Brownian motion can confound an inference will depend on how similar the spatio-temporal dependence of the Brownian motion is (as quantified by its covariance matrix) to the signal under test.
 If, for example, one is searching for the tidal force of a Planet X, this will be manifested primarily as a collective apsidal precession and a quadrupolar deviation of the orbit from its ellipse that does not evolve with time.
@@
 The Brownian noise will be somewhat orthogonal to the signal under study in the data space, and therefore cause less degradation to the Planet X inference than the per-epoch $\sigma_\phi$ would suggest.
-Another example would be a gravitational perturbation whose signature is primarily a nodal precession---we have seen that Brownian motion out of the orbital plane is $>40\times$ smaller than the $\phi$ component.
-At the other extreme, if one were trying to measure the Yarkovsky effect (which, to our knowledge, has never been detected on an MBA), this will be primarily realized as a time-invariant tangential force (torque) whose amplitude and direction depend on the (unknown) spin axis, the spin rate, and the radiant and thermal properties of the surface.
-This will lead to a linear increase or decrease in $\Delta\tau$ over the observation period.
-This signal may be  difficult to discern from the Brownian motion's random walk in $\Delta\tau$ over the same interval.
-The ratio of the Yarkovsky signal to the Brownian motion will increase as $\sqrt{T}$ and will also scale with the inverse diameter of the source.
-There will be some nominal diameter below which the unknown Yarkovsky signal leads to larger observable $\sigma_\phi$ than the $\sigma_\phi$ we have derived from Brownian motion.
-This will be good news for those wishing to measure individual bodies' Yarkovsky effect, and bad news for those wishing to use these bodies as test particles for other gravitational or non-gravitational forces.
+Another example would be a gravitational perturbation whose signature is primarily a nodal precession---we have seen that Brownian motion out of the orbital plane is $>40\times$ smaller than the $\phi$ component, and a nodal precession's observable effect is primarily in this vertical component.
+
+\subsection{Comparison to radiation forces}
+As we consider smaller MBAs as test particles, radiation pressure becomes larger relative to gravitational forces.
+It is interesting to compare the effects of radiation pressure to those of Brownian motion to see when the former are a bigger source of noise than the latter, or contrarily to see whether Brownian motion interferes with attempts to measure radiation forces on MBAs.
+At first order, radiation pressure from the Sun and the reaction from reflected sunlight cause a constant radial force on the MBA.
+The ratio of this force to the gravitational force for an asteroid of radius $R,$ geometric albedo $q$, and density $\rho$ is
+\begin{equation}
+ \Psi \equiv \frac{F_{\rm rad}}{F_{\rm grav}} = \frac{ (1+q) \pi R^2 L_\odot / 4\pi c}{4\pi GM_\odot R^3\rho/3}
+   = \frac{3(1+q)}{16\pi} \frac{L_\odot}{GM_\odot c R \rho} \approx 3\times10^{-10} 10^{0.2(H-16)},
+ \end{equation}
+ where the last expression adopts our standard values $q=0.25, \rho=2500\,{\rm kg}\, {\rm m}^{-3}.$ The observational consequence of this constant radial force is an $H$-dependent shift $\Delta r = -\Psi a$ at a fixed period of rotation, which does not resemble any expected effect of Brownian motion.
+More important is the Yarkovsky effect, in which anisotropic re-emission of the absorbed radiation leads to a net force that is fraction $g_Y$ of the incident radiation pressure.
+Taking this force to be in the prograde azimuthal direction, and using our system of units with $a=1$ and $v_c=1,$ this leads to a continuous power $dE/dt=g_Y\Psi$ which in turn implies $da/dt=2g_Y\Psi.$   The mean motion rate is hence $1-3 g_Y\Psi t$ and the accumulated shift in mean anomaly over time $T$ amounts to $\Delta\phi = -3 g_Y \Psi  T^2/2.$  The actual value of the azimuthal Yarkovsky force (torque) depends on the spin axis, the spin rate, and the radiant and thermal properties of the surface, so neither the sign nor amplitude are known for any individual MBA.
+Indeed the Yarkovsky effect has not, to our knowledge, been detected on any MBA, but measurements on NEAs suggest that values of $g_Y\lessim0.1$ might be appropriate.
+In this case, a rough comparison of the accumulated $\Delta\phi$ from the Yarkovsky effect to the $\sigma_\phi$ of noise from unmodelled Brownian motion, after a time interval $T$, is
+ \begin{equation}
+   \frac{\Delta\phi}{\sigma_\phi} \approx \frac{ 3 g_Y \Psi T^2/2}{\sqrt{3 T^3 \nIphi}} \approx \sqrt{T/10\,{\rm yr}} 10^{^{0.2(H-20.2)}
+     \label{eq:yarkovsky}
+   \end{equation}
+   if we assume $\nIphi$ from the current knowledge of asteroid masses.
+There is thus a break point near $H=20$, or $\approx100$~m diameter: for MBAs larger than this, the ephemeris uncertainties from the Yarkovsky effect will be smaller than those from mutual gravitational encounters.
+This includes all MBAs detected by Gaia and nearly all detectable by LSST.
+Below this size, the Yarkovsky displacement grows larger in a decade and has a better chance of being detectable in the presence of Brownian motion.
+Distinguishing the steady quadratic growth of the Yarkovsky $\Delta\phi$ from the random-walk growth of Brownian motion to the same final size requires observational S/N and time resolution significantly better than is needed for detecting either one individually.
+
+\subsection{Summary}
 

### 2025-02-23 main.tex
Completed draft with abstract.

@@
 \begin{abstract}
-  xxx
+  The explosion of high-precision astrometric data on main-belt asteroids (MBAs) opens the door to their use for many possible inferences on gravitational and non-gravitational forces present in this region.
+We estimate the size of positional shifts on MBAs caused by mutual gravitational encounters with other MBAs that are either omitted from ephemeris models or have uncertain mass estimates.
+In other words, what is the typical Brownian motion among MBAs that cannot be predicted from the ephemeris, and therefore serves as noise on inferences from the MBAs?
+We calculate the RMS azimuthal shift $\sigma_\phi$ of this ``Brownian noise''  by numerical estimation of the distribution of impulse sizes and directions among known MBAs, combined with analytical propagation into future positional uncertainties.
+The result depends critically on how well the masses of the larger MBAs are known.
+At current level of asteroid-mass knowledge, $\sigma_\phi$ rises to $\approx2$~km or $\approx1$~mas over $T=10$~yr, increasing as $T^{3/2}.$  This is large enough to degrade many inferences from Gaia and LSST MBA data, although the LSST data will also improve MBA mass knowledge enough to lower this Brownian uncertainty by $\ge4\times.$  Radial and vertical Brownian noise  at $T=10$~yr are suppressed by factors $\approx7$ and $\approx45,$ respectively, compared to the azimuthal signal after 10~yr, and grow more slowly with time.
+We suggest that for full exploitation of Gaia and LSST MBA data, ephemeris models should include the $\approx1000$~largest asteroids as active bodies with free masses, even if not all are well constrained.
+This will correctly propagate the uncertainties from mutual encounters into desired inferences, with the RMS value of positional errors from encounters with less-massive MBAs being limited to $\sigma_\phi\approx60$~m or 30~$\mu$as, small enough to ignore until occultation-based position data become ubiquitously available for MBAs.
 \end{abstract}
@@
 migration, \eg\ the maintainence of the near-Earth asteroid (NEA)
-population \citep{yorp}.  While most applications of precision orbit determination
+population \citep[reviewed by][]{yarkovsky}.  While most applications of precision orbit determination
 have used the 8 major planets as test particles, there are now $>10^6$
@@
 survey projects that can produce
-milliarcsecond-scale accuracy for magnitude-limited small-body populations, such as Gaia \citep{gaiass3},
+milliarcsecond-scale accuracy for magnitude-limited small-body populations, such as Gaia \citep{gaiass3,gaiafpr},
 PanSTARRS,\footnote{\url{https://www2.ifa.hawaii.edu/research/Pan-STARRS.shtml}}
@@
 gravitational forces emanating from the Sun, from the major planets and their large
-moons, from up to $\approx300$ of the largest individual MBAs, and
+moons, from up to $\approx300$ of the largest individual MBAs (typically those that can influence the exquisitely measured ranging to Mars), and
 from distributed mass annuli representing the smaller members of
@@
   \includegraphics[width=0.6\textwidth]{result.pdf}
-  \caption{Each line shows the forecast of the variance of the
+  \caption{Each line shows the forecast of the standard deviation of the
     azimuthal position of a Middle Main Belt asteroid caused by discrete
-    encounters with absolute magnitude $H$ greater than the value on
+    encounters over 10~years with absolute magnitude $H$ greater than the value on
     the (lower) $x$ axis, \ie\ the cumulative noise from all MBAs
     fainter (smaller) than the plotted $H_{\rm min.}$  The upper axis gives the $M_{\rm max}$ equivalent to $H_{\rm min}.$
-.  The left axis gives values in meters, the right
+.  The left axis gives RMS values in meters, the right
     the apparent shift in milliarseconds for a heliocentric observer.
-    Each line is for a different assumption about which asteroids' deflections are
-    part of the ephemeris model instead of being considered as noise: the $\sigma_M$
-    value gives the RMS mass uncertainty on modelled MBAs, and the $N$
-    is the number of MBAs with $M>\sigma_M$ that could be usefully
-    modelled.  The black dashed curve is a conservative estimate of what can be achieved using LSST data to constrain larger MBA's masses.}
+    Each line is for a different assumption about the RMS uncertainty $\sigma_M$ in the masses of individual asteroids gained from high-$S/N$ observations of mutual encounters.
+Each star marks the dividing point between the $N$ MBAs with $M>\sigma_M,$ and the smaller bodies with no useful measure of their mass ($M<\sigma_M$).
+The light dotted lines mark the level of azimuthal noise expected from the bodies beyond the most massive $N,$ \ie\ the RMS that an ephemeris model including $N$ bodies would not properly model.
+    The black dashed curve is a conservative estimate of what can be achieved using LSST data to constrain larger MBA's masses.
+Radial and vertical RMS positions are factors of $\approx 7$ and 45 lower than the azimuthal noise, respectively.}
   \label{fig:result}
@@
 
-\subsection{Implications for future investigations}
+\section{Consequences}
+\subsection{Implications for future inferences}
 
@@
 If the Brownian noise is below other sources of measurement noise, then the answer must be ``no.''
-The per-epoch uncertainties on solar-system object astrometry in Gaia DR3\footnote{We take one Gaia transit as an ``epoch'' here.} are reported to be as low as 0.3~mas for very bright sources ($G=13$), rising to $\approx10$~mas at its $G=21$ limit where most detections lie \citep[][Figure 6]{gaiass3}.
+The per-epoch uncertainties on solar-system object astrometry from Gaia\footnote{We take one Gaia transit as an ``epoch'' here.} are reported to be as low as 0.3~mas for very bright sources ($G=13$), rising to $\approx10$~mas at its $G=21$ limit where most detections lie \citep[][Figure 6]{gaiass3}.
 The overall RMS residual to the orbit fits are 5~mas (this is in the ``along-track'' direction).
-Thus the Brownian motion of $\approx0.4$~mas we estimate for $\phi$ (over a 10-year period) would be important for the single-epoch error budget of only the brightest Gaia MBAs.
-The Brownian motion is, however, highly correlated across epochs, whereas each epoch's measurement errors are independent.
-The Brownian motion noise hence does not average down as the square root of the number of epochs, which is $\approx20$ over 3 years for a typical MBA in the Gaia DR3 catalog.
-The Brownian motion might become relevant for analysis of a significant fraction of Gaia MBAs during analysis of the complete 11-year data.
-But with LSST estimates of deflector masses in hand c.\ 2035, with $\sigma_\phi\approx0.25$~mas, the (retrospective) use of Gaia data will probably again will have Brownian motion subdominant to measurement errors the fainter few magnitudes of its catalog.
+Thus the Brownian motion of $\approx1$~mas we estimate for $\phi$ (over a 10-year period) with current asteroid-mass knowledge would be important for the single-epoch error budget of the brighter Gaia MBAs.
+With our forecasted LSST-era knowledge of asteroid masses ($\sigma_M\approx10^{-13.6},$ RMS azimuthal errors $\approx0.25$~mas), the situation will be improved in that only the brightest few MBAs will have per-epoch Brownian motion noise comparable to the per-epoch measurement noise.
 
@@
 Objects with apparent magnitude $m\gtrsim20$ will have larger per-epoch errors due to photon noise in the images.
-There will, however, be hundreds of observations taken of each MBA over 10~yr, meaning that Brownian motion might be a significant contributor to inferences from the dynamics of most of its tracer MBAs, if our $\sigma_M$ levels were fixed near current knowledge.
-On the other hand, analyses of LSST astrometry of MBAs should certainly take advantage of its ability to improve knowledge of the masses of individual larger MBAs, and lower the ephemeris errors to $\approx0.25$~mas, well below single-epoch atmospheric turbulence noise..
-Considering the coherence of the Brownian noise over $O(100)$ LSST observations, the Brownian noise could become comparable to measurement noise for inferences using the brighter tracer MBAs in LSST data, but will remain unimportant for the great majority of LSST's measured MBAs.
+Assuming that LSST inferences will be able to exploit LSST's own determinations of MBA masses, we see that on a per-epoch basis, the Brownian noise will always be subdominant to measurement errors.
 
-An even more ambitious proposal by \citet{occultations} is to track $\approx10^6$ MBAs down to diameters of $\approx1$~km by watching them occult Gaia stars.
+An ambitious proposal by \citet{occultations} is to track $\approx10^6$ MBAs down to diameters of $\approx1$~km by watching them occult Gaia stars.
 The timing of each occultation can locate the center of mass of the target MBA along its direction of motion, with most of the information available from objects yielding uncertainties of 50--100~m on this position.
-The Brownian motion noise over a 10-year baseline will in this case become comparable to the per-epoch measurement errors.
+The Brownian noise (with LSST-level asteroid knowledge) over a 10-year baseline will in this case be comparable to the per-epoch measurement errors for most observations.
 Reducing the typical mass uncertainty of individual asteroids to $\sigma_M\approx 10^{-15.4} M_\odot$ would be needed ($N=3000$) to reduce the Brownian noise to negligible levels for this survey.
+
+These per-epoch comparisons are, however, misleading.
+The Brownian noise is highly correlated between different epochs of observation of a given asteroid, while the measurement errors are generally uncorrelated, and therefore the impact on inferences is not described solely by comparison of RMS values of each.
+The covariance structure of Brownian noise can work either to the benefit or the detriment of the inference.
+In an unfavorable case, the eigenvectors of this covariance matrix with the largest uncertainties look similar to the signal of interest in the inference, \ie\ the derivative of the MBA's position with respect to a parameter.
+If there is a single dominant eigenvector of the Brownian noise which looks just like our signal, then the effect per-epoch Brownian noise does not average down at all as we obtain $M$ observations, but the measurement noise drops as $\sqrt{M}.$  In a favorable case, all of the strong eigenvectors of the Brownian noise covariance are orthogonal to the signals of interest, and the Brownian noise causes no degradation of the inference, \ie\ its spatiotemporal structure cannot mimic that of the signal we seek.
+
+Consider first the worst-case scenario, where the Brownian noise does not average down at all across observations, but measurement errors gain a factor $\sqrt{M}.$   \citet{gaiafpr} report typically $\sim60$ epochs per asteroid in 66 months, suggesting $O(100)$ epochs per asteroid in the full 11-year observing period.
+At Gaia's faint end, this means that Brownian noise with present asteroid knowledge would have comparable impact to the measurement errors, and dominate inferences at the bright end.
+LSST will obtain several hundred epochs per asteroid.
+With an effective factor 10--20 reduction in the total measurement noise, and assuming $\sigma_\phi\approx0.25$~mas Brownian noise, the latter would become important for the brighter tracer MBAs in LSST data, but inferences with the millions of fainter objects discovered by LSST would still be limited by photon noise.
+The occultation program proposed by \citet{occultations} would obtain $<10$ measurements per decade for most MBAs and hence see less degradation of Brownian relative to measurement noise.
+
+At the other extreme are cases where the dominant modes of Brownian noise are orthogonal to the effects of parameters of interest.  
+If, for example, one is searching for the tidal force of a Planet X, this will be manifested primarily as a collective apsidal precession and a quadrupolar deviation of the orbit from its ellipse that does not evolve with time.
+These are quite distinct from the random walk in $\Delta\tau$ that is the dominant effect of Brownian motion, and therefore Brownian noise with cause less degradation to the Planet X inference than the per-epoch $\sigma_\phi$ would suggest.
+Another example would be a gravitational perturbation whose signature is primarily a nodal precession, which is a signal manifesting primarily as out-of-plane motions---we have seen that Brownian motion out of the orbital plane is $>40\times$ smaller than the $\phi$ component and grows at $T{-1/2}$ instead of $T^{3/2},$ so the per-epoch $\sigma_\phi$ of Brownian noise will grossly overestimate its effect on the inference.
 
@@
 
-The extent to which the Brownian motion can confound an inference will depend on how similar the spatio-temporal dependence of the Brownian motion is (as quantified by its covariance matrix) to the signal under test.
-If, for example, one is searching for the tidal force of a Planet X, this will be manifested primarily as a collective apsidal precession and a quadrupolar deviation of the orbit from its ellipse that does not evolve with time.
-These are quite distinct from the random walk in $\Delta\tau$ that is the dominant effect of Brownian motion.
-The Brownian noise will be somewhat orthogonal to the signal under study in the data space, and therefore cause less degradation to the Planet X inference than the per-epoch $\sigma_\phi$ would suggest.
-Another example would be a gravitational perturbation whose signature is primarily a nodal precession---we have seen that Brownian motion out of the orbital plane is $>40\times$ smaller than the $\phi$ component, and a nodal precession's observable effect is primarily in this vertical component.
+A better approach, however, would probably be to increase the number $N$ of MBAs considered as active bodies in the ephemeris fitting, leaving their masses as free parameters.
+Although the extreme case of having $N=O(10^6)$ active bodies would be a substantial computational challenge, having $N=10^3$--$10^4$ is not, especially since the MBA-to-MBA perturbations can be adequately calculated using Newtonian dynamics instead of a fully relativistic treatment.
+In this approach, all of the covariances between different observations of the same tracer, and between tracers, will be captured in the analysis for deflections due to the $N$ deflectors.
+This is true even if the ephemeris fit does not yield $S/N>1$ on every deflector.
+If we \emph{choose $N$ to be large enough that the Brownian noise from the unmodelled MBAs is negligible to the error budget,} then we need not explicitly propagate the Brownian noise.
+
+Figure~\ref{fig:result} shows us what $N$ would be required: the noise due to unmodelled deflectors is given by the point where each curve for a given $N$ departs from the upper-right envelope of the curves.
+Increasing the number of MBAs in the ephemeris from $N=100\rightarrow10^3\rightarrow10^4$ decreases the RMS displacement in 10~yrs due to the unmodelled MBAs from $\sigma_\phi=1300\rightarrow60\rightarrow10$~m, with the radial and vertical RMS being substantially lower.
+These correspond to 0.7, 0.03, and 0.005~mas, respectively.
+Incorporating $N=1000$ MBAs into the ephemeris model would render noise from the rest of the asteroids unimportant for nearly all uses of Gaia or LSST MBA data for inference.
+For occultation-based positions, one would probably want to have the $\approx3000$ most massive MBAs as active bodies with masses as free parameters in the ephemeris model in order to render encounters from unmodelled bodies as unimportant.
 
@@
 At first order, radiation pressure from the Sun and the reaction from reflected sunlight cause a constant radial force on the MBA.
-The ratio of this force to the gravitational force for an asteroid of radius $R,$ geometric albedo $q$, and density $\rho$ is
+The ratio of this force to the gravitational force for an asteroid of radius $R,$ geometric albedo $q$, and density $\rho$ is $(1+q)\Psi,$ where
 \begin{equation}
- \Psi \equiv \frac{F_{\rm rad}}{F_{\rm grav}} = \frac{ (1+q) \pi R^2 L_\odot / 4\pi c}{4\pi GM_\odot R^3\rho/3}
-   = \frac{3(1+q)}{16\pi} \frac{L_\odot}{GM_\odot c R \rho} \approx 3\times10^{-10} 10^{0.2(H-16)},
+ \Psi \equiv \frac{F_{\rm rad}}{F_{\rm grav}} = \frac{\pi R^2 L_\odot/4\pi c}{4\pi GM_\odot R^3\rho/3}
+   = \frac{3}{16\pi} \frac{L_\odot}{GM_\odot c R \rho} \approx 3\times10^{-10} 10^{0.2(H-16)},
 \end{equation}
- where the last expression adopts our standard values $q=0.25, \rho=2500\,{\rm kg}\, {\rm m}^{-3}.$ The observational consequence of this constant radial force is an $H$-dependent shift $\Delta r = -\Psi a$ at a fixed period of rotation, which does not resemble any expected effect of Brownian motion.
-More important is the Yarkovsky effect, in which anisotropic re-emission of the absorbed radiation leads to a net force that is fraction $g_Y$ of the incident radiation pressure.
-Taking this force to be in the prograde azimuthal direction, and using our system of units with $a=1$ and $v_c=1,$ this leads to a continuous power $dE/dt=g_Y\Psi$ which in turn implies $da/dt=2g_Y\Psi.$   The mean motion rate is hence $1-3 g_Y\Psi t$ and the accumulated shift in mean anomaly over time $T$ amounts to $\Delta\phi = -3 g_Y \Psi  T^2/2.$  The actual value of the azimuthal Yarkovsky force (torque) depends on the spin axis, the spin rate, and the radiant and thermal properties of the surface, so neither the sign nor amplitude are known for any individual MBA.
-Indeed the Yarkovsky effect has not, to our knowledge, been detected on any MBA, but measurements on NEAs suggest that values of $g_Y\lessim0.1$ might be appropriate.
-In this case, a rough comparison of the accumulated $\Delta\phi$ from the Yarkovsky effect to the $\sigma_\phi$ of noise from unmodelled Brownian motion, after a time interval $T$, is
+where the last expression adopts our standard values $q=0.25, \rho=2500\,{\rm kg}\, {\rm m}^{-3}.$ The observational consequence of this constant radial force is an $H$-dependent shift $\Delta r = -(1+q)\Psi a$ at a fixed period of rotation, which does not resemble any expected effect of Brownian motion.
+More important is the Yarkovsky effect, in which anisotropic re-emission of the absorbed radiation leads to a net force that is fraction $g_Y(1-q)$ of the incident radiation pressure.
+The Yarkovsky effect has been detected for several NEAs but no individual MBAs, with values $g_Y \approx 0.1$ \citep[Yarkovsky effect and data are review by][]{yarkovsky}.
+This is also consistent with detections of the Yarkovsky effect by examining the $\Delta a$ dependence on $H$ of members of asteroid \emph{families} in the Main Belt \citep[\eg][]{karin}.
+
+ The Yarkovsky effect has in common with Brownian noise that it has the strongest observable effects via changes to $a$ and the mean-motion rates, and that these changes can be of varying sign and amplitude for a given asteroid.
+In the case of the Yarkovsky effect, this variability arises because of the random orientations of spin axes, and differences in spin rates, shapes, and surface properties.
+
+Taking a case where the Yarkovsky force to be in the prograde azimuthal direction, and using our system of units with $a=1$ and $v_c=1,$ this leads to a continuous power $dE/dt\approx g_Y\Psi$ which in turn implies $da/dt\approx2g_Y\Psi.$   The mean motion rate is hence $1-3 g_Y\Psi t$ and the accumulated shift in mean anomaly over time $T$ amounts to $\Delta\phi = -3 g_Y \Psi  T^2/2.$   In this case, a rough comparison of the accumulated $\Delta\phi$ from the Yarkovsky effect to the $\sigma_\phi$ of noise from unmodelled Brownian motion, after a time interval $T$, is
 \begin{equation}
-   \frac{\Delta\phi}{\sigma_\phi} \approx \frac{ 3 g_Y \Psi T^2/2}{\sqrt{3 T^3 \nIphi}} \approx \sqrt{T/10\,{\rm yr}} 10^{^{0.2(H-20.2)}
+   \frac{\Delta\phi}{\sigma_\phi} \approx \frac{ 3 g_Y \Psi T^2/2}{\sqrt{3 T^3 \nIphi}} \approx \sqrt{T/10\,{\rm yr}} 10^{0.2(H-20.2)}
      \label{eq:yarkovsky}
@@
 if we assume $\nIphi$ from the current knowledge of asteroid masses.
-There is thus a break point near $H=20$, or $\approx100$~m diameter: for MBAs larger than this, the ephemeris uncertainties from the Yarkovsky effect will be smaller than those from mutual gravitational encounters.
+There is thus a break point near $H=20$, or $\approx100$~m diameter: for MBAs larger than this, the ephemeris uncertainties from the Yarkovsky effect will be smaller than those from mutual gravitational encounters on decade timescales.
 This includes all MBAs detected by Gaia and nearly all detectable by LSST.
@@
 
+When Yarkovsky forces on MBAs are estimated using the $H$-dependent dispersion of $a$ for asteroid families, the effective ``observation'' time period is typically several Myr.
+Over these long time scales, the temporally coherent Yarkovsky effect can dominate over Brownian motion for much larger bodies, which is why the technique is successful.
+
 \subsection{Summary}
 
-In conclusion, the Brownian motion from unmodelled or mis-modelled mutual impulses between MBAs causes RMS azimuthal displacements at the 1~km scale over 10 years, with current knowledge of the masses of large asteroids.
-This is large enough to become a significant part of the error budget for many inferences from the larger/brighter asteroids being used as tracer particles from Gaia or LSST images.
-If, however, we can use mutual events from LSST to constrain the masses of the $\approx500$ largest  asteroids with $S/N\gtrsim1,$ then the unmodelled Brownian motion shrinks to $\approx250$~m (or 0.1~mas) for MBAs, which becomes insignificantly small for most uses of Gaia or LSST astrometry.
-Occultation-based astrometry might, at this stage, have measurement errors that are similar to the uncertainties from Brownian motion on decade time scales.
+In conclusion, the Brownian motion from unmodelled or mis-modelled mutual impulses between MBAs causes RMS azimuthal displacements of $\approx2$~km (or 1 mas in angular position) over 10 years, with current knowledge of the masses of large asteroids.
+This is large enough to become a significant part of the error budget for many inferences from Gaia MBA astrometry.
+The knowledge on large-MBA masses likely to be gained from LSST tracking of MBAs should reduce this by a factor of $\ge4,$ at which point the Brownian noise will be comparable to Gaia or LSST measurement uncertainties only for the brighter targets of each survey.
+In all cases, the Brownian motion is $\approx7\times$ smaller in the radial direction and $\approx45\times$ in the vertical direction.
+The RMS azimuthal displacement grows with time as $T^{3/2}$, the vertical RMS as $T^{1/2}$, and the radial RMS eventual switches from the latter to the former at some time driven by the orbit eccentricity.
+
+Making inferences from MBA astrometry becomes more complicated when Brownian motion is a significant contributor to the error budget, but a good path forward is to do all model fitting by including the $\approx1000$ largest MBAs as active bodies in the ephemeris model and leaving their masses as free parameters.
+This will constrain the masses of these bodies to the full extent that the data allow, and will propertly propagate any remaining uncertainty in those masses into the other inferences that one is trying to make.
+The noise contributed by the remaining millions of unmodelled accuracy would be just $\approx60$~m RMS in the azimuthal direction (about 30~$\mu$as), and a few $\mu$as or less in the other axes, small enough to ignore entirely.
+Occultation-based astrometry is much more accurate (but has fewer possible measurements per tracer), and would be significantly impacted by Brownian motion until $\gtrsim1000$ MBAs have mass estimates at modest $S/N$ level, although the precise impact will depend on whether the signals one is seeking are parallel or orthogonal to the kind of variations caused by Brownian motion.
+For objects $\gtrsim100$~m in diameter, uncertainties from mutual encounters are larger than uncertainties from the Yarkovsky effect over decade timescales, at present levels of knowledge of asteroid masses.
 

### 2025-02-24 main.tex
Complete v1

@@
 \newcommand{\covm}{C}
-\newcommand{\lop}{\varpi}
+\newcommand{\lop}{\varpi}   % longitude of perihelion
+\newcommand{\ma}{\ell}  % Mean anomaly
+
 \newcommand{\Lhat}{\hat L}
@@
 \begin{abstract}
-  The explosion of high-precision astrometric data on main-belt asteroids (MBAs) opens the door to their use for many possible inferences on gravitational and non-gravitational forces present in this region.
-We estimate the size of positional shifts on MBAs caused by mutual gravitational encounters with other MBAs that are either omitted from ephemeris models or have uncertain mass estimates.
-In other words, what is the typical Brownian motion among MBAs that cannot be predicted from the ephemeris, and therefore serves as noise on inferences from the MBAs?
-We calculate the RMS azimuthal shift $\sigma_\phi$ of this ``Brownian noise''  by numerical estimation of the distribution of impulse sizes and directions among known MBAs, combined with analytical propagation into future positional uncertainties.
-The result depends critically on how well the masses of the larger MBAs are known.
-At current level of asteroid-mass knowledge, $\sigma_\phi$ rises to $\approx2$~km or $\approx1$~mas over $T=10$~yr, increasing as $T^{3/2}.$  This is large enough to degrade many inferences from Gaia and LSST MBA data, although the LSST data will also improve MBA mass knowledge enough to lower this Brownian uncertainty by $\ge4\times.$  Radial and vertical Brownian noise  at $T=10$~yr are suppressed by factors $\approx7$ and $\approx45,$ respectively, compared to the azimuthal signal after 10~yr, and grow more slowly with time.
-We suggest that for full exploitation of Gaia and LSST MBA data, ephemeris models should include the $\approx1000$~largest asteroids as active bodies with free masses, even if not all are well constrained.
-This will correctly propagate the uncertainties from mutual encounters into desired inferences, with the RMS value of positional errors from encounters with less-massive MBAs being limited to $\sigma_\phi\approx60$~m or 30~$\mu$as, small enough to ignore until occultation-based position data become ubiquitously available for MBAs.
+  The explosion of high-precision astrometric data on main-belt
+  asteroids (MBAs) opens the door to their use for many possible
+  inferences on gravitational and non-gravitational forces present in
+  this region.  We estimate the size of positional shifts on MBAs
+  caused by mutual gravitational encounters with other MBAs that are
+  either omitted from ephemeris models or have uncertain mass
+  estimates. In other words, what is the typical Brownian motion among
+  MBAs that cannot be predicted from the ephemeris, and therefore
+  serves as noise on inferences from the MBAs?  We calculate the RMS
+  azimuthal shift $\sigma_\phi$ of this ``Brownian noise''  by
+  numerical estimation of the distribution of impulse sizes and
+  directions among known MBAs, combined with analytical propagation
+  into future positional uncertainties.  The result depends strongly
+  on how well the masses of the larger MBAs are known.  At current
+  level of asteroid-mass knowledge, $\sigma_\phi$ rises to
+  $\approx2$~km or $\approx1$~mas over $T=10$~yr, increasing as
+  $T^{3/2}.$  This is large enough to degrade many inferences from
+  Gaia and LSST MBA data, although the LSST data will also improve MBA
+  mass knowledge enough to lower this Brownian uncertainty by
+  $\ge4\times.$  Radial and vertical Brownian noise  at $T=10$~yr are
+  suppressed by factors $\approx7$ and $\approx45,$ respectively,
+  compared to the azimuthal signal after 10~yr, and grow at $T^{3/2}$
+  and $T^{1/2}.$  We suggest that for full exploitation of Gaia and
+  LSST MBA data, ephemeris models should include the
+  $\approx1000$~largest asteroids as active bodies with free masses,
+  even if not all are well constrained.  This will correctly propagate
+  the uncertainties from these 1000 sources' deflections into desired
+  inferences.
+The RMS value of deflections from less-massive MBAs is limited to $\sigma_\phi\approx60$~m or 30~$\mu$as, small enough to ignore until occultation-based position data become ubiquitously available for MBAs.
 \end{abstract}
@@
 hypothetical ``Planet X'' \citep[\eg][]{holmanP9,inpopP9,trojans,occultations} or passing primordial
-black holes \citep{pbh}, determine
-the masses of known bodies \citep[\eg][]{goffin,baer}, test the laws of physics, and characterize 
+black holes \citep[\eg][]{pbh}, determine
+the masses of known bodies \citep[\eg][]{goffin,baer}, test the laws
+of physics \citep[\eg][]{inpopgraviton}, and characterize 
 the non-gravitational forces that drive many forms of planetary
@@
 by Brownian motion over a time period $T \lesssim100$~yr of human
-observation, in the present dynamical environment.  We will make use
+observation, in the present dynamical environment.  Indeed essentially
+all of the relevant deflector bodies and circumstances of encounters
+are known.  We will make use
 of the currently known population of MBAs, correcting for
@@
 from $H$ to mass; the status of future observations and $\sigma_M$ is
-also unclear.  Hence we can determine our quantities only to leading
+also unclear.  We are hence justified in retaining only the
+contributions to our results that are leading
 order in the MBA orbital eccentricity $e$ of the tracer population.
@@
 unit vectors $\rhat, \phat$ rotate with the target asteroid.  The mean
-anomaly at time $t=0$ is $M_0.$
+anomaly at time $t=0$ is $\ma_0.$
 
-All distances will be given in units of the original semimajor axis $a_0$, and all velocities in units of the circular velocity $\vcirc \equiv \sqrt{G\Msun/a_0}.$  In these units, $G\Msun=1,$ the period of the initial orbit is $2\pi,$ and the (unperturbed) mean anomaly is $M=t+M_0.$
+All distances will be given in units of the original semimajor axis $a_0$, and all velocities in units of the circular velocity $\vcirc \equiv \sqrt{G\Msun/a_0}.$  In these units, $G\Msun=1,$ the period of the initial orbit is $2\pi,$ and the (unperturbed) mean anomaly is $\ma=t+\ma_0.$
 
@@
 where the angle brackets indicate an average over possible
-realizations of the impulse history, and over the mean anomaly $M$ at the time of observation.
-To do so, we will introduce an intermediate set of 6 parameters $\vecq$ describing the deviations of the orbital elements induced by the impulses.
-The $\vecq$ components will be selected to be equal to zero before any impulses are applied, and each responds linearly to $\vecI$ at $|I|\ll 1.$
+realizations of the impulse history, and over the mean anomaly $\ma$ at the time of observation.
+To do so, we will introduce an intermediate set of 6 parameters $\vecq$ describing the orbital elements induced by the impulses.
+The $\vecq$ components will be selected to respond linearly to $\vecI$ at $|I|\ll 1.$
 We will derive the matrix $\matA$ that describes the orbital-element shifts at time $T$ that arise from an impulse at time $t_i$: 
 \begin{equation}
-  \Delta \vecq(e_0,T,t_i) = \matA(e_0, T, t_i) \vecI.
+  \Delta \vecq(e_0,T,t_i) = \matA(e_0, T, t_i) \, \vecI.
   \label{eq:A}
 \end{equation}
-In our first-order perturbation theory, $\vecq(T)$ will be the sum of the $\Delta\vecq_i$ imparted by all impulses applied at times $0<t_i<T.$  Because the impulses are uncorrelated, $\vecq$ will therefore be the result of a random walk.
-The distribution will have a covariance matrix $\covm^q(T) \equiv \left\langle \vecq(T) \vecq^T(T) \right\rangle$ whose elements are
+In our first-order perturbation theory, $\Delta\vecq(T)$ will be the sum of the $\Delta\vecq_i$ imparted by all impulses applied at times $0<t_i<T.$  Because the impulses are uncorrelated, $\vecq$ will therefore be the result of a random walk.
+The distribution will have a covariance matrix $\covm^q(T) \equiv \left\langle \Delta\vecq(T) \Delta\vecq^T(T) \right\rangle$ whose elements are
 \begin{eqnarray}
@@
 The last element of our calculation will be a conversion from the
-element shifts $\vecq$ into the observed displacements $\Delta\vecp$
+element shifts $\Delta\vecq$ into the observed displacements $\Delta\vecp$
 at time $T.$  In linear perturbation theory this will again be expressible as a matrix
@@
   \Rightarrow \quad 
-  \covm^p_{\alpha\beta}(e_0,T) & = & \sum_{jk} B_{\alpha j}(e_0,T) B_{\beta k}(e_0,T) \covm^q_{jk}(T) 
+  \covm^p_{\alpha\beta}(e_0,T) & = & \sum_{jk} B_{\alpha j}(e_0,T)
+                                     B_{\beta k}(e_0,T) \, \covm^q_{jk}(T) 
 \label{eq:covp}
 \end{eqnarray}
-which we would wish to average over the phase of the MBA orbit or mean anomaly $M$ at the time of observation $T.$
+which we would wish to average over the phase of the MBA orbit or mean anomaly $\ma$ at the time of observation $T.$
 
@@
 and $\matB(e_0,T).$  Section~\ref{sec:results} combines these and
-summarizes the results and their implications.
+summarizes the results.  The implications of these results for use of
+MBA observations for high-precision inferences are discussed in Section~\ref{sec:discuss}.
 
@@
 through numerical measurements with the orbits of all known MBAs. The
-$H$ distribution of MBAs is $dN/dH.$ We denote as $\sigma_M(H)$
-the RMS error in the mass of MBAs at $H$ from the value used in the ephemeris model.
-We use $\sigma_M$ rather than $M$ since we are seeking the size of
-deviations from the ephemeris model.  For MBAs that are not included
-in the ephemeris model, $\sigma_M=M(H),$ the full mass of the deflector.
+$H$ distribution of MBAs is $dN/dH.$ We define $\sigma_M(H)$ via
+\begin{equation}
+  \sigma_M(H) \equiv \left\{\begin{array}{cl}
+                              \sigma_M, & M(H) > \sigma_M \\
+                              M(H),        & M(H) \le \sigma_M
+                            \end{array} \right.
+  \label{eq:sigmaMH}
+\end{equation}
+such that it represents the error in the ephemeris due to uncertainty
+in the masses of MBAs included in the ephemeris (first row), and due
+to the entirety of the masses of MBAs omitted from the ephemeris
+(second row).
 
@@
 ephemeris model could include the mass of the deflector asteroid, and
-this mass could be determined from fitting the data of the tracer.
+this mass could be usefully determined from fitting the data of just this
+single tracer.
 Such encounters would thus no longer be considered source of
@@
 observations by
-\citet{negin}. The fitting of the tracer data for the mass of the
-deflector is also a small computational effort.
+\citet{negin}.
 
@@
    \left\langle n I_\gamma^2 \right\rangle = 
-2\pi \int dH \frac{dN}{dH} \sigma^2_M(H)  \int_{\sqrt{2M(H)/I_{\rm
+8\pi \int dH \frac{dN}{dH} \sigma^2_M(H)  \int_{\sqrt{2M(H)/I_{\rm
       det}}}^\infty \frac{dv}{v^2} 
@@
 The pairwise interaction rate $dn/dA\,dv$ and the impulse direction
-distributions $\langle \hat b^2_\gamma \rangle$ are estimated from a
+components $\langle \hat b^2_\gamma \rangle$ are estimated from a
 numerical integration of all of the MBA orbits available
@@
 function of $v$, all plotted in Figure~\ref{fig:rtz}.
-Here it is apparent that the $\hat b_\phi$ component is
+Here it is apparent that the $\hat b_\phi$ component is, on average,
 larger than $\hat b_z$ and $\hat b_r,$ but there is significant
@@
 events by \citet{goffin} reports typical $\sigma_M\approx
-5\times10^{-13}$ for 30 MBAs. But we can expect substantial improvement
+5\times10^{-13}$ for 30 MBAs (although the list is biased toward
+better-measured events). But we can expect substantial improvement
 in the LSST era when a millions of potential tracer asteroids attain
-mas-level astrometry.
-\citet{negin} use a simulated LSST sequence of observations to estimate the impulse (and mass) uncertainties from analyzing LSST's data on mutual encounters of the known asteroids.
-The tracer MBAs for those events are essentially an ubiased selection from the known MBAs, and the left-hand plot of Figure~\ref{fig:sigmaI} shows the distribution of the uncertainty $\sigma_I$ derived from each tracer's LSST data after marginalizing over its initial state vector.
-This histogram marginalizes over the apparent magnitude and observing conditions of the tracer, the date of the encounter, and the direction $\bhat$ of the impulse.
-The right-hand plot of Figure~\ref{fig:sigmaI} forecasts the distribution of $\sigma_M$ that we can expect for deflectors.
-This forecast is made as follows: we treat each of the 5000 tracer MBAs that we have selected from each region as deflectors, since we have a list of all of their encounters with known MBAs with $b<0.03$~AU.
-For each such encounter, we select at random one $\sigma_I$ from the distribution found by \citet{negin} (Figure~\ref{fig:sigmaI}, left), and calculate $\sigma_M=(bv/2)\sigma_I.$  For each deflector, we then calculate (a) the individual encounter with lowest $\sigma_M$; (b) the combined constraint $\left(\sum \sigma_{M,i}^2\right)^{-1/2}$ from the 10 lowest $\sigma_M$ encounters for that deflector; and (c) the combined constraint for all $b<0.03$~AU encounters (typically $\sim1000$ of them).
-The right-hand plot of Figure~\ref{fig:sigmaI} gives the histograms of these three measures of $\sigma_M.$  We see that most of the available information on any given deflector's mass comes from its 10 most-informative encounters.
-It would be computationally very easy to solve for the masses of $O(10^4)$ MBAs by including the deflections they impart on $O(10^5)$ other MBAs in an ephemeris fit.
+mas-level astrometry. \citet{negin} use a simulated LSST sequence of
+observations to estimate the impulse (and mass) uncertainties from
+analyzing LSST's data on mutual encounters of the known asteroids. The
+tracer MBAs for those events are essentially an unbiased selection
+from the known MBAs, and the left-hand plot of Figure~\ref{fig:sigmaI}
+shows the distribution of the uncertainty $\sigma_I$ derived from each
+tracer's LSST data after marginalizing over its initial state vector.
+This histogram marginalizes over the apparent magnitude and observing
+conditions of the tracer, the date of the encounter, and the direction
+$\bhat$ of the impulse.  The right-hand plot of
+Figure~\ref{fig:sigmaI} forecasts the distribution of $\sigma_M$ that
+we can expect for deflectors.  This forecast is made as follows: we
+treat each of the 5000 tracer MBAs that we have selected from each
+region as deflectors, since we have a list of all of their encounters
+with known MBAs with $b<0.03$~AU.  For each such encounter, we select
+at random one $\sigma_I$ from the distribution for LSST data on known
+MBAs forecasted by \citet{negin} (Figure~\ref{fig:sigmaI}, left), and
+calculate $\sigma_M=(bv/2)\sigma_I.$  For each deflector, we then
+calculate (a) the individual encounter with lowest $\sigma_M$; (b) the
+combined constraint $\left(\sum \sigma_{M,i}^2\right)^{-1/2}$ from the
+10 lowest $\sigma_M$ encounters for that deflector; and (c) the
+combined constraint for all $b<0.03$~AU encounters (typically
+$\sim1000$ of them).  The right-hand plot of Figure~\ref{fig:sigmaI}
+gives the histograms of these three measures of $\sigma_M.$  We see
+that most of the available information on any given deflector's mass
+comes from its 10 most-informative encounters.  It should not be a
+computational barrier to solve for the masses of $O(10^4)$ MBAs by including the deflections they impart on $O(10^5)$ other MBAs in an ephemeris fit.
 The RMS $\sigma_M$ using LSST data from the 10 most-informative tracer encounters with a given deflector is estimated as $\sigma_M=2.6\times10^{-14}M_\odot.$ There are $\approx450$ MBAs with mass above this forecasted RMS $\sigma_M.$  This is a conservative estimate in that we ignore any observational information before LSST, and we also do not consider the information increase that will come from tracking the $\approx10$-fold increase in number of known MBAs that LSST will enable.
@@
 \label{sec:elements}
-
-Five of our orbital element perturbations $\vecq$ will be taken from constants of the motion.
-The mean anomaly $M$ is not appropriate as the sixth, time-dependent element of $\vecq$ because the derivative $dM/dI$ can diverge as $e_0\rightarrow 0.$ Instead we introduce $\tau = M+\lop,$ which we will show does have a shift $\Delta\tau$ during an impulse that has finite and linear response to $\vecI$.
-Between impulses, $\tau$ advances with the mean motion as $a^{-3/2}t.$  We define a coordinate system that is uniformly rotating with the unperturbed $\tau_0=t+M_0.$ The two smoothly rotating unit vectors $\uhat_\parallel=(\cos \tau_0, \sin \tau_0)$ and
+The orbital shifts from an impulse can be derived from Gauss's
+equations, \eg\ as presented by \citet[][Section 1.9.2]{tremaine}, but
+we provide here a derivation customized to our case.
+Five of our orbital elements $\vecq$ will be taken from constants of the motion.
+The mean anomaly $\ma$ is not appropriate as the sixth, time-dependent element of $\vecq$ because the derivative $d\ma/dI$ can diverge as $e_0\rightarrow 0.$ Instead we introduce $\tau = \ma+\lop,$ which we will show does have a shift $\Delta\tau$ during an impulse that has finite and linear response to $\vecI$.
+Between impulses, $\tau$ advances with the mean motion as $a^{-3/2}t.$  We define a coordinate system that is uniformly rotating with the unperturbed $\tau_0=t+\ma_0.$ The two smoothly rotating unit vectors $\uhat_\parallel=(\cos \tau_0, \sin \tau_0)$ and
 $\uhat_\perp=(-\sin \tau_0, \cos \tau_0)$ satisfy $\uhat_\parallel
@@
 particular, the components $(e_\parallel,e_\perp)$ of the ellipticity
-vector $\vece$ have the values $e_\parallel=e\cos M$ and $e_\perp=-e\sin M$ in the unperturbed orbit.
+vector $\vece$ have the values $e_\parallel=e\cos \ma$ and $e_\perp=-e\sin \ma$ in the unperturbed orbit.
 The true anomaly $\nu,$ radius $r$, and azimuthal angle $\phi,$ and
@@
 \begin{eqnarray}
-  M & = & M_0 + t = \tau_0,  \nonumber \\
-  \nu & = & M - 2 e_\perp \nonumber \\
+  \ma & = & \ma_0 + t = \tau_0,  \nonumber \\
+  \nu & = & \ma - 2 e_\perp \nonumber \\
   \phi & = & \nu + \lop = \tau - 2e_\perp, \nonumber\\
@@
 
-The first element adopted for $\vecq$ is the shift $\Delta a$, derivable from the orbital energy $E=-1/2a.$
+The first element adopted for $\vecq$ is the semimajor axis $a$, with
+shift derivable from the orbital energy $E=-1/2a.$
 \begin{eqnarray}
@@
   \Delta \mathbf{L} & = & \vecr \times \vecI = (1-e_\parallel) I_\phi \zhat - (1-e_\parallel) I_z \phat \\
-  \quad \Rightarrow \quad \Delta \Lhat_\parallel & = & -2e_\perp I_z
+  \quad \Rightarrow \quad \Delta L_\parallel & = & -2e_\perp I_z
   \label{eq:dLpar}\\
-  \Delta \Lhat_\perp & = & -(1-e_\parallel)  I_z
+  \Delta L_\perp & = & -(1-e_\parallel)  I_z
                   \label{eq:dLperp}
 \end{eqnarray}
-We adopt the changes in direction $(\Delta\Lhat_x, \Delta\Lhat_y)$ as two elements of $\vecq$ specifying the inclination and ascending node $\Omega$ of the perturbed orbit.
-These are a rotation of $(\Delta\Lhat_\parallel, \Delta\Lhat_\perp)$ by the angle $\tau.$  The conversion from the first line above to the subsequent two makes use of the decomposition $\phat=\uhat_\perp + 2e_\perp \uhat_\parallel$ derivable from the value of $\phi$ in Equations~(\ref{eq:kepler}).
+We adopt the $x$ and $y$ components of the orbital angular momentum
+$(L_x, L_y)$ as two elements of $\vecq$ specifying the inclination and
+ascending node $\Omega$ of the perturbed orbit.  These are a rotation
+of $(\Delta L_\parallel, \Delta L_\perp)$ by the angle $\tau.$  The conversion from the first line above to the subsequent two makes use of the decomposition $\phat=\uhat_\perp + 2e_\perp \uhat_\parallel$ derivable from the value of $\phi$ in Equations~(\ref{eq:kepler}).
 
@@
 \end{eqnarray}
-We adopt $\Delta e_x$ and $\Delta e_y$ as two additional components of $\vecq,$ again related through a rotation by $\tau$ to the quantities in Equations~(\ref{eq:depar}) and (\ref{eq:deperp}).
+We adopt $e_x$ and $e_y$ as two additional components of $\vecq,$ again related through a rotation by $\tau$ to the quantities in Equations~(\ref{eq:depar}) and (\ref{eq:deperp}).
 
-The final element of $\vecq$ will be the shift $\Delta\tau$ induced in
-$M+\lop$ by the impulse.  This can be obtained by enforcing the
+The final element of $\vecq$ will be $\tau=\ma+\lop.$  The shift
+induced by an impulse can be obtained by enforcing the
 conditions that the radius $r$ or the azimuthal angle $\phi$ must
-remain constant during the impulse.  To find the latter, we need a formula for $\nu(M)$ as a power series in $e.$  Standard formulae in the literature lead to:
+remain constant during the impulse.  To find the latter, we need a
+formula for $\nu(\ma)$ as a power series in $e.$  Standard formulae in
+the literature \citep[\eg][]{tremaine} lead to:
 \begin{eqnarray}
-  \nu & = & M + 2e \sin M + \frac{5e^2}{4} \sin 2M + O(e^3) \\
+  \nu & = & \ma + 2e \sin \ma + \frac{5e^2}{4} \sin 2\ma + O(e^3) \\
   \label{eq:nue2}
-    & = & M + 2(e\sin M) + \frac{5}{2} (e \sin M) (e \cos M) + O(e^3) 
+    & = & \ma + 2(e\sin \ma) + \frac{5}{2} (e \sin \ma) (e \cos \ma) + O(e^3) 
 \end{eqnarray}
@@
 to zero.
-But the derivatives of $M$ and $\lop$ with impulse can become infinite.
-Instead we introduce a perturbation $\Delta\tau = \Delta(M+\lop).$ We can now express several post-impact quantities as deviations from the unperturbed quantities (subscripted with 0):
+But the derivatives of $\ma$ and $\lop$ with impulse can become infinite.
+Instead we introduce a perturbation $\Delta\tau = \Delta(\ma+\lop).$ We can now express several post-impact quantities as deviations from the unperturbed quantities (subscripted with 0):
 \begin{eqnarray}
-  M & = & \tau-\lop = \tau_0 -\lop + \Delta\tau \\
-  e \sin M & = & e \sin (\tau_0 -\lop) + \Delta\tau \left[ e \cos  (\tau_0 -\lop)\right] \nonumber \\
+  \ma & = & \tau-\lop = \tau_0 -\lop + \Delta\tau \\
+  e \sin \ma & = & e \sin (\tau_0 -\lop) + \Delta\tau \left[ e \cos  (\tau_0 -\lop)\right] \nonumber \\
 \label{eq:esinM}
     & = & -e_{0\perp}-\Delta e_\perp + e_{0\parallel} \Delta\tau +O(e^2) \\
-  e \cos M & = & e \cos (\tau_0 -\lop) - \Delta\tau \left[ e \sin  (\tau_0 -\lop)\right] \nonumber\\
+  e \cos \ma & = & e \cos (\tau_0 -\lop) - \Delta\tau \left[ e \sin  (\tau_0 -\lop)\right] \nonumber\\
 \label{eq:ecosM}
@@
 
-After the impulse, the mean anomaly $M$ advances at a rate of $a^{-3/2}t,$ meaning that an additional term $\Delta\tau = -3\Delta a (t-t_i)/2$ accrues by time $t$.
-The total perturbation of $\tau$ from the nominal value $t+M_0$ becomes
+After the impulse, the mean anomaly $\ma$ advances at a rate of $a^{-3/2}t,$ meaning that an additional term $\Delta\tau = -3\Delta a (t-t_i)/2$ accrues by time $t$.
+The total perturbation of $\tau$ from the nominal value $t+\ma_0$ becomes
 \begin{equation}
@@
 We now have all the information we need to create the $\matA$ matrix.
-Rotating to $\vecq=\{\Delta a, \Delta \tau, \Delta e_x, \Delta e_y, \Delta\Lhat_x, \Delta\Lhat_y\}$ and with $\vecI=\{I_r, I_\phi, I_z\}$ we have, to $O(e)$,
+Rotating to $\vecq=\{a, \tau, e_x, e_y, L_x, L_y\}$ and with $\vecI=\{I_r, I_\phi, I_z\}$ we have, to $O(e)$,
 \begin{equation}
@@
   \Var(e_x) =  \Var(e_y) & = & \nIr/2+2\nIphi \nonumber\\
-  \Var(\Lhat_x) =  \Var(\Lhat_y) & = & \nIz/2
+  \Var( L_x) =  \Var( L_y) & = & \nIz/2
 \label{eq:Cq}
@@
   \label{eq:Ee2}
-  \cos E  =  e\cos M - (e \sin M)^2 + O(e^3).
+  e \cos E  =  e\cos \ma - (e \sin \ma)^2 + O(e^3).
 \end{equation}
@@
 \begin{eqnarray}
-  \Var(r) & \approx & \Var(a) + \frac{1}{2}\Var(e_x) + \frac{1}{2}\Var(e_y) + \frac{\langle e^2\rangle}{2} \Var(\tau) \\
+  \sigma^2_r = \Var(r) & \approx & \Var(a) + \frac{1}{2}\Var(e_x) + \frac{1}{2}\Var(e_y) + \frac{\langle e^2\rangle}{2} \Var(\tau) \\
           & \approx & \frac{T}{2} \nIr + \left( 6T + \frac{3\langle e^2 \rangle}{2} T^3\right) \nIphi.
@@
 \begin{eqnarray}
-  \Var(\phi) & \approx & \Var(\tau) + 2\Var(e_x) + 2\Var(e_y) \\
+\sigma^2_\phi =   \Var(\phi) & \approx & \Var(\tau) + 2\Var(e_x) + 2\Var(e_y) \\
              & \approx & 6T\nIr + \left(8T+3T^3\right) \nIphi.
@@
 \begin{eqnarray}
-  \Delta\theta & = & -\Delta\Lhat_x \frac{x}{r} - \Delta\Lhat_y \frac{y}{r} \\
-  \Rightarrow \quad \Var(\theta) & \approx & \langle\cos^2\tau\rangle \Var(\Lhat_x) + \langle\sin^2\tau\rangle \Var(\Lhat_y) \\
+  \Delta\theta & = & -\Delta L_x \frac{x}{r} - \Delta L_y \frac{y}{r} \\
+  \Rightarrow \quad \sigma^2_\theta = \Var(\theta) & \approx & \langle\cos^2\tau\rangle \Var(L_x) + \langle\sin^2\tau\rangle \Var(L_y) \\
                & \approx & \frac{T}{2} \nIz.
@@
 To convert their variances into linear displacements in the $r,\phi,z$ directions, multiply them all by $a^2,$ or equivalently scale the standard deviations by $a.$
-\item The terms proportional to $T$ show typical diffusive growth $\Delta \propto \sqrt{T},$ and arise from perturbations to the 5 constants of Keplerian motion.
-They represent epicyclic alterations to the original orbit that grow in amplitude as $\sqrt{T}.$  The $T^3$ terms dominate $\Var(\phi)$ by the end of a full orbit, resulting from accumulating delays/advances $\Delta\tau$ in orbital phase $M$ as $a$ and the mean motion diffuse.
+\item The terms proportional to $T$ show typical diffusive growth
+  $\sigma \propto \sqrt{T},$ and arise from random walks in the 5 constants of Keplerian motion.
+They represent epicyclic alterations to the original orbit that grow in amplitude as $\sqrt{T}.$  The $T^3$ terms dominate $\Var(\phi)$ by the end of a full orbit, resulting from accumulating delays/advances $\Delta\tau$ in orbital phase $\ma$ as $a$ and the mean motion diffuse.
 Thus it is the energy-changing impulses, namely $I_\phi,$ that cause the most Brownian motion.
-\item The radial variance also becomes dominated by the $T^3$ term if $e>0,$ as this term represents advances/delays in the radial epicycle phase.
+\item The radial variance also becomes dominated by the $T^3$ term
+  within 2~years at typical MBA values of $e$ (although
+  nearly-circular orbits have $\Var(r)\propto T$)---this term arises from advances/delays in the radial epicycle phase.
 This is the only part of the Brownian motion that whose leading order in $e$ is non-zero, \ie\ which we would have missed by assuming circular initial orbits.
@@
 \emph{The radial noise results look almost identical, but $\approx 7\times$ lower RMS, and likewise the vertical Brownian motion is  very similar but scaled $\approx45\times$ lower than the azimuthal RMS.}
-\item For longer time periods $T$, the azimuthal noise will scale as
+\item For longer time periods $T$, the azimuthal and radial noise will scale as
 $T^{3/2},$ the vertical
-noise as $T^{1/2},$ and the radial noise between these.
+noise as $T^{1/2}.$
 \item The largest contributions to the unmodelled Brownian motion are
@@
 including these 100 MBAs will accrue RMS errors of $\approx 2$~km in
-the azimuthal position, or 1~mas in observed ecliptic longitude.
+the azimuthal position, or 1~mas in ecliptic longitude.
 RMS range and vertical ephemeris errors would be closer to 300 and 40~m, respectively.  We find that, roughly speaking, $\sigma_\phi$
-drops with the number of modeled MBAs as $1/N,$ and
-\begin{equation} \sigma_\phi \equiv \sqrt{\Var(\phi}) \approx
-(200\,\textrm{m}) \times \left(\sigma_M /
-10^{-14}M_\odot\right)^{0.6}.
-  \label{eq:varscale}
-\end{equation}
+drops with the number of modeled MBAs as $1/N.$
+
  In the next decade, LSST will significantly
-increase the number $N$ and lower $\sigma_M$ of deflector MBAs with
+increase the number $N$ of deflector MBAs with
 usefully measured masses from mutual encounters by tracking hugely
-more potential tracers.
+more potential tracers and lowering $\sigma_M.$.
 The conservative estimate given in Section~\ref{sec:mbaprops} is that LSST will lower $\sigma_M$ to $\approx10^{-13.7},$ which our calculations indicate lowers the $\sigma_\phi$ values to $\approx500$~m or 0.25~mas.
@@
 impulses, the astrometric signal will be smaller, but the Brownian
-motion noise is lower too.  Since $\Delta\phi/\sigma_\phi \propto
-T^{-1/2},$ this signal-to-noise ratio would still be $\gg 5$ for
-surveys of several decades' duration.  We conclude that our choice of $I_{\rm
+motion noise is lower too. 
+We conclude that our choice of $I_{\rm
 det}$ is not ruined by the presence of Brownian motion noise on the
-tracer asteroids, rather it will be driven by the observational
-uncertainties.
+tracer asteroids, impulses of this size will remain strongly detected.
 
 \section{Consequences}
+\label{sec:discuss}
 \subsection{Implications for future inferences}
 
-An important question is: will the positional noise due to unmodelled MBA encounters degrade inferences being made with measurements of tracer MBAs?
-If the Brownian noise is below other sources of measurement noise, then the answer must be ``no.''
+An important question is: will the positional noise due to unmodelled
+MBA encounters degrade inferences being made with measurements of
+tracer MBAs?  If the Brownian noise is well below other sources of measurement noise, then the answer must be ``no.''
 The per-epoch uncertainties on solar-system object astrometry from Gaia\footnote{We take one Gaia transit as an ``epoch'' here.} are reported to be as low as 0.3~mas for very bright sources ($G=13$), rising to $\approx10$~mas at its $G=21$ limit where most detections lie \citep[][Figure 6]{gaiass3}.
@@
 
-These per-epoch comparisons are, however, misleading.
-The Brownian noise is highly correlated between different epochs of observation of a given asteroid, while the measurement errors are generally uncorrelated, and therefore the impact on inferences is not described solely by comparison of RMS values of each.
-The covariance structure of Brownian noise can work either to the benefit or the detriment of the inference.
-In an unfavorable case, the eigenvectors of this covariance matrix with the largest uncertainties look similar to the signal of interest in the inference, \ie\ the derivative of the MBA's position with respect to a parameter.
-If there is a single dominant eigenvector of the Brownian noise which looks just like our signal, then the effect per-epoch Brownian noise does not average down at all as we obtain $M$ observations, but the measurement noise drops as $\sqrt{M}.$  In a favorable case, all of the strong eigenvectors of the Brownian noise covariance are orthogonal to the signals of interest, and the Brownian noise causes no degradation of the inference, \ie\ its spatiotemporal structure cannot mimic that of the signal we seek.
+\emph{These per-epoch comparisons are, however, misleading.} The
+Brownian noise is highly correlated between different epochs of
+observation of a given asteroid, while the measurement errors are
+generally uncorrelated, and therefore the impact on inferences is not
+described solely by comparison of RMS values of each.  The covariance
+structure of Brownian noise can work either to the benefit or the
+detriment of the inference.  In an unfavorable case, the eigenvectors
+of the Brownian covariance matrix with the largest uncertainties look similar to the signal of interest in the inference, \ie\ the derivative of the MBA's position with respect to a parameter.
+If there is a single dominant eigenvector of the Brownian noise which looks just like our signal, then the per-epoch Brownian noise does not average down at all as we obtain $M$ observations, but the measurement noise drops as $\sqrt{M}.$  In a favorable case, all of the strong eigenvectors of the Brownian noise covariance are orthogonal to the signals of interest, and the Brownian noise causes no degradation of the inference, \ie\ its spatiotemporal structure cannot mimic that of the signal we seek.
 
-Consider first the worst-case scenario, where the Brownian noise does not average down at all across observations, but measurement errors gain a factor $\sqrt{M}.$   \citet{gaiafpr} report typically $\sim60$ epochs per asteroid in 66 months, suggesting $O(100)$ epochs per asteroid in the full 11-year observing period.
+Consider first the worst-case scenario, where the Brownian noise does
+not average down at all across observations, but measurement errors
+drop by a factor $\sqrt{M}.$   \citet{gaiafpr} report typically $\sim60$ epochs per asteroid in 66 months, suggesting $O(100)$ epochs per asteroid in the full 11-year observing period.
 At Gaia's faint end, this means that Brownian noise with present asteroid knowledge would have comparable impact to the measurement errors, and dominate inferences at the bright end.
@@
 These are quite distinct from the random walk in $\Delta\tau$ that is the dominant effect of Brownian motion, and therefore Brownian noise with cause less degradation to the Planet X inference than the per-epoch $\sigma_\phi$ would suggest.
-Another example would be a gravitational perturbation whose signature is primarily a nodal precession, which is a signal manifesting primarily as out-of-plane motions---we have seen that Brownian motion out of the orbital plane is $>40\times$ smaller than the $\phi$ component and grows at $T{-1/2}$ instead of $T^{3/2},$ so the per-epoch $\sigma_\phi$ of Brownian noise will grossly overestimate its effect on the inference.
+Another example would be a gravitational perturbation whose signature is primarily a nodal precession, which is a signal manifesting primarily as out-of-plane motions---we have seen that Brownian motion out of the orbital plane is $>40\times$ smaller than the $\phi$ component and grows at $T^{-1/2}$ instead of $T^{3/2},$ so the per-epoch $\sigma_\phi$ of Brownian noise will grossly overestimate its effect on the inference.
 
@@
 
-A better approach, however, would probably be to increase the number $N$ of MBAs considered as active bodies in the ephemeris fitting, leaving their masses as free parameters.
+A better approach, however, would probably be to increase the number
+$N$ of MBAs considered as active bodies in the ephemeris fitting,
+leaving their masses as free parameters whether or not we expect them
+to be constrained to $S/N\gtrsim1$ by the data.
 Although the extreme case of having $N=O(10^6)$ active bodies would be a substantial computational challenge, having $N=10^3$--$10^4$ is not, especially since the MBA-to-MBA perturbations can be adequately calculated using Newtonian dynamics instead of a fully relativistic treatment.
-In this approach, all of the covariances between different observations of the same tracer, and between tracers, will be captured in the analysis for deflections due to the $N$ deflectors.
-This is true even if the ephemeris fit does not yield $S/N>1$ on every deflector.
+In this approach, all of the covariances between different observations of the same tracer, and between tracers, will be captured in the analysis for deflections due to these $N$ deflectors.
 If we \emph{choose $N$ to be large enough that the Brownian noise from the unmodelled MBAs is negligible to the error budget,} then we need not explicitly propagate the Brownian noise.
@@
 \end{equation}
-where the last expression adopts our standard values $q=0.25, \rho=2500\,{\rm kg}\, {\rm m}^{-3}.$ The observational consequence of this constant radial force is an $H$-dependent shift $\Delta r = -(1+q)\Psi a$ at a fixed period of rotation, which does not resemble any expected effect of Brownian motion.
-More important is the Yarkovsky effect, in which anisotropic re-emission of the absorbed radiation leads to a net force that is fraction $g_Y(1-q)$ of the incident radiation pressure.
+where the last expression adopts our standard values $q=0.25,
+\rho=2500\,{\rm kg}\, {\rm m}^{-3}.$ The observational consequence of
+this constant radial force is an $H$-dependent shift $\Delta r =
+-(1+q)\Psi a$ at a fixed period of rotation, which does not resemble
+any expected effect of Brownian motion.  More important is the
+Yarkovsky effect, in which anisotropic re-emission of the absorbed
+radiation leads to a net force that is a fraction $g_Y(1-q)$ of the incident radiation pressure.
 The Yarkovsky effect has been detected for several NEAs but no individual MBAs, with values $g_Y \approx 0.1$ \citep[Yarkovsky effect and data are review by][]{yarkovsky}.
@@
 
-Taking a case where the Yarkovsky force to be in the prograde azimuthal direction, and using our system of units with $a=1$ and $v_c=1,$ this leads to a continuous power $dE/dt\approx g_Y\Psi$ which in turn implies $da/dt\approx2g_Y\Psi.$   The mean motion rate is hence $1-3 g_Y\Psi t$ and the accumulated shift in mean anomaly over time $T$ amounts to $\Delta\phi = -3 g_Y \Psi  T^2/2.$   In this case, a rough comparison of the accumulated $\Delta\phi$ from the Yarkovsky effect to the $\sigma_\phi$ of noise from unmodelled Brownian motion, after a time interval $T$, is
+Taking a case where the Yarkovsky force is in the prograde azimuthal direction, and using our system of units with $a=1$ and $v_c=1,$ this leads to a continuous power $dE/dt\approx g_Y\Psi$ which in turn implies $da/dt\approx2g_Y\Psi.$   The mean motion rate is hence $1-3 g_Y\Psi t$ and the accumulated shift in mean anomaly over time $T$ amounts to $\Delta\phi = -3 g_Y \Psi  T^2/2.$   In this case, a rough comparison of the accumulated $\Delta\phi$ from the Yarkovsky effect to the $\sigma_\phi$ of noise from unmodelled Brownian motion, after a time interval $T$, is
 \begin{equation}
-   \frac{\Delta\phi}{\sigma_\phi} \approx \frac{ 3 g_Y \Psi T^2/2}{\sqrt{3 T^3 \nIphi}} \approx \sqrt{T/10\,{\rm yr}} 10^{0.2(H-20.2)}
+\frac{\textrm{Yarkovsky}}{\textrm{Brownian}} = 
+   \frac{\Delta\phi}{\sigma_\phi} \approx \frac{ 3 g_Y \Psi
+     T^2/2}{\sqrt{3 T^3 \nIphi}} \approx \sqrt{T/10\,{\rm yr}} \times 10^{0.2(H-20.2)}
      \label{eq:yarkovsky}
@@
 
-In conclusion, the Brownian motion from unmodelled or mis-modelled mutual impulses between MBAs causes RMS azimuthal displacements of $\approx2$~km (or 1 mas in angular position) over 10 years, with current knowledge of the masses of large asteroids.
-This is large enough to become a significant part of the error budget for many inferences from Gaia MBA astrometry.
-The knowledge on large-MBA masses likely to be gained from LSST tracking of MBAs should reduce this by a factor of $\ge4,$ at which point the Brownian noise will be comparable to Gaia or LSST measurement uncertainties only for the brighter targets of each survey.
-In all cases, the Brownian motion is $\approx7\times$ smaller in the radial direction and $\approx45\times$ in the vertical direction.
-The RMS azimuthal displacement grows with time as $T^{3/2}$, the vertical RMS as $T^{1/2}$, and the radial RMS eventual switches from the latter to the former at some time driven by the orbit eccentricity.
+In conclusion, the Brownian motion from unmodelled or mis-modelled
+mutual impulses between MBAs causes RMS azimuthal displacements of
+$\approx2$~km (or 1 mas in angular position) over 10 years, with
+current knowledge of the masses of large asteroids.  This is large
+enough to become a significant part of the error budget for many
+inferences from Gaia MBA astrometry.  The knowledge on large-MBA
+masses likely to be gained from LSST tracking of MBAs should reduce
+this by a factor of $\ge4,$ at which point the Brownian noise will be
+comparable to Gaia or LSST measurement uncertainties only for the
+brighter targets of each survey.  In all cases, the Brownian motion is
+$\approx7\times$ smaller in the radial direction and $\approx45\times$
+in the vertical direction.  The RMS azimuthal and radial displacements
+grow with time as $T^{3/2}$, and the vertical RMS as $T^{1/2}.$
 
-Making inferences from MBA astrometry becomes more complicated when Brownian motion is a significant contributor to the error budget, but a good path forward is to do all model fitting by including the $\approx1000$ largest MBAs as active bodies in the ephemeris model and leaving their masses as free parameters.
-This will constrain the masses of these bodies to the full extent that the data allow, and will propertly propagate any remaining uncertainty in those masses into the other inferences that one is trying to make.
-The noise contributed by the remaining millions of unmodelled accuracy would be just $\approx60$~m RMS in the azimuthal direction (about 30~$\mu$as), and a few $\mu$as or less in the other axes, small enough to ignore entirely.
+Making inferences from MBA astrometry becomes more complicated when
+Brownian motion is a significant contributor to the error budget, but
+a good path forward is to do all model fitting by including the
+$\approx1000$ largest MBAs as active bodies in the ephemeris model and
+leaving their masses as free parameters.  This will constrain the
+masses of these bodies to the full extent that the data allow, and
+will properly propagate any remaining uncertainty in those masses into
+the other inferences that one is trying to make.  The noise
+contributed by the remaining millions of unmodelled MBAs would be just
+$\approx60$~m RMS in the azimuthal direction (about 30~$\mu$as), and a
+few $\mu$as or less in the other axes, small enough to ignore entirely
+for the accuracy of imaging astrometry for the foreseeable future.
 Occultation-based astrometry is much more accurate (but has fewer possible measurements per tracer), and would be significantly impacted by Brownian motion until $\gtrsim1000$ MBAs have mass estimates at modest $S/N$ level, although the precise impact will depend on whether the signals one is seeking are parallel or orthogonal to the kind of variations caused by Brownian motion.
@@
 \begin{acknowledgments}
-  This work was supported by NSF grant AST-????.
+  This work was supported by NSF grant AST-2205808.  GMB is grateful for
+  the thoughts and efforts of Negin Najafi, Daniel Gomes, Matt Holman,
+  David Trilling, and Will Grundy.
   This research has made use of data and/or services provided by the International Astronomical Union's Minor Planet Center. 

### 2025-03-02 main.tex
Edit run-through done, just need equation check.

@@
 \newcommand{\vcirc}{v_c}
-\newcommand{\vrel}{v_{\rm r}}
 \newcommand{\Msun}{M_\odot}
@@
 the non-gravitational forces that drive many forms of planetary
-migration, \eg\ the maintainence of the near-Earth asteroid (NEA)
+migration, such as the maintainence of the near-Earth asteroid (NEA)
 population \citep[reviewed by][]{yarkovsky}.  While most applications of precision orbit determination
@@
 to include 250 MBAs as gravitating bodies in the fit to the full
-corpus of observations at the time.  For only $\approx 1/2$ of the
+corpus of observations at the time.  For only $\approx 1/2$ of these
 bodies did the fit yield masses $>3\times$ the uncertainties.  With
@@
 contributions to our results that are leading
-order in the MBA orbital eccentricity $e$ of the tracer population.
+order in the orbital eccentricity $e$ of the tracer MBA.
 We will assume throughout that the joint
@@
 unit vectors $\rhat, \phat$ rotate with the target asteroid.  The mean
-anomaly at time $t=0$ is $\ma_0.$
+anomaly at time $t=0$ is $\ma_0.$  In general, subscripts of $0$ will
+indicate properties of the unperturbed orbit.
 
@@
 We describe all encounters with other asteroids in the impulse approximation, defining $\vecI$ as the $\Delta\vecv$ imparted on the target by the deflector.
-In our units, the gravitational impulse imparted by a deflector of mass $M_d$ approaching at impact parameter $\vecb=b\bhat$ and relative velocity $\vrel$ is
+In our units, the gravitational impulse imparted by a deflector of mass $M_d$ approaching at impact parameter $\vecb=b\,\bhat$ and relative velocity $v$ is
 \begin{equation}
@@
 where the angle brackets indicate an average over possible
-realizations of the impulse history, and over the mean anomaly $\ma$ at the time of observation.
-To do so, we will introduce an intermediate set of 6 parameters $\vecq$ describing the orbital elements induced by the impulses.
-The $\vecq$ components will be selected to respond linearly to $\vecI$ at $|I|\ll 1.$
+realizations of the impulse history, and over the mean anomaly $\ma$
+at the time of observation.  To do so, we will introduce an
+intermediate set of 6 orbital elements $\vecq,$ selected to respond linearly to $\vecI$ at $|I|\ll 1.$
 We will derive the matrix $\matA$ that describes the orbital-element shifts at time $T$ that arise from an impulse at time $t_i$: 
@@
 \end{eqnarray}
-which we would wish to average over the phase of the MBA orbit or mean anomaly $\ma$ at the time of observation $T.$
+which we will average over the phase of the MBA orbit at the
+observation time $T,$ \ie\ average over mean anomaly $\ma.$
 
@@
 random walk of the asteroid's orbits can be expressed as an integral
-over the absolute magnitude $H$ of the deflectors, the velocity $v$,
+over the absolute magnitude $H$ of the deflectors, the relative velocity $v$,
 impact parameter $b,$ and polar/azimuthal angles $\theta,\phi$ of the
@@
                                              \frac{dN}{dH} \frac{dn}{dv\,db\,d\theta\,d\phi}
-                                              I^2_\gamma(H,v,b,\theta\phi)
+                                              I^2_\gamma(H,v,b,\theta,\phi)
   \\
@@
     (arbitrary) normalization, so the relative heights properly
-    reflect the frequency of encounters between different MBA
-    regions.}
+    reflect the relative frequency of encounters among
+    pairs of deflector and tracer regions.}
   \label{fig:rtz}
@@
     the functional form from \eqq{eq:lsstmba} shown as the dotted
-    curve.}
+    curve. The Inner and Outer regions are slightly rescaled to help
+    visual comparison of the curves' shapes in the three MBA regions.}
   \label{fig:counts}
@@
 attainable from following a few individual tracers, and calculate
-the Brownian-motion noise as a function of $\sigma_M.$  We take
-\begin{equation}
-  \sigma_M(H) = \textrm{min}\left[ \sigma_M, M(H) \right],
-\end{equation}
-because a given MBA could have a mass determination accurate to
-$\sigma_M$ if included in the ephemeris model, or if $M(H)<\sigma_M,$
-it would be omitted from the ephemeris model.
+the Brownian-motion noise as a function of $\sigma_M.$
 
@@
 \subsubsection{Individually detectable impulses}
-Examining the $\sigma_I$ distribution forecasted for LSST data on tracers (Figure~\ref{fig:sigmaI}), we find that for impulses occurring near the midpoint of the survey, $\approx80\%$ of all cases yield $\sigma_I<10^{-4}\,{\rm m}\,{\rm
+Examining the $\sigma_I$ distribution forecasted for LSST data on
+tracers by \citet{negin}, we find that for impulses occurring near the midpoint of the survey, $\approx80\%$ of all cases yield $\sigma_I<10^{-4}\,{\rm m}\,{\rm
   s}^{-1},$ some an order of magnitude lower.  We conservatively
@@
 The true anomaly $\nu,$ radius $r$, and azimuthal angle $\phi,$ and
-the velocity components of the initial orbit are, to first order in $e$,
+the velocity components of the unperturbed orbit are, to first order in $e$,
 \begin{eqnarray}
@@
 induced by an impulse can be obtained by enforcing the
-conditions that the radius $r$ or the azimuthal angle $\phi$ must
+condition that the radius $r,$ or the azimuthal angle $\phi,$ must
 remain constant during the impulse.  To find the latter, we need a
@@
 But the derivatives of $\ma$ and $\lop$ with impulse can become infinite.
-Instead we introduce a perturbation $\Delta\tau = \Delta(\ma+\lop).$ We can now express several post-impact quantities as deviations from the unperturbed quantities (subscripted with 0):
+Instead we introduce a perturbation $\Delta\tau = \Delta(\ma+\lop).$ We can now express several post-impact quantities as deviations from the unperturbed quantities:
 \begin{eqnarray}
@@
 
-After the impulse, the mean anomaly $\ma$ advances at a rate of $a^{-3/2}t,$ meaning that an additional term $\Delta\tau = -3\Delta a (t-t_i)/2$ accrues by time $t$.
-The total perturbation of $\tau$ from the nominal value $t+\ma_0$ becomes
-\begin{equation}
-  \Delta\tau= \left[-2+\frac{3e_\parallel}{2} +3 e_\perp (t-t_i)\right] I_r  - \left[3(1+e_\parallel)(t-t_i) + e_\perp\right] I_\phi.
+After the impulse, the mean anomaly $\ma$ advances at a rate of $a^{-3/2}t,$ meaning that an additional term $\Delta\tau = -3\Delta a (T-t_i)/2$ accrues by time $T$.
+The total perturbation of $\tau$ from the nominal value $T+\ma_0$ becomes
+\begin{equation}
+  \Delta\tau= \left[-2+\frac{3e_\parallel}{2} +3 e_\perp (T-t_i)\right] I_r  - \left[3(1+e_\parallel)(T-t_i) + e_\perp\right] I_\phi.
   \label{eq:dtau}
@@
 \begin{itemize}
-\item We find that \emph{the linear displacements are nearly identical for Inner, Middle, and Outer regions of the asteroid belt.}  The angular displacements vary slightly, inversely to their distance from the Sun.
-\item We plot the azimuthal noise.
-\emph{The radial noise results look almost identical, but $\approx 7\times$ lower RMS, and likewise the vertical Brownian motion is  very similar but scaled $\approx45\times$ lower than the azimuthal RMS.}
+\item \emph{The linear displacements are nearly identical for Inner,
+    Middle, and Outer regions of the asteroid belt.}  The angular
+  displacements vary slightly, inversely to their distance from the
+  Sun.  Figure~\ref{fig:result} plots only the Middle MBAs'
+  deviations, although the count $N$ of modelled MBAs is the sum for
+  all three regions.
+\item We plot the azimuthal noise.  \emph{The radial noise results
+    have nearly identical scaling with $H$ and $\sigma_M$, but
+    $\approx 7\times$ lower RMS. Likewise the vertical Brownian motion
+    has the same shape but $\approx45\times$ lower than the azimuthal RMS.}
 \item For longer time periods $T$, the azimuthal and radial noise will scale as
@@
 usefully measured masses from mutual encounters by tracking hugely
-more potential tracers and lowering $\sigma_M.$.
-The conservative estimate given in Section~\ref{sec:mbaprops} is that LSST will lower $\sigma_M$ to $\approx10^{-13.7},$ which our calculations indicate lowers the $\sigma_\phi$ values to $\approx500$~m or 0.25~mas.
+more potential tracers and lowering $\sigma_M.$  The conservative estimate given in Section~\ref{sec:mbaprops} is that LSST will lower $\sigma_M$ to $\approx10^{-13.7},$ which our calculations indicate lowers the $\sigma_\phi$ values to $\approx500$~m or 0.25~mas.
 
@@
 det}$ is not ruined by the presence of Brownian motion noise on the
-tracer asteroids, impulses of this size will remain strongly detected.
+tracer asteroids---impulses of this size will remain strongly detected.
 
@@
 
-The algorithms used to make inferences based on range or astrometric data will need to be changed when Brownian motion becomes a significant source of error, to accomodate the fact that positional errors on a given tracer are highly correlated between epochs.
-The full covariance matrix of the tracer's observations will be needed rather than typical practice of considering measurements as independent.
-A small modification to our $\matB$ matrices would be needed to calculate the covariance of the Brownian noise between observations at different times $T$ and $T^\prime.$  The Brownian motion should be fairly uncorrelated between different MBAs, because the varying encounter geometries randomize the influence of any particular deflector's mass error on the trajectories of the tracers it influences.
+The algorithms used to make inferences based on range or astrometric
+data will need to be changed when Brownian motion becomes a
+significant source of error, to accomodate the fact that positional
+errors on a given tracer are highly correlated between epochs.  The
+full covariance matrix of the tracer's observations will be needed
+rather than typical practice of considering measurements as
+independent.  A small modification to our $\matB$ matrices would be
+needed to calculate the covariance of the Brownian noise between
+observations at different times $T$ and $T^\prime.$  The Brownian
+motion should be fairly uncorrelated between different MBAs, because
+the varying encounter geometries randomize the influence of any
+particular deflector's mass error on the trajectories of the tracers
+it influences.
+
+A further difficulty in propagating unmodelled deflections into
+inferences is that the probability distribution of the position shifts
+will be non-Gaussian, because the RMS is dominated by the
+contributions of relatively few encounters with the largest unmodelled
+asteroids.  The central limit theorem will not be applicable over
+decade timescales.
 
@@
 leaving their masses as free parameters whether or not we expect them
-to be constrained to $S/N\gtrsim1$ by the data.
-Although the extreme case of having $N=O(10^6)$ active bodies would be a substantial computational challenge, having $N=10^3$--$10^4$ is not, especially since the MBA-to-MBA perturbations can be adequately calculated using Newtonian dynamics instead of a fully relativistic treatment.
+to be constrained to $S/N\gtrsim1$ by the data.  Although the extreme
+case of having $N=O(10^6)$ active bodies' masses as free parameters in
+the ephemeris fit would be a substantial computational challenge,
+having $N=10^3$--$10^4$ should be manageable.
+The MBA-to-MBA perturbations can be adequately calculated using Newtonian dynamics instead of a fully relativistic treatment.
 In this approach, all of the covariances between different observations of the same tracer, and between tracers, will be captured in the analysis for deflections due to these $N$ deflectors.
@@
  \Psi \equiv \frac{F_{\rm rad}}{F_{\rm grav}} = \frac{\pi R^2 L_\odot/4\pi c}{4\pi GM_\odot R^3\rho/3}
-   = \frac{3}{16\pi} \frac{L_\odot}{GM_\odot c R \rho} \approx 3\times10^{-10} 10^{0.2(H-16)},
+   = \frac{3}{16\pi} \frac{L_\odot}{GM_\odot c R \rho} \approx
+   3\times10^{-10} \times 10^{0.2(H-16)},
 \end{equation}
 where the last expression adopts our standard values $q=0.25,
-\rho=2500\,{\rm kg}\, {\rm m}^{-3}.$ The observational consequence of
-this constant radial force is an $H$-dependent shift $\Delta r =
--(1+q)\Psi a$ at a fixed period of rotation, which does not resemble
-any expected effect of Brownian motion.  More important is the
+\rho=2500\,{\rm kg}\, {\rm m}^{-3}.$ The primary observational consequence of
+this constant radial force is a time-independent, $H$-dependent 
+shift $\Delta r = -(1+q)\Psi a$ of the radial position a body should
+have at a given period.  
+This does not resemble
+any expected effect of Brownian motion, and also very difficult to
+measure since radar ranging to MBAs is infeasible.  More important for
+observations is the
 Yarkovsky effect, in which anisotropic re-emission of the absorbed
@@
 
-When Yarkovsky forces on MBAs are estimated using the $H$-dependent dispersion of $a$ for asteroid families, the effective ``observation'' time period is typically several Myr.
-Over these long time scales, the temporally coherent Yarkovsky effect can dominate over Brownian motion for much larger bodies, which is why the technique is successful.
+When Yarkovsky forces on MBAs are estimated using the $H$-dependent
+dispersion of $a$ for asteroid families, the effective ``observation''
+time period is typically several Myr.  Over these long time scales,
+\eqq{eq:yarkovsky} shows that the temporally coherent Yarkovsky effect
+can dominate over Brownian motion for bodies up to 10's or 100's of km
+in diameter, which is why the technique is successful.
 

### 2025-03-03 main.tex
Fix A table for dex, dey
For PSJ submission

@@
   The explosion of high-precision astrometric data on main-belt
-  asteroids (MBAs) opens the door to their use for many possible
+  asteroids (MBAs) enables new
   inferences on gravitational and non-gravitational forces present in
-  this region.  We estimate the size of positional shifts on MBAs
+  this region.  We estimate the size of MBA motions
   caused by mutual gravitational encounters with other MBAs that are
@@
   MBAs that cannot be predicted from the ephemeris, and therefore
-  serves as noise on inferences from the MBAs?  We calculate the RMS
+  serves as noise on inferences from the MBAs?  We estimate the RMS
   azimuthal shift $\sigma_\phi$ of this ``Brownian noise''  by
@@
   directions among known MBAs, combined with analytical propagation
-  into future positional uncertainties.  The result depends strongly
-  on how well the masses of the larger MBAs are known.  At current
-  level of asteroid-mass knowledge, $\sigma_\phi$ rises to
+  into future positional uncertainties. At current
+  levels of asteroid-mass knowledge, $\sigma_\phi$ rises to
   $\approx2$~km or $\approx1$~mas over $T=10$~yr, increasing as
-  $T^{3/2}.$  This is large enough to degrade many inferences from
-  Gaia and LSST MBA data, although the LSST data will also improve MBA
+  $T^{3/2},$  large enough to degrade many inferences from
+  Gaia and LSST MBA data. LSST data will, however, improve MBA
   mass knowledge enough to lower this Brownian uncertainty by
   $\ge4\times.$  Radial and vertical Brownian noise  at $T=10$~yr are
-  suppressed by factors $\approx7$ and $\approx45,$ respectively,
-  compared to the azimuthal signal after 10~yr, and grow at $T^{3/2}$
-  and $T^{1/2}.$  We suggest that for full exploitation of Gaia and
+  factors $\approx7$ and $\approx45,$ respectively,
+  lower than the azimuthal noise, and grow as $T^{3/2}$
+  and $T^{1/2}.$  For full exploitation of Gaia and
   LSST MBA data, ephemeris models should include the
@@
   the uncertainties from these 1000 sources' deflections into desired
-  inferences.
-The RMS value of deflections from less-massive MBAs is limited to $\sigma_\phi\approx60$~m or 30~$\mu$as, small enough to ignore until occultation-based position data become ubiquitously available for MBAs.
+  inferences. The RMS value of deflections from less-massive MBAs is
+  then just $\sigma_\phi\approx60$~m or 30~$\mu$as, small enough to ignore until occultation-based position data become ubiquitously available for MBAs.
 \end{abstract}
 
-\keywords{asteroids}
+\keywords{Main belt asteroids (2036), Asteroid dynamics (2210), Astrometry (80)}
 
@@
  -2+e\left[1.5\cos\tau-3(T-t_i)\sin\tau\right] & e\sin\tau -3(1+e\cos\tau)(T-t_i) & 0\\
- -\cos\tau-2e\sin^2\tau & 2\cos\tau - 1.5e\sin 2\tau & 0\\
-  -\sin\tau + e\sin 2\tau & 2\sin\tau-3e\cos^2\tau & 0 \\
+ \sin\tau+e\sin 2\tau & 2\cos\tau - 3e\sin^2\tau & 0\\
+  -\cos\tau - 2e\sin^2\tau & 2\sin\tau-1.5e\sin 2\tau & 0 \\
 0 & 0 & \sin\tau + 0.5 e \sin 2\tau \\
@@
 
-% \begin{equation}
-%   \matC^2 = \left(\begin{array}{cccccc}
-%     4T \nIphi & -3T^2\nIphi & e\left(\nIr + 5\nIphi\right) & e\nIr & 0 & 0 \\
-% -3T^2\nIphi & 4\nIr + 3T^3\nIphi & -e(11T\nIr/4 + 15T^2\nIphi/4) & e\left[(2T-3T^2/4)\nIr+(-T+9T^2/4)\nIphi\right] & 0 & 0\\
-%  e\left(\nIr + 5\nIphi\right) & -e(11T\nIr/4 + 15T^2\nIphi/4) & \nIr/2+2\nIphi & 0 & 0 & 0 \\
-%  e\nIr &  e\left[(2T-3T^2/4)\nIr+(-T+9T^2/4)\nIphi\right]  & 0 & \nIr/2+2\nIphi & 0 & 0 \\
-% 0 & 0 & 0 & 0 & \nIz/2 & 0 \\
-% 0 & 0 & 0 & 0 & 0 \nIz/2 \\
-%\end{array}\right)
-% \end{equation}
  \subsection{Astrometric and ranging shifts from elements shifts}

### 2025-03-03 main.tex
As submitted / posted

@@
-%\documentclass[onecolumn]{aastex631}
-\documentclass[linenumbers, onecolumn]{aastex631}
+\documentclass[onecolumn]{aastex631}
+%\documentclass[linenumbers, onecolumn]{aastex631}
 
@@
 For the values of $b_{\rm min}$ and $b_{\rm max},$ we adopt heuristics
-for the nature of impulses of interest.  We set $b_{\rm max}$ by the
+for the nature of impulses of interest.  We set $b_{\rm max}$ by 
 requiring that $b/v<1,$ such that the ``impulse'' is being applied
@@
 If, for example, one is searching for the tidal force of a Planet X, this will be manifested primarily as a collective apsidal precession and a quadrupolar deviation of the orbit from its ellipse that does not evolve with time.
-These are quite distinct from the random walk in $\Delta\tau$ that is the dominant effect of Brownian motion, and therefore Brownian noise with cause less degradation to the Planet X inference than the per-epoch $\sigma_\phi$ would suggest.
+These are quite distinct from the random walk in $\Delta\tau$ that is the dominant effect of Brownian motion, and therefore Brownian noise will cause less degradation to the Planet X inference than the per-epoch $\sigma_\phi$ would suggest.
 Another example would be a gravitational perturbation whose signature is primarily a nodal precession, which is a signal manifesting primarily as out-of-plane motions---we have seen that Brownian motion out of the orbital plane is $>40\times$ smaller than the $\phi$ component and grows at $T^{-1/2}$ instead of $T^{3/2},$ so the per-epoch $\sigma_\phi$ of Brownian noise will grossly overestimate its effect on the inference.

### 2025-03-31 main.tex
Fix Yarkovsky discussion

@@
-\documentclass[onecolumn]{aastex631}
-%\documentclass[linenumbers, onecolumn]{aastex631}
+%\documentclass[onecolumn]{aastex631}
+\documentclass[linenumbers, onecolumn]{aastex631}
 
@@
 \begin{abstract}
-  The explosion of high-precision astrometric data on main-belt
+  % The explosion of high-precision astrometric data on main-belt
+  % asteroids (MBAs) enables new
+  % inferences on gravitational and non-gravitational forces present in
+  % this region.  We estimate the size of MBA motions
+  % caused by mutual gravitational encounters with other MBAs that are
+  % either omitted from ephemeris models or have uncertain mass
+  % estimates. In other words, what is the typical Brownian motion among
+  % MBAs that cannot be predicted from the ephemeris, and therefore
+  % serves as noise on inferences from the MBAs?  We estimate the RMS
+  % azimuthal shift $\sigma_\phi$ of this ``Brownian noise''  by
+  % numerical estimation of the distribution of impulse sizes and
+  % directions among known MBAs, combined with analytical propagation
+  % into future positional uncertainties. At current
+  % levels of asteroid-mass knowledge, $\sigma_\phi$ rises to
+  % $\approx2$~km or $\approx1$~mas over $T=10$~yr, increasing as
+  % $T^{3/2},$  large enough to degrade many inferences from
+  % Gaia and LSST MBA data. LSST data will, however, improve MBA
+  % mass knowledge enough to lower this Brownian uncertainty by
+  % $\ge4\times.$  Radial and vertical Brownian noise  at $T=10$~yr are
+  % factors $\approx7$ and $\approx45,$ respectively,
+  % lower than the azimuthal noise, and grow as $T^{3/2}$
+  % and $T^{1/2}.$  For full exploitation of Gaia and
+  % LSST MBA data, ephemeris models should include the
+  % $\approx1000$~largest asteroids as active bodies with free masses,
+  % even if not all are well constrained.  This will correctly propagate
+  % the uncertainties from these 1000 sources' deflections into desired
+  % inferences. The RMS value of deflections from less-massive MBAs is
+  % then just $\sigma_\phi\approx60$~m or 30~$\mu$as, small enough to
+  % ignore until occultation-based position data become ubiquitously
+  % available for MBAs.
+%
+High-precision astrometric data on main-belt
 asteroids (MBAs) enables new
 inferences on gravitational and non-gravitational forces present in
-  this region.  We estimate the size of MBA motions
-  caused by mutual gravitational encounters with other MBAs that are
-  either omitted from ephemeris models or have uncertain mass
-  estimates. In other words, what is the typical Brownian motion among
-  MBAs that cannot be predicted from the ephemeris, and therefore
-  serves as noise on inferences from the MBAs?  We estimate the RMS
-  azimuthal shift $\sigma_\phi$ of this ``Brownian noise''  by
-  numerical estimation of the distribution of impulse sizes and
-  directions among known MBAs, combined with analytical propagation
+this region.  We estimate the typical Brownian motion due to mutual encounters among
+MBAs that cannot be predicted from the ephemeris of larger bodies, and therefore
+serves as noise on inferences from the MBAs.  The RMS
+azimuthal shift $\sigma_\phi$ of this ``Brownian noise''  is predicted
+using the distribution of known MBA orbits, combined with analytical propagation
 into future positional uncertainties. At current
@@
 mass knowledge enough to lower this Brownian uncertainty by
-  $\ge4\times.$  Radial and vertical Brownian noise  at $T=10$~yr are
-  factors $\approx7$ and $\approx45,$ respectively,
-  lower than the azimuthal noise, and grow as $T^{3/2}$
-  and $T^{1/2}.$  For full exploitation of Gaia and
+$\ge4\times.$  Radial (vertical) Brownian noise  at $T=10$~yr is
+$\approx7\times$ ($\approx45\times$) 
+lower than the azimuthal noise.  For full exploitation of Gaia and
 LSST MBA data, ephemeris models should include the
@@
 Yarkovsky effect, in which anisotropic re-emission of the absorbed
-radiation leads to a net force that is a fraction $g_Y(1-q)$ of the incident radiation pressure.
-The Yarkovsky effect has been detected for several NEAs but no individual MBAs, with values $g_Y \approx 0.1$ \citep[Yarkovsky effect and data are review by][]{yarkovsky}.
-This is also consistent with detections of the Yarkovsky effect by examining the $\Delta a$ dependence on $H$ of members of asteroid \emph{families} in the Main Belt \citep[\eg][]{karin}.
+radiation leads to a net force that is a fraction $g_Y(1-q)$ of the
+incident radiation pressure.  The Yarkovsky effect has been detected
+for several NEAs but no individual MBAs, with values $g_Y \approx 0.1$
+\citep[Yarkovsky effect and data are review by][]{yarkovsky}.
+\edited{A value of $g_Y\approx0.1$ is also the largest that can be obtained
+  from a model of a spherical body with uniform surface properties and
+  Lambertian emission.} This is also consistent with detections of the
+Yarkovsky effect by examining the $\Delta a$ dependence on $H$ of
+members of asteroid \emph{families} in the Main Belt
+\citep[\eg][]{karin}.  \edited{We will assume an RMS value of $g_Y=0.05,$
+the influence of the surface absorption coefficient being $<1.$}
 
@@
    \frac{\Delta\phi}{\sigma_\phi} \approx \frac{ 3 g_Y \Psi
-     T^2/2}{\sqrt{3 T^3 \nIphi}} \approx \sqrt{T/10\,{\rm yr}} \times 10^{0.2(H-20.2)}
+     T^2/2}{\sqrt{3 T^3 \nIphi}} \approx \sqrt{T/10\,{\rm yr}} \times 10^{0.2(H-12.7)}
      \label{eq:yarkovsky}
 \end{equation}
-if we assume $\nIphi$ from the current knowledge of asteroid masses.
-There is thus a break point near $H=20$, or $\approx100$~m diameter: for MBAs larger than this, the ephemeris uncertainties from the Yarkovsky effect will be smaller than those from mutual gravitational encounters on decade timescales.
-This includes all MBAs detected by Gaia and nearly all detectable by LSST.
-Below this size, the Yarkovsky displacement grows larger in a decade and has a better chance of being detectable in the presence of Brownian motion.
-Distinguishing the steady quadratic growth of the Yarkovsky $\Delta\phi$ from the random-walk growth of Brownian motion to the same final size requires observational S/N and time resolution significantly better than is needed for detecting either one individually.
+\edited{if we assume $\nIphi$ from the LSST-era knowledge of asteroid
+  masses.   For MBAs fainter than $H\approx12.7,$ or
+  $\lesssim7$~km in diameter, the RMS uncertainty in position from
+  Yarkovsky forces exceeds the Brownian motion noise.   With current-day knowledge
+  of asteroid masses, the threshold for Yarkovsky dominance is a bit
+  fainter than $H=12.7.$  The optimist's view of this is that Brownian
+  motion will not interfere with attempts to measure the Yarkovsky
+  effect on individual km-scale or smaller MBAs.  The pessimists view is that uncertainty in
+ the  Yarkovsky effect will limit ephemeris accuracy before Brownian
+  motion will.  In either viewpoint, the steady $T^2$ accumulation of
+  $\Delta\phi$ displacements from the Yarkovsky effect is
+  distinguishable from the $T^{3/2}$ random walk of Brownian motion as
+  data quality and duration improve, so one separate these effects to
+  some extent when fitting an ephemeris---\citet{negin} quantify the
+  ability to distinguish the Yarkovsky effect from mutual encounters.}
 
@@
 \eqq{eq:yarkovsky} shows that the temporally coherent Yarkovsky effect
-can dominate over Brownian motion for bodies up to 10's or 100's of km
+can dominate over Brownian motion for bodies up to hundreds of km
 in diameter, which is why the technique is successful.
@@
 for the accuracy of imaging astrometry for the foreseeable future.
-Occultation-based astrometry is much more accurate (but has fewer possible measurements per tracer), and would be significantly impacted by Brownian motion until $\gtrsim1000$ MBAs have mass estimates at modest $S/N$ level, although the precise impact will depend on whether the signals one is seeking are parallel or orthogonal to the kind of variations caused by Brownian motion.
-For objects $\gtrsim100$~m in diameter, uncertainties from mutual encounters are larger than uncertainties from the Yarkovsky effect over decade timescales, at present levels of knowledge of asteroid masses.
+Occultation-based astrometry is much more accurate (but has fewer
+possible measurements per tracer), and would be significantly impacted
+by Brownian motion until $\gtrsim1000$ MBAs have mass estimates at
+modest $S/N$ level, although the precise impact will depend on whether
+the signals one is seeking are parallel or orthogonal to the kind of
+variations caused by Brownian motion.  \edited{For objects
+  $\lesssim7$~km in diameter, uncertainties from mutual encounters are
+  smaller than uncertainties from the Yarkovsky effect over decade
+  timescales, at near-future levels of knowledge of asteroid masses.
+  Yarkovsky forces must be included in dynamical models in order to
+  reach the precision levels where Brownian motion becomes a limitation.}
 

### 2025-08-19 response.txt
ready for revisions

Response to first referee report.

I thank the referee for their concise and insightful review of the paper.  The
main issue raised is that this paper relies heavily on the companion paper
analysing LSST constraints on asteroid masses from mutual encounters.  That
paper was submitted to PSJ some weeks before this one, and has since been
refereed, revised, accepted, and published in the journal, which I hope will
assuage this worry.

One of the revisions made to that paper addressed an issue also raised by this
referee, which was the rather high value of albedo assumed for the main-belt
asteroid population.  To address this concern for the other paper, we changed
our conversion from (well-observed) H values to (physically relevant) mass
substantially,  Instead of choosing a representative value of  both density and
albedo of the population, we surveyed the literature for individual main-belt
asteroids (MBAs) with mass determinations of any useful precision, then examined
the ratio $M/ 10^{-0.6H}$ directly.  The number of such bodies is actually
larger than the number for which both density and albedo have been determined
to similar precision, so is a simpler and slightly more accurate means of deriving
the "mass-to-light" conversion.  See Figure 1 and Section 2.1 of the published
paper.

The adopted result was that the distribution of mass-to-light
ratios of this (incomplete, heterogeneous, almost certainly biased) population
could be crudely characterized by taking a "high" and "low" value, which differ
by a factor of 5.  The main change made to this Brownian-motion paper is to
adopt these two values and recalculate the resulting integrals that yield the
estimated RMS 10-year uncertainties that are the key result of this paper.
These mass-to-light ratio changes appear as multiplicative scalings of the
previous result.  The thresholds for individually-detectable impulses in Section
3.2.3 have also been updated in concert with the updates to the
mass-determination paper.

Thus the numerical results presented in this paper have been adjusted, but as
per the referee's comments, the methodology remains unchanged.  Changed text is
in red in the draft, and figures that have changed have this noted in their
captions.

A few minor replies to the referee's comments are below.

> This report evaluates the manuscript "Brownian Motion of Main-Belt
> Asteroids on Human Timescales," which examines how gravitational
> interactions between asteroids introduce stochastic positional
> uncertainties over decadal timescales. The study provides a novel
> approach to understanding perturbative effects in asteroid
> ephemerides, with implications for high-precision astronomy.
> 
> The central hypothesis of the study is that the long-term orbital
> evolution of main-belt asteroids is affected by random gravitational
> impulses from other asteroids, leading to a Brownian motion-like
> positional drift. This effect is expected to be particularly
> significant over human timescales (10-100 years), impacting asteroid
> tracking and planetary defense efforts. The study proposes that this
> motion is primarily azimuthal and that current ephemeris models
> underestimate this stochastic contribution.

We did not intend to give the impression that Brownian motion is larger or more
important on decadal scales than over millennia.  Rather our purpose is to
quantify the effect of the Brownian motion on "real-time" dynamics of the solar
system, i.e. the development of ephemerides that fit obervations and adjust
parameters of the solar system's members and of the laws of gravitation.
Understanding the long-term dynamical consequences of the Brownian motion for
population evolution is beyond our scope.

> 
> The study is based on several key assumptions:
> 
> - Asteroid Encounters as Stochastic Events: The model assumes that
> gravitational encounters between asteroids can be treated as a series
> of independent random impulses.
> 
> - Negligibility of Non-Gravitational Effects: The study assumes that
> radiation forces (e.g., Yarkovsky effect) and thermal recoil forces
> have a minor influence compared to gravitational interactions for
> asteroids larger than ~100 m.
> 
> - Mass Uncertainty from LSST Data: The study uses projected LSST data
> to estimate mass uncertainties, heavily relying on Bernstein et
> al. (2025) for these calculations.
> 
> - Choice of Geometric Albedo: The author chooses the geometric albedo
> value that might, in this evaluator's judgment, be too "bright" to be
> representative of the entire population of asteroids in the main
> belt. The problem with that choice lies in the fact that, with that
> value, the mass distribution estimate would be misrepresented. A value
> of 0.15 might be more representative of the average, so I recommend
> its use.
>

See the note at top for a major revision to our estimation of asteroid masses
from their absolute magnitudes.

> In terms of methodology, the study employs a combination of
> theoretical modeling, statistical analysis, and observational
> constraints:
> 
> -Impulse-Based Orbital Perturbation Model: The author developed a
> model where asteroid encounters introduce small, random velocity
> perturbations that accumulate over time, following a Brownian motion
> framework.
> 
> -Empirical Scaling Relations: The study derives scaling relations for
> the expected positional variance over time.
> 
> -Comparison with Gaia and LSST Predictions: The paper estimates how
> well future astronomical surveys can track and mitigate these
> stochastic uncertainties.  Both the mathematical approach and its
> development are appropriate, and I have not detected any errors in
> them.
> 
> Regarding the results, the manuscript highlights some points that
> might be very interesting (the reasons for using "might" are explained
> below):
> 
> - the RMS azimuthal shift due to Brownian motion reaches ~2 km over 10
> years (~1 mas in angular uncertainty).
> 
> - Radial and vertical displacements are found to be significantly
> smaller, with the azimuthal motion dominating the effect.
> 
> - LSST mass uncertainty reductions could improve orbit predictions but
> do not eliminate Brownian motion noise.
> 
> Although the study presents an interesting and underexplored
> perspective on stochastic effects affecting asteroid motion and has a
> robust framework with an application of "Brownian motion" to celestial
> mechanics, it is fully justified to rely on a paper that is in
> preparation (Bernstein et al., 2025) and has clearly not gone through
> the peer review process. This is the reason for using "might" in the
> results section, making it difficult to analyze the final results. If
> the author can improve on this point, the paper would be an
> interesting contribution to the field.
>

### 2025-08-20 response.txt
Submission 2

@@
 the "mass-to-light" conversion.  See Figure 1 and Section 2.1 of the published
-paper.
+paper on mass estimation for details.
 
-The adopted result was that the distribution of mass-to-light
-ratios of this (incomplete, heterogeneous, almost certainly biased) population
-could be crudely characterized by taking a "high" and "low" value, which differ
-by a factor of 5.  The main change made to this Brownian-motion paper is to
-adopt these two values and recalculate the resulting integrals that yield the
-estimated RMS 10-year uncertainties that are the key result of this paper.
-These mass-to-light ratio changes appear as multiplicative scalings of the
-previous result.  The thresholds for individually-detectable impulses in Section
-3.2.3 have also been updated in concert with the updates to the
-mass-determination paper.
+The adopted result of plotting the objects with known H and M was that the
+distribution of mass-to-light ratios of this (incomplete, heterogeneous, almost
+certainly biased) population could be crudely characterized by taking a "high"
+and "low" value, which differ by a factor of 5.  The main change made to this
+Brownian-motion paper is to adopt these two values and recalculate the resulting
+integrals that yield the estimated RMS 10-year uncertainties that are the key
+result of this paper.  These mass-to-light ratio changes appear as
+multiplicative scalings of the previous result.  The thresholds for
+individually-detectable impulses in Section 3.2.3 have also been updated in
+concert with the updates to the mass-determination paper.
 
-Thus the numerical results presented in this paper have been adjusted, but as
-per the referee's comments, the methodology remains unchanged.  Changed text is
-in red in the draft, and figures that have changed have this noted in their
-captions.
+Thus the numerical results presented in this paper have been adjusted, and this
+has led to substantial revisions of the sections conveying results, the
+discussion sections, and the abstract.  We have changed the main results in
+Figure 3 to show both the high- and low-mass MBA cases, and added Table 1 to
+summarize the numerical results because there are now a multiplicity of cases to
+keep track of.  One important change to the conclusions is that this Brownian
+noise is important to consider by people analyzing upcoming Gaia astrometry on
+MBAs.
+
+Changed text is in red in the draft, and figures that have changed have this
+noted in their captions. As per the referee's comments, the methodology remains
+largely unchanged.  Throughout the text, we have made small changes to the
+wording of sections where we now see clarity can be improved, and have fixed
+some small typos and errors.
 
@@
 We did not intend to give the impression that Brownian motion is larger or more
-important on decadal scales than over millennia.  Rather our purpose is to
+important on decadal scales than over Myr or Gyr scales.  Rather our purpose is to
 quantify the effect of the Brownian motion on "real-time" dynamics of the solar
@@
 >
+
+Note in section 6.2, we calculate that Yarkovsky effects are significant in
+comparison to the Brownian noise for asteroids of diameter >10 km.
+
 > - Mass Uncertainty from LSST Data: The study uses projected LSST data

### 2025-08-20 main.tex
Submission 2

@@
 \newcommand{\eg}{\textit{e.g.\/}}
-\newcommand\edited[1]{{\color{blue} {#1}}}
+\newcommand\edited[1]{{\color{red} {#1}}}
 \newcommand\gary[1]{{\color{red} {\textbf{GMB}: #1}}}
@@
 \newcommand{\ma}{\ell}  % Mean anomaly
+\newcommand{\uas}{$\mu$as}
 
@@
 using the distribution of known MBA orbits, combined with analytical propagation
-into future positional uncertainties. At current
-levels of asteroid-mass knowledge, $\sigma_\phi$ rises to
-$\approx2$~km or $\approx1$~mas over $T=10$~yr, increasing as
-$T^{3/2},$  large enough to degrade many inferences from
-Gaia and LSST MBA data. LSST data will, however, improve MBA
-mass knowledge enough to lower this Brownian uncertainty by
-$\ge4\times.$  Radial (vertical) Brownian noise  at $T=10$~yr is
-$\approx7\times$ ($\approx45\times$) 
-lower than the azimuthal noise.  For full exploitation of Gaia and
+into future positional uncertainties---radial and vertical perturbations are
+much smaller. At current
+levels of asteroid-mass knowledge, \edited{we estimate $\sigma_\phi$ accumulates
+to 2--9~km, or 1--5~mas over $T=10$~yr,  increasing as
+$T^{3/2}.$  This must be taken into consideration for proper inferences from Gaia astrometry.}
+LSST data will, however, improve MBA
+mass knowledge enough to lower this Brownian uncertainty \edited{to 260--400~m,
+  or 140--210~\uas, which in many but not all cases could be small enough to
+  ignore in inferences using Gaia or LSST astrometry of MBAs.}
+% Radial (vertical) Brownian noise  at $T=10$~yr is
+% $\approx7\times$ ($\approx45\times$) 
+% lower than the azimuthal noise.
+For full exploitation of Gaia and
 LSST MBA data, ephemeris models should include the
-$\approx1000$~largest asteroids as active bodies with free masses,
+\edited{$\approx10,000$ largest} asteroids as active bodies with free masses,
 even if not all are well constrained.  This will correctly propagate
-the uncertainties from these 1000 sources' deflections into desired
+the uncertainties \edited{in these bodies' gravity} into desired
 inferences. The RMS value of deflections from less-massive MBAs is
-then just $\sigma_\phi\approx60$~m or 30~$\mu$as, small enough to ignore until occultation-based position data become ubiquitously available for MBAs.
+then just \edited{$\sigma_\phi\approx17$--60~m, or 9--40~\uas, depending on the
+  mean mass of MBAs,} small enough to ignore until occultation-based position
+data become ubiquitously available for MBAs. 
 \end{abstract}
@@
 the non-gravitational forces that drive many forms of planetary
-migration, such as the maintainence of the near-Earth asteroid (NEA)
+migration, such as the maintenance of the near-Earth asteroid (NEA)
 population \citep[reviewed by][]{yarkovsky}.  While most applications of precision orbit determination
@@
 \section{Calculation overview}
-We aim to calculate the Brownian motion variance to an accuracy of a
+We aim to derive the Brownian motion variance to an accuracy of a
 factor $\approx 2.$
-Greater accuracy in the calculation is not warranted
-since some of the principal inputs have substantial uncertainty.  The
+Greater accuracy in the \edited{methodology is not warranted
+since some of the principal inputs have substantial uncertainty, \eg\ the}
 number of MBAs vs mass is 
 uncertain because of unknowns in the conversion
-from $H$ to mass; the status of future observations and $\sigma_M$ is
+from $H$ to mass, and the status of future observations and $\sigma_M$ is
 also unclear.  We are hence justified in retaining only the
@@
 normal to the initial orbital plane, and $\phi=z=0$ toward the
-perhelion of the initial orbit, \ie\ both the initial ascending node
+perihelion of the initial orbit, \ie\ both the initial ascending node
 $\Omega_0$ and longitude of perihelion $\lop_0$ equal to zero.  The
@@
   \\
-  & = & \int dH \frac{dN}{dH} \int dv
+  & = & \int dH \frac{dN}{dH} \int dv\,\edited{d\theta\,d\phi}
         \frac{dn}{dA\,dv\,d\theta\,d\phi} \hat b_\gamma^2(\theta,\phi)
@@
 
-In the third line, we execute the integral over $b$ and introduce the
-average geometry factors $\langle \hat b^2_\gamma\rangle$ of encounters as
-a function of $v$.  It must be true that
+In the third line, we execute the integral over $b,$ and \edited{condense the integrals
+over the impulse direction $(\theta,\phi)$ into the angle-averaged
+geometry factors $\langle \hat b^2_\gamma\rangle$ of encounters as
+a function of $v$.  These will be derived from the actual MBA orbit
+distributions.} It must be true that
 \begin{equation}
@@
 \eqq{eq:nI2} reduces the calculation to a double integral over the
-deflector $H$ and the tracer-deflector relative velocity $v.$  The
+deflector's $H$ and the tracer-deflector relative velocity $v.$  The
 constituents of this calculation are estimated as follows.
@@
 \label{sec:mbaprops}
-We crudely approximate the population as having a geometric albedo of 0.25,
-density of 2500~kg~m$^{-3},$ and spherical shapes, which leads to the
-conversion
-\begin{equation}
-  M(H) = 1.2\times10^{-17} M_\odot \times 10^{-0.6(H-15)}
-\end{equation}
-
+We \edited{begin with} the $H$ distribution of potential deflectors,
+\edited{since this is directly observable.}.
+Figure~\ref{fig:counts} plots our assumptions for $dN/dH.$  The MPC
+catalog is likely close to complete in all three regions for $H\le17,$
+so we will use the MPC counts directly in this regime.  Note that the
+three regions have similar shapes of $dN/dH.$  For $H>17,$ we adopt
+the functional form for the cumulative MBA counts given by
+initial LSST predictions \citep{lsstbook}:
+\begin{equation}
+  N(<H) \propto \frac{10^{0.43(H-15.7)}}{10^{0.18(H-15.7)} +
+    10^{-0.18(H-15.7)}},
+\label{eq:lsstmba}
+\end{equation}
+normalizing this curve to the observed $H<17$ counts for each region.
 \begin{figure}
@@
 \end{figure}
-We also require the $H$ distribution of potential deflectors.
-Figure~\ref{fig:counts} plots our assumptions for $dN/dH.$  The MPC
-catalog is likely close to complete in all three regions for $H\le17,$
-so we will use the MPC counts directly in this regime.  Note that the
-three regions have similar shapes of $dN/dH.$  For $H>17,$ we adopt
-the functional form for the cumulative MBA counts given by
-initial LSST predictions \citep{lsstbook}:
-\begin{equation}
-  N(<H) \propto \frac{10^{0.43(H-15.7)}}{10^{0.18(H-15.7)} +
-    10^{-0.18(H-15.7)}},
-\label{eq:lsstmba}
-\end{equation}
-normalizing this curve to the observed $H<17$ counts for each region.
+
+\edited{Next we require a mean mass of MBAs with a given $H,$ ignoring the fact
+  that there will be a substantial spread of $M$ vs $H$.  While one can
+  calculate a mass under assumptions of a Lambertian spherical body, a known
+  geometric albedo, and a known mean density, each of these assumptions would be
+  inaccurate and based on a limited set of well-measured MBAs.  It is more
+  effective to examine $M(H)$ directly from the slightly-less-limited set of
+  MBAs with usefully constrained masses.  If albedo, density, and shape vary
+  weakly with $H$, we will expect a relation of the form
+  \begin{equation}
+        M(H) = M_{10} 10^{-0.6(H-10)}.
+        \label{m10}
+      \end{equation}
+The distribution of $M_{10}$ values for MBAs with masses found in the literature
+is presented in Figure~1 and Section~2.1 of \citet{negin}.  We adopt their
+selection of a ``high-mass-to-light'' value of
+$M_{10}=2\times10^{17}\,\text{kg}=10^{-13} M_\odot$ and a ``low-$M/L$'' value of
+$M_{10}=4\times10^{16}\,\text{kg}=2\times10^{-14} M_\odot$ as providing a rough
+estimate of the range that the true mean $M_{10}$ might take.}
+%       We crudely approximate the population as having a geometric albedo of 0.25,
+% density of 2500~kg~m$^{-3},$ and spherical shapes, which leads to the
+% conversion
+% \begin{equation}
+%   M(H) = 1.2\times10^{-17} M_\odot \times 10^{-0.6(H-15)}
+% \end{equation}
 
@@
 each deflector MBA's encounters, and the measurement errors on its tracers.  These
-factors are independent of the mass of the deflector.  We will
-assign a value $\sigma_M$ to the RMS uncertainty in MBA mass
-attainable from following a few individual tracers, and calculate
-the Brownian-motion noise as a function of $\sigma_M.$
-
-\begin{figure}
-  \plottwo{sigmaI.pdf}{sigmaM.pdf}
-  \caption{Forecasts of uncertainties derived from LSST tracking of the known MBAs.
-The left-hand plot shows the distribution of (log) uncertainties $\sigma_I$ on the amplitude of the impulse applied to the tracer MBA, incorporating forecasted LSST observations and marginalizing over the tracer's initial state vector.
-The right-hand side shows the distribution of the uncertainties on the mass, $\sigma_M=(bv/2)\sigma_I,$ for deflector MBAs expected from LSST.
-The histograms show the results from using only the single tracer yielding the lowest $\sigma_M$ for that deflector; combining the 10 best-$\sigma_M$ tracers; or combining all impulses at $b<0.03$~AU.
-The RMS value of $\sigma_M$ over the deflector population is given in the legend.
-Masses are in solar units, and velocities in units of the circular velocity of the deflector.}
-  \label{fig:sigmaI}
-\end{figure}
+factors are independent of the mass of the deflector.
 
@@
 observations to estimate the impulse (and mass) uncertainties from
-analyzing LSST's data on mutual encounters of the known asteroids. The
-tracer MBAs for those events are essentially an unbiased selection
-from the known MBAs, and the left-hand plot of Figure~\ref{fig:sigmaI}
-shows the distribution of the uncertainty $\sigma_I$ derived from each
-tracer's LSST data after marginalizing over its initial state vector.
-This histogram marginalizes over the apparent magnitude and observing
-conditions of the tracer, the date of the encounter, and the direction
-$\bhat$ of the impulse.  The right-hand plot of
-Figure~\ref{fig:sigmaI} forecasts the distribution of $\sigma_M$ that
-we can expect for deflectors.  This forecast is made as follows: we
-treat each of the 5000 tracer MBAs that we have selected from each
-region as deflectors, since we have a list of all of their encounters
-with known MBAs with $b<0.03$~AU.  For each such encounter, we select
-at random one $\sigma_I$ from the distribution for LSST data on known
-MBAs forecasted by \citet{negin} (Figure~\ref{fig:sigmaI}, left), and
-calculate $\sigma_M=(bv/2)\sigma_I.$  For each deflector, we then
-calculate (a) the individual encounter with lowest $\sigma_M$; (b) the
-combined constraint $\left(\sum \sigma_{M,i}^2\right)^{-1/2}$ from the
-10 lowest $\sigma_M$ encounters for that deflector; and (c) the
-combined constraint for all $b<0.03$~AU encounters (typically
-$\sim1000$ of them).  The right-hand plot of Figure~\ref{fig:sigmaI}
-gives the histograms of these three measures of $\sigma_M.$  We see
-that most of the available information on any given deflector's mass
-comes from its 10 most-informative encounters.  It should not be a
-computational barrier to solve for the masses of $O(10^4)$ MBAs by including the deflections they impart on $O(10^5)$ other MBAs in an ephemeris fit.
-The RMS $\sigma_M$ using LSST data from the 10 most-informative tracer encounters with a given deflector is estimated as $\sigma_M=2.6\times10^{-14}M_\odot.$ There are $\approx450$ MBAs with mass above this forecasted RMS $\sigma_M.$  This is a conservative estimate in that we ignore any observational information before LSST, and we also do not consider the information increase that will come from tracking the $\approx10$-fold increase in number of known MBAs that LSST will enable.
+analyzing LSST's data on mutual encounters of the known asteroids.
+\edited{ They find
+that the median $\sigma_M$ for MBAs depends substantially on the quality of
+pre-LSST astrometry of tracer bodies, varying from $10^{-13.5} M_\odot$ if no
+pre-LSST data are used, to $10^{-14.1} M_\odot$ if pre-LSST data fully determine
+the tracer state vector at the start of LSST (see their Figure~4 and Section~5).  These values assume that the 20
+most informative tracers of a given target asteroid are combined to find the
+best estimate of the deflector mass.  We will adopt a value of
+$\sigma_M=10^{-14} M_\odot$  as the nominal situation post-LSST. This is on the
+optimistic side of the value of pre-LSST data, but also ignores information that
+might be gained from mutual encounters with the $>10^7$ new MBAs that LSST is
+expected to discover and have available as tracer particles.}
+
+% We will
+% assign a value $\sigma_M$ to the RMS uncertainty in MBA mass
+% attainable from following a few individual tracers, and calculate
+% the Brownian-motion noise as a function of $\sigma_M.$
+
+% \begin{figure}
+%   \plottwo{sigmaI.pdf}{sigmaM.pdf}
+%   \caption{Forecasts of uncertainties derived from LSST tracking of the known MBAs.
+The left-hand plot shows the distribution of (log) uncertainties $\sigma_I$ on the amplitude of the impulse applied to the tracer MBA, incorporating forecasted LSST observations and marginalizing over the tracer's initial state vector.
+The right-hand side shows the distribution of the uncertainties on the mass, $\sigma_M=(bv/2)\sigma_I,$ for deflector MBAs expected from LSST.
+The histograms show the results from using only the single tracer yielding the lowest $\sigma_M$ for that deflector; combining the 10 best-$\sigma_M$ tracers; or combining all impulses at $b<0.03$~AU.
+The RMS value of $\sigma_M$ over the deflector population is given in the legend.
+Masses are in solar units, and velocities in units of the circular velocity of the deflector.}
+%   \label{fig:sigmaI}
+% \end{figure}
+
+% The
+% tracer MBAs for those events are essentially an unbiased selection
+% from the known MBAs, and the left-hand plot of Figure~\ref{fig:sigmaI}
+% shows the distribution of the uncertainty $\sigma_I$ derived from each
+% tracer's LSST data after marginalizing over its initial state vector.
+% This histogram marginalizes over the apparent magnitude and observing
+% conditions of the tracer, the date of the encounter, and the direction
+% $\bhat$ of the impulse.  The right-hand plot of
+% Figure~\ref{fig:sigmaI} forecasts the distribution of $\sigma_M$ that
+% we can expect for deflectors.  This forecast is made as follows: we
+% treat each of the 5000 tracer MBAs that we have selected from each
+% region as deflectors, since we have a list of all of their encounters
+% with known MBAs with $b<0.03$~AU.  For each such encounter, we select
+% at random one $\sigma_I$ from the distribution for LSST data on known
+% MBAs forecasted by \citet{negin} (Figure~\ref{fig:sigmaI}, left), and
+% calculate $\sigma_M=(bv/2)\sigma_I.$  For each deflector, we then
+% calculate (a) the individual encounter with lowest $\sigma_M$; (b) the
+% combined constraint $\left(\sum \sigma_{M,i}^2\right)^{-1/2}$ from the
+% 10 lowest $\sigma_M$ encounters for that deflector; and (c) the
+% combined constraint for all $b<0.03$~AU encounters (typically
+% $\sim1000$ of them).  The right-hand plot of Figure~\ref{fig:sigmaI}
+% gives the histograms of these three measures of $\sigma_M.$  We see
+% that most of the available information on any given deflector's mass
+% comes from its 10 most-informative encounters.  It should not be a
+% computational barrier to solve for the masses of $O(10^4)$ MBAs by including the deflections they impart on $O(10^5)$ other MBAs in an ephemeris fit.
+% The RMS $\sigma_M$ using LSST data from the 10 most-informative tracer encounters with a given deflector is estimated as $\sigma_M=2.6\times10^{-14}M_\odot.$ There are $\approx450$ MBAs with mass above this forecasted RMS $\sigma_M.$  This is a conservative estimate in that we ignore any observational information before LSST, and we also do not consider the information increase that will come from tracking the $\approx10$-fold increase in number of known MBAs that LSST will enable.
 
@@
 Examining the $\sigma_I$ distribution forecasted for LSST data on
-tracers by \citet{negin}, we find that for impulses occurring near the midpoint of the survey, $\approx80\%$ of all cases yield $\sigma_I<10^{-4}\,{\rm m}\,{\rm
-  s}^{-1},$ some an order of magnitude lower.  We conservatively
+tracers by \citet{negin} \edited{(their Figure 3)}, we find that for impulses occurring near the midpoint of the survey, $\approx80\%$ of all cases yield $\sigma_I<10^{-4}\,{\rm m}\,{\rm
+  s}^{-1},$ some an order of magnitude lower, \edited{even when pre-LSST information on
+the tracer MBA is uninformative.}  We conservatively
 assume that any encounter producing  an impulse $>5\times$ this value, $I_{\rm det}=5\times 10^{-4}\,{\rm m}\,{\rm
   s}^{-1},$ can be used to directly constrain the relevant deflector and include it in the ephemeris model.
-When normalized by $v_c,$ we have $I_{\rm  det}=10^{-8}$ at 10\% accuracy across the main belt.
+When normalized by $v_c,$ we have \edited{$I_{\rm  det}<3\times10^{-8}$ across the main belt.}
 
@@
 
-\section{Results and conclusions}
+\section{Results}
 \label{sec:results}
@@
 \item The population has $\langle e^2 \rangle=0.03,$ which is the value for the cataloged MBAs of the Inner and Middle belts, and a slight overestimate for the Outer belt.
-\item The $5\sigma$ impulse size $I_{\rm det}$, at which we deem it possible to isolate the effect of a single deflector on a tracer and hence usefully include that deflector's mass in the ephemeris model, is taken to be $I_{\rm det} = 10^{-8}v_c.$ In practice this will depend on the geometry of the impulse and on the quality of the tracer's observations.
-But we find that changing $I_{\rm det}$ by a factor of 10 alters the noise from unmodeled impulses by $<20\%.$
-  \item We will vary $\sigma_M,$ the RMS uncertainty on the mass of modelled asteroids, such that the number $N$ of MBAs  with $M>\sigma_M$ varies from 100 to 10,000.
-The number of MBAs included in the ephemeris model would be $\approx N.$
+\item The $5\sigma$ impulse size $I_{\rm det}$, at which we deem it possible to isolate the effect of a single deflector on a tracer and hence usefully include that deflector's mass in the ephemeris model, is taken to be $I_{\rm det} = 3\times10^{-8}v_c.$ In practice this will depend on the geometry of the impulse and on the quality of the tracer's observations.
+But we find that changing $I_{\rm det}$ by a factor of 10 alters the noise from unmodelled impulses by $<20\%.$
+\item We will vary $\sigma_M,$ the RMS uncertainty on the mass of modelled asteroids, such that the number $N$ of MBAs  with $M>\sigma_M$ varies from 200 to 10,000.
+The number of MBAs included in the ephemeris model would be $\approx N.$  \edited{We also examine the forecasted LSST results at $\sigma_M=10^{-14}M_\odot.$}
+  \item \edited{We consider the two cases of high or low mean $M/L$ ratio for
+      MBAs, which differ by a factor of 5 in asteroid mass at fixed $H.$ This
+      yields nearly a factor 5 difference in the RMS Brownian motion at a given $N.$}
 \end{itemize}
@@
   \centering
-  \includegraphics[width=0.6\textwidth]{result.pdf}
-  \caption{Each line shows the forecast of the standard deviation of the
-    azimuthal position of a Middle Main Belt asteroid caused by discrete
-    encounters over 10~years with absolute magnitude $H$ greater than the value on
-    the (lower) $x$ axis, \ie\ the cumulative noise from all MBAs
-    fainter (smaller) than the plotted $H_{\rm min.}$  The upper axis gives the $M_{\rm max}$ equivalent to $H_{\rm min}.$
-.  The left axis gives RMS values in meters, the right
+  % \includegraphics[width=0.6\textwidth]{result.pdf}
+  \includegraphics[width=0.48\textwidth]{result_loML.pdf}
+  \includegraphics[width=0.48\textwidth]{result_hiML.pdf}
+  \caption{Each curve shows the forecast of the standard deviation of the
+    azimuthal position of a Middle Main Belt asteroid \edited{caused by discrete
+    encounters, over 10~years, with other MBAs that are smaller/fainter than the
+    $H_{\rm min}$ value on the (lower) $x$ axis.  The upper axis gives the $M_{\rm max}=M(H_{\rm min}).$}
+The left axis gives RMS values in meters, the right
 the apparent shift in milliarseconds for a heliocentric observer.
+\edited{The left and right panels differ in their assumptions about the scaling between $H$ value and mass for MBAs, as noted within each frame.}
     Each line is for a different assumption about the RMS uncertainty $\sigma_M$ in the masses of individual asteroids gained from high-$S/N$ observations of mutual encounters.
-Each star marks the dividing point between the $N$ MBAs with $M>\sigma_M,$ and the smaller bodies with no useful measure of their mass ($M<\sigma_M$).
-The light dotted lines mark the level of azimuthal noise expected from the bodies beyond the most massive $N,$ \ie\ the RMS that an ephemeris model including $N$ bodies would not properly model.
-    The black dashed curve is a conservative estimate of what can be achieved using LSST data to constrain larger MBA's masses.
-Radial and vertical RMS positions are factors of $\approx 7$ and 45 lower than the azimuthal noise, respectively.}
+Each star marks the dividing point between the $N$ MBAs with $M>\sigma_M,$ and
+the smaller bodies with no useful measure of their mass ($M<\sigma_M$). The
+light dotted lines mark the level of azimuthal noise from deflections only by
+the bodies with $M<\sigma_M.$ 
+The black dashed curve is a conservative estimate of what can be achieved using LSST data to constrain MBA's masses.
+Radial and vertical RMS positions are factors of $\approx 7$ and 45 lower than the azimuthal noise, respectively.
+\edited{This figure has been split into two panels for two different assumptions about mean MBA mass.}}
   \label{fig:result}
@@
 
-Figure~\ref{fig:result} plots the resulting noise level $\sigma_\phi=\sqrt{\Var(\phi)}$ of the calculation for the Middle regions of MBAs, showing both the linear RMS displacements on the left axis and the angular displacements on the right axis.
+Figures~\ref{fig:result} plot the resulting noise level $\sigma_\phi=\sqrt{\Var(\phi)}$ of the calculation for the Middle regions of MBAs, showing both the linear RMS displacements on the left axis and the angular displacements on the right axis.
 Each point on the curve is the $\sigma_\phi$ attributable to deflector MBAs that are smaller (fainter) than the $H$ value on the lower axis, or the mass on the top axis.
@@
 Different color lines show the results for the values of $N$ and $\sigma_M$ as marked astride each line.
+\edited{The left-hand plot assumes the lower value of $M_{10},$ the right-hand panel the higher value---note the shift of the upper axis scale.}
 Notable features of these result are:
@@
 that determines the Brownian motion is dropping.
+\item \edited{The $5\times$ larger mass at fixed $H$ translates fairly directly
+    into a $5\times$ higher RMS Brownian motion at the same $H$ cutoff.  The
+    higher $M_{10}$ also means, however, that $\approx2\times$ more (and
+    fainter) asteroids will attain useful constraints on their masses from LSST
+    observations of mutual interactions.  This ameliorates the increase in
+    unmodelled Brownian motion from the higher $M/L.$}
 \end{itemize}
 
-We take as representative of the present state of the art the case
-where the $N=100$ most massive MBAs (aside from Ceres and Vesta) have
-masses determined to $\sigma_M\approx 10^{-12.7} M_\odot$ by mutual
-encounters.  Our calculations suggest then that a 10-year ephemeris
-including these 100 MBAs will accrue RMS errors of $\approx 2$~km in
-the azimuthal position, or 1~mas in ecliptic longitude.
-RMS range and vertical ephemeris errors would be closer to 300 and 40~m, respectively.  We find that, roughly speaking, $\sigma_\phi$
-drops with the number of modeled MBAs as $1/N.$
-
- In the next decade, LSST will significantly
-increase the number $N$ of deflector MBAs with
-usefully measured masses from mutual encounters by tracking hugely
-more potential tracers and lowering $\sigma_M.$  The conservative estimate given in Section~\ref{sec:mbaprops} is that LSST will lower $\sigma_M$ to $\approx10^{-13.7},$ which our calculations indicate lowers the $\sigma_\phi$ values to $\approx500$~m or 0.25~mas.
+We take as roughly representative of the present state of the art the case
+where the \edited{$N=200$} brightest and most massive MBAs (aside from Ceres and Vesta) have
+masses determined \edited{to $>1\sigma$ accuracy}
+% .  to $\sigma_M\approx 10^{-12.7} M_\odot$
+by mutual encounters.  Our calculations then suggest that a 10-year ephemeris
+including these 200 MBAs will accrue RMS errors in the range \edited{2--8.5~km in
+the azimuthal position, or 1.0--4.4~mas in ecliptic longitude.
+This wide range arises from uncertainty in what the actual masses of the MBAs are.
+These results are given as the ``Total RMS'' in the column of Table~\ref{tab:rms} labelled ``Current.''}
+% RMS range and vertical ephemeris errors would be closer to 300 and 40~m, respectively.
+
+\begin{deluxetable}{lcccc}
+  \tablewidth{0pt}
+\tablecaption{Forecasted RMS Brownian motion}
+\tablehead{\colhead{} & \colhead{Current} & \colhead{LSST} &   \colhead{Future 1} & \colhead{Future 2}}
+\startdata
+$\sigma_M/M_\odot$ & $10^{-12.8}$--$10^{-12.1}$ & $10^{-14.0}$
+& $10^{-15.1}$--$10^{-14.4}$ 
+& $10^{-15.8}$--$10^{-15.1}$ \\
+Number $M>\sigma_M$ & 200 & 850--1700 & 3000 & 10,000 \\
+Total RMS & 2.0--8.5~km & 260--400~m & 50--240~m & 23--110~m \\
+& 1.0--4.4~mas & 140--210~\uas & 26--120~\uas & 12--60~\uas \\[5pt]
+$M<\sigma_M$ only & 1.2--5~km & 130--270~m & 35--170~m & 17--80~m \\
+& 0.6--2.8~mas & 70--140~\uas & 20--90~\uas & 9--40~\uas
+\enddata
+\label{tab:rms}
+\tablecomments{\edited{This table is new.} Each column gives the forecasted RMS positional uncertainty after 10~years from caused by deflections by asteroids of unknown and uncertain mass.
+Each column assumes either a total number $N$ of MBAs with mass $M$ above the mean MBA mass uncertainty $\sigma_M$, or in the case of LSST forecasts, assumes a value of $\sigma_M.$ The range is due to uncertainty in the mean mass of MBAs at a fixed $H$ value.
+The ``Total'' rows are the RMS of this Brownian motion caused by the $N$ bodies with mass known to accuracy $\sigma_M$, plus the remainder of the MBA population at $M<\sigma_M$.
+The bottom two rows include only the latter population's effects.}
+\end{deluxetable}
+%
+\edited{ The next curve in Figures~\ref{fig:result} and column in Table~\ref{tab:rms} adopt our estimate that 10-year LSST astrometry will reduce $\sigma_M$ to $\approx10^{-14}M_\odot,$ putting 800--1700 MBAs at $M>\sigma_M,$ depending on the $M/L$ ratio.
+[LSST should also greatly reduce the factor-5 uncertainty that we have assigned to $M/L$!]  The total RMS Brownian motion from ignorance of MBA masses is reduced substantially, to 260--400~m or 140--210~\uas.}
+
+\edited{We also examine scenarios where 3000 or 10,000 MBAs have $>1\sigma$ mass
+  measurements available.  The plots show that as the number of bodies $N$ with
+  $M>\sigma_M$ increases, RMS Brownian motion $\sigma_\phi$ drops as $\approx1/N.$}
+
+% In the next decade, LSST will significantly
+% increase the number $N$ of deflector MBAs with
+% usefully measured masses from mutual encounters by tracking hugely
+% more potential tracers and lowering $\sigma_M$  \edited{toward $10^{-14}M_\odot,$}
+% % The conservative estimate given in Section~\ref{sec:mbaprops} is that LSST will lower $\sigma_M$ to $\approx10^{-13.7},$
+% which our calculations indicate lowers the $\sigma_\phi$ values to $\approx500$~m or 0.25~mas.
 
@@
 $\Delta\phi\approx35$~mas.  This is much larger than the
-stochastic $\sigma_\phi\approx1$~mas expected from unmodelled masses at the
+stochastic \edited{$\sigma_\phi < 5$~mas} expected from unmodelled masses at the
 current state of knowledge.  For primarily radial or vertical
@@
 
-An important question is: will the positional noise due to unmodelled
+\edited{The main question we wish to answer is:} will the positional noise due to unmodelled
 MBA encounters degrade inferences being made with measurements of
 tracer MBAs?  If the Brownian noise is well below other sources of measurement noise, then the answer must be ``no.''
-The per-epoch uncertainties on solar-system object astrometry from Gaia\footnote{We take one Gaia transit as an ``epoch'' here.} are reported to be as low as 0.3~mas for very bright sources ($G=13$), rising to $\approx10$~mas at its $G=21$ limit where most detections lie \citep[][Figure 6]{gaiass3}.
-The overall RMS residual to the orbit fits are 5~mas (this is in the ``along-track'' direction).
-Thus the Brownian motion of $\approx1$~mas we estimate for $\phi$ (over a 10-year period) with current asteroid-mass knowledge would be important for the single-epoch error budget of the brighter Gaia MBAs.
-With our forecasted LSST-era knowledge of asteroid masses ($\sigma_M\approx10^{-13.6},$ RMS azimuthal errors $\approx0.25$~mas), the situation will be improved in that only the brightest few MBAs will have per-epoch Brownian motion noise comparable to the per-epoch measurement noise.
+The per-epoch uncertainties on solar-system object astrometry from
+Gaia\footnote{We take one Gaia transit as an ``epoch'' here.} are reported to be
+as low as 0.3~mas for very bright sources ($G=13$), rising to $\approx10$~mas at
+its $G=21$ limit where most detections lie \citep[][Figure 6]{gaiass3}.  The
+overall RMS residual to the orbit fits are 5~mas (this is in the ``along-track''
+direction).  Thus the Brownian motion of \edited{1--5~mas we estimate for $\phi$
+  (over a 10-year period) with current asteroid-mass knowledge would be
+  important for the single-epoch error budget of many MBAs.}  With the
+forecasted LSST-era knowledge of asteroid masses, \edited{RMS azimuthal shifts
+  from MBA mass ignorance are reduced to $\lesssim 200$~\uas, such that only the
+  brightest few MBAs will have per-epoch Brownian motion noise comparable to the
+  per-epoch Gaia measurement noise.}
 
@@
 The timing of each occultation can locate the center of mass of the target MBA along its direction of motion, with most of the information available from objects yielding uncertainties of 50--100~m on this position.
-The Brownian noise (with LSST-level asteroid knowledge) over a 10-year baseline will in this case be comparable to the per-epoch measurement errors for most observations.
-Reducing the typical mass uncertainty of individual asteroids to $\sigma_M\approx 10^{-15.4} M_\odot$ would be needed ($N=3000$) to reduce the Brownian noise to negligible levels for this survey.
+The Brownian noise over a 10-year baseline, with LSST-level asteroid knowledge, will \edited{exceed the per-epoch measurement errors from occultations.
+Reducing the Brownian noise to be below or comparable to the occultation measurement errors will require some combination of reaching $1\sigma$ mass measurement for 3000-10,000 MBAs, and having the asteroids' $M(H)$ relation be near the lighter end of our nominal range.}
 
@@
 drop by a factor $\sqrt{M}.$   \citet{gaiafpr} report typically $\sim60$ epochs per asteroid in 66 months, suggesting $O(100)$ epochs per asteroid in the full 11-year observing period.
-At Gaia's faint end, this means that Brownian noise with present asteroid knowledge would have comparable impact to the measurement errors, and dominate inferences at the bright end.
-LSST will obtain several hundred epochs per asteroid.
-With an effective factor 10--20 reduction in the total measurement noise, and assuming $\sigma_\phi\approx0.25$~mas Brownian noise, the latter would become important for the brighter tracer MBAs in LSST data, but inferences with the millions of fainter objects discovered by LSST would still be limited by photon noise.
-The occultation program proposed by \citet{occultations} would obtain $<10$ measurements per decade for most MBAs and hence see less degradation of Brownian relative to measurement noise.
+\edited{In this case, we should be comparing the Brownian noise to $1/\sqrt{100}$ of Gaia's per-epoch uncertainties.
+This would mean that Brownian noise, at current asteroid mass knowledge, will dominate the error budget for measurements of most Gaia stars' properties, and with LSST-level asteroid mass knowledge, will still be a significant source of error for the brighter end of Gaia's catalog.}
+
+LSST will obtain several hundred epochs per asteroid, \edited{so in this
+  worst-case scenario of the Brownian noise correlations paralleling some
+  dynamical signal of interest, the Brownian noise is significant if
+  $\sigma_\phi \gtrsim (2\,\text{mas}/\sqrt{400}=100$~\uas.
+We have seen that the Brownian noise with LSST-level knowledge of MBA masses is comparable to this value, and therefore would be significant.
+The ``Future'' scenarios of mass knowledge could reduce Brownian noise to insignificance for all LSST inferences, especially if the true MBA $M/L$ is near our low estimates.}
+The occultation program proposed by \citet{occultations} would only obtain
+$\lesssim10$ measurements per decade for most MBAs, but its extremely small
+measurement errors \edited{mean it will remain important to incorporate the Brownian
+noise into inferences.}
 
@@
 These are quite distinct from the random walk in $\Delta\tau$ that is the dominant effect of Brownian motion, and therefore Brownian noise will cause less degradation to the Planet X inference than the per-epoch $\sigma_\phi$ would suggest.
-Another example would be a gravitational perturbation whose signature is primarily a nodal precession, which is a signal manifesting primarily as out-of-plane motions---we have seen that Brownian motion out of the orbital plane is $>40\times$ smaller than the $\phi$ component and grows at $T^{-1/2}$ instead of $T^{3/2},$ so the per-epoch $\sigma_\phi$ of Brownian noise will grossly overestimate its effect on the inference.
+Another example would be a gravitational perturbation whose signature is primarily a nodal precession, which is a signal manifesting primarily as out-of-plane motions---we have seen that Brownian motion out of the orbital plane is $>40\times$ smaller than the $\phi$ component and grows at $T^{1/2}$ instead of $T^{3/2},$ so the per-epoch $\sigma_\phi$ of Brownian noise will grossly overestimate its effect on the inference.
 
@@
 errors on a given tracer are highly correlated between epochs.  The
-full covariance matrix of the tracer's observations will be needed
+full covariance matrix of the Brownian model noise at the tracer's observations will be needed
 rather than typical practice of considering measurements as
@@
 A further difficulty in propagating unmodelled deflections into
-inferences is that the probability distribution of the position shifts
+inferences is that the probability distribution of the position shifts \edited{imposed on any given MBA}
 will be non-Gaussian, because the RMS is dominated by the
@@
 
-A better approach, however, would probably be to increase the number
+A better approach \edited{to propagating this mutual-encounter noise into parameter influence would be} to increase the number
 $N$ of MBAs considered as active bodies in the ephemeris fitting,
@@
 the ephemeris fit would be a substantial computational challenge,
-having $N=10^3$--$10^4$ should be manageable.
-The MBA-to-MBA perturbations can be adequately calculated using Newtonian dynamics instead of a fully relativistic treatment.
-In this approach, all of the covariances between different observations of the same tracer, and between tracers, will be captured in the analysis for deflections due to these $N$ deflectors.
-If we \emph{choose $N$ to be large enough that the Brownian noise from the unmodelled MBAs is negligible to the error budget,} then we need not explicitly propagate the Brownian noise.
-
-Figure~\ref{fig:result} shows us what $N$ would be required: the noise due to unmodelled deflectors is given by the point where each curve for a given $N$ departs from the upper-right envelope of the curves.
-Increasing the number of MBAs in the ephemeris from $N=100\rightarrow10^3\rightarrow10^4$ decreases the RMS displacement in 10~yrs due to the unmodelled MBAs from $\sigma_\phi=1300\rightarrow60\rightarrow10$~m, with the radial and vertical RMS being substantially lower.
-These correspond to 0.7, 0.03, and 0.005~mas, respectively.
-Incorporating $N=1000$ MBAs into the ephemeris model would render noise from the rest of the asteroids unimportant for nearly all uses of Gaia or LSST MBA data for inference.
-For occultation-based positions, one would probably want to have the $\approx3000$ most massive MBAs as active bodies with masses as free parameters in the ephemeris model in order to render encounters from unmodelled bodies as unimportant.
+having $N=10^3$--$10^4$ should be manageable.  The MBA-to-MBA perturbations can
+be adequately calculated using Newtonian dynamics instead of a fully
+relativistic treatment.  In this approach, all of the covariances between
+different observations of the same tracer, and between tracers, will be captured
+in the analysis for deflections due to these $N$ deflectors.  If we \emph{choose
+  $N$ to be large enough that the Brownian noise from the remainder of the MBAs is negligible to the error budget,} then we need not explicitly propagate the Brownian noise.
+
+\edited{The dotted lines in Figures~\ref{fig:result} and the bottom 2 rows of Table~\ref{tab:rms} show us what positional uncertainties arise purely from the effects of ignorance of MBA masses \emph{beyond} the brightest $N$ MBAs.} The noise due to unmodelled deflectors is given by the point where each curve for a given $N$ departs from the upper-right envelope of the curves.
+Increasing the number of MBAs in the ephemeris \edited{to 3000, say, decreases the angular noise from other MBAs down to 20--90~\uas, and unless the true $M/L$ is at the high end of our range, this could be safely neglected for inferences from Gaia and LSST astrometry.
+For occultation-based positions, with their 50-100~m accuracy, one would probably want to have the $\approx10,000$ most massive MBAs as active bodies with masses as free parameters in the ephemeris model in order to render encounters from unmodelled bodies as unimportant; perhaps $>10^4$ would be needed if MBA $M/L$ values are high.}
 
@@
 At first order, radiation pressure from the Sun and the reaction from reflected sunlight cause a constant radial force on the MBA.
-The ratio of this force to the gravitational force for an asteroid of radius $R,$ geometric albedo $q$, and density $\rho$ is $(1+q)\Psi,$ where
+The ratio of this force to the gravitational force for spherical Lambertian asteroid of radius $R,$ bolometric albedo $q$, and density $\rho$ is $(1+q)\Psi,$ where
+\edited{
 \begin{equation}
@@
    = \frac{3}{16\pi} \frac{L_\odot}{GM_\odot c R \rho} \approx
-   3\times10^{-10} \times 10^{0.2(H-16)},
-\end{equation}
-where the last expression adopts our standard values $q=0.25,
-\rho=2500\,{\rm kg}\, {\rm m}^{-3}.$ The primary observational consequence of
+   2\times10^{-11} \times 10^{0.2(H-10)},
+ \end{equation}
+ }
+where the last expression adopts \edited{a geometric albedo $q=0.15$ and 
+$\rho=2000\,{\rm kg}\, {\rm m}^{-3},$ which would yield our low-$M/L$ value of $M_{10}=2\times10^{-14}M_\odot.$}
+The primary observational consequence of
 this constant radial force is a time-independent, $H$-dependent 
@@
 This does not resemble
-any expected effect of Brownian motion, and also very difficult to
+any expected effect of Brownian motion, and is also very difficult to
 measure since radar ranging to MBAs is infeasible.  More important for
@@
 incident radiation pressure.  The Yarkovsky effect has been detected
-for several NEAs but no individual MBAs, with values $g_Y \approx 0.1$
+at $g_Y\sim0.1$ for several NEAs, but not for any individual MBAs
 \citep[Yarkovsky effect and data are review by][]{yarkovsky}.
-\edited{A value of $g_Y\approx0.1$ is also the largest that can be obtained
+A value of $g_Y\approx0.1$ is also the largest that can be obtained
   from a model of a spherical body with uniform surface properties and
-  Lambertian emission.} This is also consistent with detections of the
+  Lambertian emission \citep{tremaine}, \edited{and this maximal value requires a spin axis
+  normal to the orbital plane.} This is also consistent with detections of  the
 Yarkovsky effect by examining the $\Delta a$ dependence on $H$ of
 members of asteroid \emph{families} in the Main Belt
-\citep[\eg][]{karin}.  \edited{We will assume an RMS value of $g_Y=0.05,$
-the influence of the surface absorption coefficient being $<1.$}
+\citep[\eg][]{karin}.  We will assume an RMS value of $g_Y=0.05.$
 
@@
 Taking a case where the Yarkovsky force is in the prograde azimuthal direction, and using our system of units with $a=1$ and $v_c=1,$ this leads to a continuous power $dE/dt\approx g_Y\Psi$ which in turn implies $da/dt\approx2g_Y\Psi.$   The mean motion rate is hence $1-3 g_Y\Psi t$ and the accumulated shift in mean anomaly over time $T$ amounts to $\Delta\phi = -3 g_Y \Psi  T^2/2.$   In this case, a rough comparison of the accumulated $\Delta\phi$ from the Yarkovsky effect to the $\sigma_\phi$ of noise from unmodelled Brownian motion, after a time interval $T$, is
+\edited{
 \begin{equation}
@@
    \frac{\Delta\phi}{\sigma_\phi} \approx \frac{ 3 g_Y \Psi
-     T^2/2}{\sqrt{3 T^3 \nIphi}} \approx \sqrt{T/10\,{\rm yr}} \times 10^{0.2(H-12.7)}
+     T^2/2}{\sqrt{3 T^3 \nIphi}} \approx \sqrt{T/10\,{\rm yr}} \times 10^{0.2(H-12.5)}
      \label{eq:yarkovsky}
 \end{equation}
-\edited{if we assume $\nIphi$ from the LSST-era knowledge of asteroid
-  masses.   For MBAs fainter than $H\approx12.7,$ or
-  $\lesssim7$~km in diameter, the RMS uncertainty in position from
-  Yarkovsky forces exceeds the Brownian motion noise.   With current-day knowledge
-  of asteroid masses, the threshold for Yarkovsky dominance is a bit
-  fainter than $H=12.7.$  The optimist's view of this is that Brownian
+if we assume $\nIphi$ from the LSST-era knowledge of asteroid
+masses and take the low value for $M_{10},$ which yields the higher Yarkovsky/Brownian ratio.}
+For MBAs fainter than \edited{$H\approx12.5,$ or
+  $\lesssim10$~km in diameter,} the RMS uncertainty in position from
+Yarkovsky forces exceeds the Brownian motion noise.
+%  With current-day knowledge
+%  of asteroid masses, the threshold for Yarkovsky dominance is a bit
+%  fainter than $H=12.7.$
+The optimist's view of this is that Brownian
   motion will not interfere with attempts to measure the Yarkovsky
-  effect on individual km-scale or smaller MBAs.  The pessimists view is that uncertainty in
- the  Yarkovsky effect will limit ephemeris accuracy before Brownian
-  motion will.  In either viewpoint, the steady $T^2$ accumulation of
+  effect on individual km-scale or smaller MBAs.  The pessimist's view is that uncertainty in
+ the  Yarkovsky effect \edited{will damage ephemeris accuracy more than Brownian
+  motion will for $H>12.5.$}  In either viewpoint, the steady $T^2$ accumulation of
   $\Delta\phi$ displacements from the Yarkovsky effect is
   distinguishable from the $T^{3/2}$ random walk of Brownian motion as
-  data quality and duration improve, so one separate these effects to
+  data quality and duration improve, so one can separate these effects to
   some extent when fitting an ephemeris---\citet{negin} quantify the
-  ability to distinguish the Yarkovsky effect from mutual encounters.}
+  ability to distinguish the Yarkovsky effect from mutual encounters.
 
@@
 
-In conclusion, the Brownian motion from unmodelled or mis-modelled
+The Brownian motion from unmodelled or mis-modelled
 mutual impulses between MBAs causes RMS azimuthal displacements of
-$\approx2$~km (or 1 mas in angular position) over 10 years, with
+\edited{2--9~km, or 1--5 mas in observed position,} over 10 years, with
 current knowledge of the masses of large asteroids.  This is large
 enough to become a significant part of the error budget for many
-inferences from Gaia MBA astrometry.  The knowledge on large-MBA
+inferences from Gaia MBA astrometry.  \edited{The large spread of the estimate is primarily due to uncertainty in the masses of typical MBAs at a fixed absolute magnitude $H$.}
+
+The knowledge on large-MBA
 masses likely to be gained from LSST tracking of MBAs should reduce
-this by a factor of $\ge4,$ at which point the Brownian noise will be
-comparable to Gaia or LSST measurement uncertainties only for the
-brighter targets of each survey.  In all cases, the Brownian motion is
+this \edited{to 260--400~m, or 140--210~\uas, at which point the Brownian noise will be
+well below the single-epoch measurement uncertainties for all but the brightest
+Gaia targets and all LSST measurements.  This does not, however, necessarily
+render the Brownian noise insignificant, because it is correlated across years'
+worth of measurements for a given MBA whereas measurement errors are
+uncorrelated and average down.  It is, therefore, still important to propagate
+the uncertainty in mutual-encounter deflections through any Gaia or LSST
+inferences from precision MBA astrometry.  Full exploitation of the
+$\lesssim100$~m positional accuracy attainable through timing of MBA's
+occultations of Gaia stars will required improving our knowledge of MBA masses
+beyond the capabilities of the nominal 10-year LSST.}
+% comparable to Gaia or LSST measurement uncertainties only for the
+% brighter targets of each survey.
+In all cases, the Brownian motion is
 $\approx7\times$ smaller in the radial direction and $\approx45\times$
-in the vertical direction.  The RMS azimuthal and radial displacements
+in the vertical direction at $T=10$~yr.  The RMS azimuthal and radial displacements
 grow with time as $T^{3/2}$, and the vertical RMS as $T^{1/2}.$
@@
 Making inferences from MBA astrometry becomes more complicated when
-Brownian motion is a significant contributor to the error budget, but
-a good path forward is to do all model fitting by including the
-$\approx1000$ largest MBAs as active bodies in the ephemeris model and
+Brownian motion is a significant contributor to the error budget, \edited{because of the
+correlations across observations and the non-Gaussian distribution of
+displacements at any given observation.}
+A good path forward is to do all model fitting by including the 1000--10,000
+largest MBAs as active bodies in the ephemeris model and
 leaving their masses as free parameters.  This will constrain the
@@
 the other inferences that one is trying to make.  The noise
-contributed by the remaining millions of unmodelled MBAs would be just
-$\approx60$~m RMS in the azimuthal direction (about 30~$\mu$as), and a
-few $\mu$as or less in the other axes, small enough to ignore entirely
-for the accuracy of imaging astrometry for the foreseeable future.
-Occultation-based astrometry is much more accurate (but has fewer
-possible measurements per tracer), and would be significantly impacted
-by Brownian motion until $\gtrsim1000$ MBAs have mass estimates at
-modest $S/N$ level, although the precise impact will depend on whether
-the signals one is seeking are parallel or orthogonal to the kind of
-variations caused by Brownian motion.  \edited{For objects
-  $\lesssim7$~km in diameter, uncertainties from mutual encounters are
+contributed by the remaining millions of unmodelled MBAs \edited{beyond the top
+  10,000} would be just 17--80~m RMS in the azimuthal direction (9--30~\uas), and a
+few \uas\ or less in the other axes, small enough to ignore entirely
+for the accuracy of imaging astrometry for the foreseeable future \edited{and not
+dominate the measurement errors for
+occultation-based astrometry.}
+%is much more accurate (but has fewer
+%possible measurements per tracer), and would be significantly impacted
+%by Brownian motion until $\gtrsim1000$ MBAs have mass estimates at
+%modest $S/N$ level, although
+\edited{The precise impact that the Brownian noise will have on future
+inferences from  precision MBA astrometry will depend, first, on what the masses
+of the typical $8<H<15$ MBAs are, but also on whether
+the position-vs-time signals one is trying to infer are parallel or orthogonal to the kind of
+variations caused by Brownian motion.  For objects with $H>12.5$ 
+ (diameter $\lesssim10$~km),} uncertainties from mutual encounters are
   smaller than uncertainties from the Yarkovsky effect over decade
-  timescales, at near-future levels of knowledge of asteroid masses.
+  timescales, at LSST levels of knowledge of asteroid masses. For such MBAs,
   Yarkovsky forces must be included in dynamical models in order to
-  reach the precision levels where Brownian motion becomes a limitation.}
+  reach the precision levels where Brownian motion becomes a limitation.
 
